### 2021-04-26 manuscript.tex
Initial Overleaf Import

%%
%% This is file `./samples/longsample.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% apa7.dtx  (with options: `longsample')
%% ----------------------------------------------------------------------
%% 
%% apa7 - A LaTeX class for formatting documents in compliance with the
%% American Psychological Association's Publication Manual, 7th edition
%% 
%% Copyright (C) 2019 by Daniel A. Weiss <daniel.weiss.led at gmail.com>
%% 
%% This work may be distributed and/or modified under the
%% conditions of the LaTeX Project Public License (LPPL), either
%% version 1.3c of this license or (at your option) any later
%% version.  The latest version of this license is in the file:
%% 
%% http://www.latex-project.org/lppl.txt
%% 
%% Users may freely modify these files without permission, as long as the
%% copyright line and this statement are maintained intact.
%% 
%% This work is not endorsed by, affiliated with, or probably even known
%% by, the American Psychological Association.
%% 
%% ----------------------------------------------------------------------
%% 
\documentclass[man]{apa7}

\usepackage[american]{babel}
\usepackage{nicefrac}
\usepackage{todonotes}
\usepackage{epigraph}
% For sorting references from same year in order instead of on title
\newcommand{\SR}[1]{} %for EJ's bibtex file
\newcommand{\EJ}[1]{\todo[inline, color=green]{  #1 }}
\newcommand{\JR}[1]{\todo[inline, color=gray]{  #1 }}
\newcommand{\MK}[1]{\todo[inline, color=red]{  #1 }}

\usepackage{graphicx}
\graphicspath{ {images/} }

\usepackage{csquotes}
\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{bibliography.bib}

\title{Trust Me, I'm a Bayesian: Building Back Trust in Science Using Bayes}
\shorttitle{Bayes in Science Education}

\author{Joshua M. Rosenberg}
\affiliation{University of Tennessee, Knoxville}

\author{Marcus Kubsch}
\affiliation{IPN - Leibniz Institute for Science and Mathematics Education}

\author{Eric-Jan Wagenmakers}
\affiliation{University of Amsterdam}

\leftheader{Weiss}

\abstract{

Uncertainty is ubiquitous in science and this uncertainty takes many forms, but the rhetoric around scientific knowledge and negotiating standards of evidence is often portrayed as immutable. Adding to this problem around the limited recognition of the tentative nature of science, when scientists try to communicate their uncertainty, research has shown us that many—including both learners and experts—struggle to interpret the statistics that are often used to make inferences from scientific data. We argue that a Bayesian approach can offer a different foundation and language around a core but sometimes unsettling tenet and persistent feature, uncertainty. Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains. In this essay, we a) briefly introduce Bayes theorem and wider ideas around Bayesian epistemology, b) elaborate on the importance of probability and uncertainty in science and science education and review how people learn about these concepts and where they face difficulties, and c) provide conceptual and practical examples within the JASP statistical software of how Bayes could help to address these issues. We conclude with implications and directions for future research focused around bolstering trust in science outside of classroom settings (for which Bayesian approaches could serve as a crosscutting part of science literacy) and in pre-collegiate science classrooms, where Bayes can provide a coherent approach to engaging in numerous science practices such as arguing from evidence and analyzing and interpreting data in a principled and flexible way.
}

\keywords{APA style, demonstration}

\authornote{
   \addORCIDlink{Joshua M. Rosenberg}{0000-0003-2170-0447}
   \addORCIDlink{Marcus Kubsch}{0000-0000-0000-0000}
   \addORCIDlink{Eric-Jan Wagenmakers}{0000-0003-1596-1034} 
}

\begin{document}
\maketitle

\setlength{\epigraphwidth}{1.0\textwidth} % set for individual epigraph
\epigraph{It is seen in this essay that the theory of probabilities is at bottom only common sense reduced to calculus; it makes us appreciate with exactitute that which exact minds feel by a sort of instinct without being able ofttimes to give a reason for it . . . we shall see that there is no science more worthy of our meditations, and that no more useful one could be incorporated in the system of public instruction.}{Pierre-Simon Laplace, 1829}
\setlength{\epigraphwidth}{0.4\textwidth} % return to default for the next epigraphs
%\nocite{Laplace18291902} %[p. 196]

\EJ{This is a comment.}

Uncertainty is ubiquitous in science and this uncertainty takes many forms. For example, measurement uncertainty is present when there are differences (small or large) in the repeated measures of the same object or phenomenon \parencite{pls03} or when the object that is measured, like a plant, is inherently variable in its characteristics \parencite{pls03}. Another form of uncertainty is its role within explanatory theories, such as those associated with evolution and quantum physics, for which central models take the form of probablistic statments, i.e., uncertainty is inherent in these models (Gigerenzer, 2000).

Negotiating these many forms of uncertainty in argumentation—and building toward consensus at the same time that uncertainty is present—is a key part of the scientific process (Thagard, 2000). In consequence, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions about the world (Giere, 2010; Godfrey-Smith, 2006; Nersessian, 2002; Reichenbach, 1977).

However, while scientists and philosophers of science may be aware of the uncertainty (and limitations) surrounding scientific knowledge and negotiating standards of evidence interpretation, scientific knowledge is still often portrayed as immutable in educational contexts (Duschl et al., 1990). In consequence, trust in science can erode when scientists individually or collectively change their views, such as during the first phase of the COVID-19 crisis in Germany, when scientists faced serious criticism from politicians and the media for changing their opinion based on new information. A similar tension was present in the United States and around the world regarding potential COVID-19 medication, when initial scientific data and changing interpretations of its utility collided with citizens’ skepticism (and, partisan political views).

Adding to this problem around the limited recognition of the tentative nature of science, when scientists try to communicate their uncertainty, research has shown us that many—including learners and experts alike—struggle to interpret statistics such as confidence intervals or error bars (Gigerenzer et al. 2004; Sedlmeier, 2007; Tversky & Kahneman, 1977). Similarly, difficulties in understanding the language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution of nuclear decay that are modelled in terms of probabilistic statements (e.g., Fiedler et al., 2019).

Based on accumulating evidence from cognitive psychology, neuroscience, and education, we build on and extend beyond the science and engineering practice of argumentation the foundational work of science education (e.g., Nussbaum, 2011; Szu & Osborne, 2012) and statistics education scholars (e.g., Albert, 2002; Bolstad, 2002; Hoegh et al., 2020; Hu, 2020; Martignon and Erickson, 2014, Sedlmeier, 2007; Witmer, 2017) to argue for a relatively novel approach to uncertainty and probability. Specifically, we argue that it is helpful to consider uncertainty in terms of a Bayesian epistemological framework to help people to understand uncertainty and probability better and to engage with them in a way that leverages their existing ways of thinking. Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains (Gopnik, 2012). In this way, we make this conjecture not only on the merits of Bayes Theorem as a mathematical or statistical tool, but based on a body of psychological and educational research that shows that such an approach is an intuitive way for people to think about data and evidence in the world (Gopnik, 2012) and in classroom contexts (Leuders & Loibl, 2020; Nussbaum, 2011; Sedlmeier, 2007; Szu & Osborne, 2012; Witmer, 2017).
The remainder of this paper is structured in the following way. We will first briefly introduce Bayes theorem and wider ideas around Bayesian epistemology. Next, we will elaborate on the importance of probability and uncertainty in science (education) and review how people learn about these concepts and where they face difficulties. Then, we will provide examples of how Bayes could help to address these issues before concluding with implications and directions for future research.

\section{Primer on Bayes Theorem and Bayesian Views of Epistemology}

In essence, Bayes’ theorem provides a framework for updating one’s beliefs about the world from observations in a way that takes prior knowledge into account in a principled manner. Bayes theorem (Equation 1) states that our knowledge about the world after making an observation can be expressed as a probability distribution over the possible outcomes of the phenomenon we observed—the posterior distribution. To calculate the posterior, we express our knowledge about the world before the observation in terms of a probability distribution, the prior, and multiply that with the likelihood which summarizes the information in our observation. These are then divided by the average likelihood to yield a posterior distribution that sums to one.:

Posterior = Prior x LikelihoodAverage Likelihood

\EJ{So I thought about introducing this more elaborately with an example where interest centers on what proportion of students is using
Windows vs. the Mac. Data from the first 10 students are $\{W, W, M, W, M, W, W, M, W, M\}$. I'm using a beta$(2,2)$ prior. The results are shown in the two figures below.}

These are big ideas, and, fortunately, there are a number of intuitive explanations of Bayesian methods, including Kurt (2019), Kruschke and Liddell (2018a) Wagenmakers et al. (2016). A recent example from physics can serve to illustrate this principle in the context of science education. In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light (Brumfiel, 2011) - and thus defying Einstein’s theory of relativity. However, instead of celebrating a historic discovery, the scientists started to look for what had gone wrong in their observations.

How can Bayes’ theorem help to make sense of this? With nearly a century of accumulated evidence that the theory of relativity holds, scientists had a very strong prior belief about the speed of light. To change this belief, extraordinary evidence would have been required. Thus, a single instance of apparently faster-than light elementary particles does not give a likelihood strong enough to change the prior in a meaningful way and in consequence the posterior, the knowledge about the world after making an observation, does not change in a meaningful way.

Thus, rather than changing their belief about the speed of light, scientists looked for things that could have gone wrong with the measurement. This example shows how a qualitative take on Bayes’ theorem can explain and predict many aspects of how animals (Okasha et al., 2012), people (Griffiths et al. 2008, Tenenbaum et al. 2006), and organisations (Kruschke et al., 2012) learn from observations. Recently, advances in computational power and techniques have started the more and more widespread application of this principle in a quantitative way that allows researchers to precisely express their knowledge about the world and then learn from observations.

Expressed in the Bayesian learning cycle (Figure 1), a Bayesian reasoning framework allows scientists to integrate the strengths of both, inductive and deductive research approaches, in a formalized way. Further, when specifying the prior, scientists openly and explicitly need to describe the (un)certainty in the current state of knowledge and the posterior captures how the data has changed the knowledge and the (un)certainty in that knowledge. Thus, Bayesian reasoning is, at least in principle, open and transparent by design.

Figure 1
The Bayesian learning cycle adapted from Wagenmakers and Gronau (2018).
\includegraphics[scale = .75]{learning.png}

Having introduced Bayes theorem and an application of Bayes theorem as a cycle through which individuals can learn, we describe the role of uncertainty in science, with a focus on the role of probability.

\section{Uncertainty in Science and the Role of Probability}

Probability and uncertainty are ubiquitous in science—in scientific methodologies, the scientific concepts, or in how science is communicated (Gougis et al., 2016). The practice of science often begins with measurement, and with measurement comes measurement error, which has a random component. This random component arises from random variations in the measurement process and limits the certainty that one can have in the measurements. Reducing that measurement error has been critical for many discoveries in science, such as in sorting out the periodic table as we know it today (Fontani, 2014). However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., the recent study on the relationship between SARS-CoV-2 viral load and patient age by Jones et al. (2020) was heavily debated in the scientific community (Frick, 2020).

Probability and uncertainty are not only key concepts in science practices such as analyzing and interpreting data, but also part of the scientific concepts themselves. One example is Heisenberg’s uncertainty principle or quantum mechanics more generally  (Feynman, 1951). In quantum mechanics, confidence in measurement is limited not only by measurement error but in principle—even in the perfect experimental setup there is no deterministic outcome. Instead, one has to calculate the probability of each possible outcome using the Born rule. To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunnelling (e.g., Caroll [2020]). Another example of how probability and uncertainty are central to learning scientific concepts is the probabilistic nature of evolution (Fiedler et al., 2017). Evolution is a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically. Further, the continued survival of these variations is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modelled probabilistically. Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers (Millstein, 2014).

Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public (e.g., Gigerenzer 2004, McShane & Gal, 2017, Cumming, 2014). For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign higher probability to the center of the interval (Kruschke & Liddell, 2018a) or p-values are interpreted as effect sizes (Gelman & Carlin, 2017; Nuzzo; 2014) although p-values describe how (in)compatible a set of data is with a set of assumptions. In sum, misinterpretations of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work.

\section{Uncertainty in Science Education}

Given the above-noted roles of probability in scientific concepts and theories, it is unsurprising that educational researchers have documented how students learn about probability. As one example, probability and chance events play an important role in the life sciences (Garfield, 2003; Garvin-Doxas & Klymkowski, 2017), particularly in evolutionary processes and learning about them (Tibell & Harms, 2017). Recently, Fiedler et al. (2019) investigated the relation between statistical reasoning and acceptance and knowledge about evolution in a large sample of nearly 500 university students in the US. They found that students’ statistical reasoning capabilities are a strong predictor of both acceptance of evolution and knowledge about evolution. Thus, this study suggests (but does not directly show) that instructional strategies for supporting students’ statistical reasoning could improve their learning about specific evolutariony concepts.

In the physical sciences, students’ struggles with the probabilistic nature of quantum physics (Bao & Redish, 2002) and nuclear decay (Santostasi et al., 2017) are known, but the underlying reasons and mechanisms remain less researched than in evolution. Research into students’ misconceptions inspired by conceptual change research  (e.g., Marshman & Singh, 2015, Aguilar et al., 2014, Stefani & Tsarparlis, 2009) has started to address this gap. Findings seem to mirror the position of Batanero et al. (1994) in that the concepts of probability are taught too abstractly and in a way that does not build upon and connect with the ideas that students’ already hold, the experiences that students make, and the language that students use. Thus, these studies affirm students' struggles with scientific concepts that are probabilistic but hint at the underlying issue being foundational, conceptual struggles with ideas about probability not limited to science learning. Indeed, the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners. The work of Tversky and Kahneman (1974) points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit. A prominent example is the neglect of base rate information in the (timely) context of medical testing, that is, the probability of having a medical condition on the basis of a positive test result is often hugely overestimated because information about the base rate of people affected from the condition is neglected (Kahneman, 2012).

More recent work by Gigerenzer and colleagues has provided important insights into the underlying mechanisms for the occurence of these biases (e.g., Gigerenzer & Hoffrange, 1995; Jenny et al., 2018). Their research suggests that the biases are at least partly the result of information about probability and uncertainty being provided in a way that is incompatible with the heuristics that people develop from their everyday experiences. Based on this assumption, Jenny et al. (2018) were able to demonstrate that students judge probability significantly better, when language is adopted that better aligns with the heuristics that most learners develop from their everyday experiences. In sum, changing representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that these support instead of inhibit learning seems feasible.

\section{How Learners and People Make Sense of Data in Light of Uncertainty}

Recent work in cognitive and psychological science on how children learn argues that children understand and learn about the world through the intuitive application of Bayesian ideas (Gopnik, 2012; Tenenbaum et al., 2006; Griffiths et al., 2008). This idea has led current research to conceptualize human learning in terms of Bayesian probabilistic models of learning. In other words, people (and children) may view (or come to view) the world probabilistically, and in ways that happen to align with Bayesian interpretations of probability.

This view of how children learn from data has become especially prominent in developmental science and is successfully applied there (Gopnik and Tenenbaum, 2007; Gopnik & Wellman, 2012). One might assume that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education. Developmentalists have pointed to the potential and opportunity here. Gopnik (2012) wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p. 1627).

Despite arguments for integrating developmental and educational research, it has largely not been the case that—with respect to advances in Bayesian methods—“science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” (Gopnik, 2012, p. 1627), as some science education scholars have bemoaned (Lehrer & Schauble, 2006). Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas (Piaget & Inhelder, 1969) that represent the greatest contribution of developmental ideas to education (Lehrer & Schauble, 2007; Metz, 1995). We argue that this is still largely the case in the present, despite advances in both developmental research as well as statistical methodology related to Bayesian ideas—and that it is time to start exploring how bringing out the inherent Bayesian ways of thinking in education may provide support for learners’ efforts to make sense of the world.

Related science education research emphasizes the role of prior knowledge that students bring with them to the analysis of data and how that prior knowledge impacts the inferences they make about data. For example, Masnick et al. (2007) summarized findings across a number of studies showing that when elementary grade students reasoned with data in a context about which they had strong and valid prior understandings that informed their interpretation of data (about force and motion in the context of ramps), and they were able to explain sources of error in the data (Masnick & Klahr, 2003), but when the data was in a context they did not understand well, they struggled to distinguish between random variation due to sources of error and sources of variation due to something that was causing a difference in the data (generated by measuring the motion of pendulums). Later design- and laboratory-based research has shown that students can develop an understanding of scientific ideas through the analysis of data (Lehrer & Schauble, 2004; Masnick et al., 2017).

Laboratory-based research, especially, has examined how learners’ prior understanding may have a bearing on how they make sense of how likely different outcomes are (Klahr & Dunbar, 1998; Masnick et al., 2007; Schwartz et al., 2007). While these studies focused on how students’ prior ideas and data interact, Bayesian frameworks were not used to formalize this process of integrating data in light of prior beliefs about scientific phenomena. In short, this work aligns closely with the Bayesian models of how children learn from data described by Gopnik (2012), but this research does not explicitly interpret their findings in light of these ideas about how children learn from data in a Bayesian way.

Overall, developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in science education to design instruction based around eliciting and understanding students’ ideas (Gotwals & Birmingham, 2012; Haverly et al., 2019; Windschitl et al., 2012). There is an opportunity to use Bayesian ideas as a framework for formalizing these ideas about how prior ideas affect students’ reasoning about data and how teachers can leverage students’ ideas to design instruction, yet Bayes has mostly been absent from research in science education.
After having reviewed how people learn about probability and uncertainty and where they face difficulties, we want to present how Bayes could be part of a solution. First, we will make the case how Bayes could support science literacy at large before delving into more concrete topics.

\section{Bayes and Science Literacy}

\subsection{Overview of Bayes and Science Literacy}

Having described the role of probability and uncertainty in science and science education, we now describe what these Bayesian ideas can mean for science literacy. We see two primary ways in which Bayesian methods can enhance science literacy (National Academies of Sciences, Engineering, and Medicine, 2016), or a range of capabilities that people may possess which have value (individually or collectively) to them. These capabilities range from foundational literacies, including numeracy and textual literacy (what is most thought of concerning literacy in common parlance), content knowledge, and epistemic knowledge and other ways of thinking and habits of mind (National Academies of Sciences, Engineering, and Medicine, 2016). One way that Bayesian ideas can inform discussions of science literacy is to provide individuals with a technique for combining their prior understanding with new evidence. Along this line, Bayes and the Bayesian learning cycle more specifically can provide a conceptual tool for people to combine their understanding with new evidence (and data) in a rigorous and coherent way, as we discuss next.

The history of science is, in many ways, a story of established ideas being built upon or supplanted (Fara, 2010), and the contemporary emphasis on viewing science as a practice was sparked by Kuhn’s (1962) historical-epistemological account on what this suggests about the nature and philosophy of science (Godfrey-Smith, 2009). Yet, our ways of interpreting scientific evidence do rarely recognize this, perhaps for good reason: Most of the time, new data or information do not contradict foundational scientific ideas (Kuhn, 1962). The same phenomenon, though, is present on a more micro-level: While scientists may be familiar with the data they collect not eliding with ideas that are a part of the scientific consensus, in media and the public discourse scientific ideas are often presented as fixed and certain—even in cases for which consensus has not been reached (Collins, 1987).

These historical and philosophical ideas align with a Bayesian perspective on science literacy. Established scientific ideas can be considered to be those that individual people - psychologically - should be more confident about, whereas newer ideas, or data suggesting new ideas, can be weighted less. In this way, Bayes can provide a simple heuristic for people to evaluate new information or evidence; by asking the question of how informative is the evidence in light of the prior. Thus, individuals possess a tool, one that is non-dichotomous, in that the evidence is not true or false, and also probabilistic, in that what individuals can think, going forward, is about the extent to which a particular belief has merit and is informative.

A concrete example can help to illustrate how a Bayesian perspective on a topic relevant to science literacy may offer a different way forward. Consider a salient topic, medical vaccines for diseases, such as Polio and the disease caused by the coronavirus SARS-CoV-2, COVID-19. Most evidence suggests that vaccines have been safe, historically, and are generally safe, at present (Centers for Disease Control and Prevention, 2021). Nevertheless, they are not perfectly safe, as unlikely, adverse outcomes do happen, and it is possible that there are adverse side effects from some vaccines that research has not yet revealed. In light of this evidence, consider the development of a new vaccine, such as those available at the time of this writing for COVID-19. Some use long-established methods of vaccine development (namely, deactivating SARS-CoV-2 particles but retaining cell signalling proteins to which humans develop an immune response—and, after, a degree of immunity) and injecting these deactivated virus particles into people), whereas others use a new mechanism (namely, injecting messenger RNA into people; this messenger RNA is then transcribed by cells in people’s bodies to produce a particular protein common to SARS-CoV-2 particles, which human’s bodies than learn to recognize to develop immunity.

How should we consider evidence about the vaccines? Prior to any data from medical clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and on the basis of similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are. After initial trials were completed, the degree of safety that individuals associate with these vaccines—in light of the prior and the limited evidence—could increase, but still not be near the level of safety that many would consider acceptable. Following the completion of clinical trials that show minimal side effects (holding aside their efficacy), one’s estimates of the safety of these vaccines could increase by a substantial amount, but concerns could still remain, particularly for messenger RNA vaccines, which are newer, and may, plausibly, have unintended side effects.

\subsection{Strategies for Supporting Science Literacy}

What is the result of this entire process of weighting new evidence in light of a generic prior that we discussed above? We argue that there are three ways that Bayesian methods could support science literacy. 
First, the result is a non-binary inference, one that doesn’t speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty. Though a small difference, discussing the safety of vaccines on these terms can provide a more solid foundation for principled disagreements and arguments.

Second, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in-line with how most of us likely already interpret data. Specifically Bayes provides people with a way to talk about weighing between ideas and their prior beliefs and new information and data. In short, Bayes gives people a chance to deal with uncertainty in a principled but accessible way.

Last, his process could be formalized mathematically.

I think there are several very general ideas that could be illustrated with Bayes’ rule. These include:

\begin{enumerate}
    \item When we gain new information, our updated knowledge of the world is a compromise between our old knowledge and the new information.
    \item Today’s posterior is tomorrow’s prior 
    \item Hypotheses/claims that predict new data well receive a boost in credibility, whereas those that predict the data poorly suffer a decline.
    \item All knowledge is relative
    \item Avoid absolute beliefs (Cromwell’s rule; these are not amenable to learning)
    \item Surprise lost is credibility gained
    \item Polya’s fundamental inductive pattern: “This inductive pattern says nothing surprising. On the contrary, it expresses a belief which no reasonable person seems to doubt: The verification of a consequence renders a conjecture more credible. With a little attention, we can observe countless reasonings in everyday life, in the law courts, in science, etc., which appear to confirm to our pattern.” (Polya, 1954, pp. 4-5)
    \item Extraordinary claims require extraordinary evidence
\end{enumerate}

\section{Bayes and Science Education}

\subsection{Overview of Bayes and Science Education}

Research on the role of Bayes in science education has mostly been carried out in the context of undergraduate statistics education, in which there have been a number of calls to action (Gould et al., 2018; Hoegh et al., 2020), examples and design options (Albert, 2002; Bolstad, 2002; Hu, 2020; Gelman, 2008; Witmer, 2017) and debate (Johnson et al., 2020) over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs. The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power (Gould et al., 2018) as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers (Albert & Hu, 2020). As evidenced by the recent special issue of the Journal of Statistics Education of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.

There is little research outside of statistics education in the broader science education community. Two publications merit mention though. The work of Szu and Osborne (2012) and Nussbaum (2011) both applied Bayesian perspectives to the science practice of argumentation. Szu and Osborne presented the case for how and why Bayes Theorem can apply to research and practice on scientific reasoning, broadly, and argumentation, particularly. This paper represents the most comprehensive account of the relevance of Bayesian methods for science education. Szu and Osborne argue that Bayes is useful as both a formal mathematical tool, and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes Theorem that follow more or less deterministically from “external, objectively probabilistic systems” (p. 61)—is “the key leap that characterizes the debate about the value of Bayesian inference as a model of scientific reasoning” (p. 61). In this way, Szu and Osborne argue that the greatest use of Bayes theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community (e.g., Batanero et al., 2016; Makar & Rubin, 2018).  They offer some research-related backing for the use of Bayes Theorem (e.g., noting how conceptual change research, particularly, and constructivist research, broadly, can both be explained coherently within a Bayesian approach) as well as some practical, instructional recommendations, some of which we detail later in this section.

Nussbaum’s (2011) work on Bayesian approaches to argumentation complement the work of Szu and Osborne in that Nussbaum describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation. Concerning the latter, Nussbaum describes how the social issue of raising taxes to provide resources to homeless individuals was a rich context for students to engage in forms of Bayesian reasoning. In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what Szu and Osborne (2012) characterize as the more externally objective use of Bayes Theorem—and how the estimates that result from applying Bayes theorem led students to re-evaluate their initial arguments.

In summary, the work of Szu and Osborne (2012) and Nussbaum begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process. Curiously, no work has applied Bayesian methods to another science practice, that of analyzing and interpreting data; one for which Bayesian methods have clear conceptual applicability and utility, but one for which the absence of suitable tools may provide to be a barrier. In this way, Bayesian methods could serve as a cross-cutting concept (National Research Council, 2012) that unites the practice of argumentation with analyzing and interpreting data, and using mathematics and computational thinking.

\subsection{Strategies for Supporting Science Education}

Bayesian methods clearly have some utility in science education, even if that utility is under-realized at present. Next, we present some of the specific ways Bayesian methods could enhance science education—and could do so in a way that makes Bayes more tractable.

First, we should consider “unplugged” versions of Bayesian analyses, akin to how computer science education scholars have emphasized that computational thinking does not necessarily require the use of a computer (Yadav et al., 2018). These could interface with the science and engineering practice of modeling; could students use diagrammatic models in conjunction with Bayesian “heuristics” to revise their models in a similar manner to how this practice can unfold in classrooms (see Schwarz et al., 2009).

Another advance should be to build out curricular strategies, elaborating on the ideas of Szu and Osborne (2011) and Nussbaum (2011). At a high-level, Szu and Osborne (2011) write that in order to reason about data in light of their initial beliefs (and to construct a kind of informal likelihood statistic), students “need to see judgments about data and evidence being an assessment not only of the probability of the hypothesis being correct but also of it being wrong” (p. 67). Filling in more details, speaking from both the perspective of a researcher (of argumentation) and an educator (or, a designer of classroom activities), Nussbaum summarizes that using Bayes “forces one to (a) think about an issue from different perspectives, (b) reflect on one’s range of certainty and uncertainty, and (c) conduct a simple com- puter simulation (e.g., calculating the product) that integrates various facets of the issue together and affords further reflection” (p. 99).

These are helpful suggestions, and they complement finer-grained instructional techniques, especially those around how students can reason about uncertainty probabilistically. Particularly, Binder et al. (2018), for instance, suggest replacing probabilities with natural frequencies (in alignment with a great deal of work on how humans understand probability; Gigerenzer & Hoffrange, 1995) as well as using and interpreting tree diagrams to depict natural frequencies. Martignon and Erickson (2014)—building on much earlier experimental work by Wason and Shapiro (1966)—provide a simple example of how even fourth graders, with the right tools to represent probabilities, can support students to transition from logical to probabilistic reasoning. These could also support the aim of developing unplugged versions of Bayesian analyses.

Third, we should, of course, consider teachers and how teachers can (or already) support students to reason about uncertainty, and, possibly, to use Bayesian ideas. While some scholars have contributed foundational work demonstrating how this uncertainty can be productive as a part of teaching practice (Manz & Suarez, 2018), a broad and formal perspective on how uncertainty can be approached in science education is needed. There is the need to explore the potential of a principled approach to uncertainty, one that may be intuitive to learners and useful to teachers, but is as yet under-represented in the science education literature. If learners and people better understand uncertainty, they can better understand the scientific process which helps to build trust in science. Further, a better understanding of uncertainty can help to learn about scientific--and other--concepts where probability plays a key role; it may be that learning about how Bayes rule can be used in legal or medical contexts could impress upon students the cross-cutting nature of probability and uncertainty. This is especially helpful in light of the teacher's practice of noticing and responding to students' ideas and the process through which students change their understanding.

Finally, recently developed tools, particularly the JASP statistical software, may be helpful. JASP allows for both conventional frequentist as well as Bayesian versions of commonly used statistical tests, such as t-tests and regression analyses. There are a number of accessible guides available for specifying Bayesian versions of analyses (van Doorn et al., 2020; Wagenmakers et al., 2018), and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms. Another possibly useful tool could be Bayesian networks (see Bayes Box; Bayes Box, 2021), which could be used to model complex system in a way that is similar to other tools used in science education for modeling complex systems, such as SageModeler (2020).

\section{Summary and Conclusion}

While some educational researchers may state that such theoretical, abstract ideas about the philosophical and statistical underpinnings of working with data are irrelevant to how students work with data, we make the case, here, that these ideas can inform instructional practice, assessments, and the design of activities that involve students in working with data, especially given some of the clear problems that have been articulated around the present frequentist Null Hypothesis Statistical Testing (NHST)-based approaches (Cumming, 2014; Gigerenzer, 2004).

Science education researchers may note that these ideas are conceptually-focused, rather than aligning with the present emphasis in science education upon engaging students in science and engineering practice in order to make sense of core ideas and to engage with cross-cutting concepts (National Research Council, 2012). We argue that Bayes is unique in that it, perhaps akin to a cross-cutting concept in the sense of a style of reasoning (Szu & Osborne, 2012; Osborne et al. 2018) that blends conceptual ideas with practices. Bayes is both a philosophical and psychological perspective on how to learn from data as well as a statistical framework. Indeed, a key potential benefit of Bayes may be that it provides a framework for helping students to transition from their conceptual understanding of scientific ideas to quantitatively analyzing data.

In this paper, we have introduced Bayes as a powerful way of conceptualizing probability and uncertainty. These are two concepts that are ubiquitous in the learning, doining, and communicating of science and therefore critical aspects of science literacy (National Academies of Sciences, Engineering, and Medicine, 2016). However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike (Gigerenzer et al. 2004; Sedlmeier, 2007; Tversky & Kahneman, 1977). Thus, it is not surprising when the public misinterprets statements about probability and uncertainty but the consequences may be dire when trust in science is eroded, and, sadly, examples of this are plentiful during current the COVID-19, e.g., the efficiency of masks in preventing the spread of COVID and the efficiency of the AstraZeneca vaccine were strongly questioned, delaying efforts to comb at the pandemic. The public, however, can hardly be blamed for this in light of the prevalent findings that the frequentist conceptualizations of probability and uncertainty in school are challenging and uncommonly align with or build on peoples’ intuitions. As research from cognitive and psychological science as well as emerging work from statistics and statistics data science education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions (Kruschke & Liddell, 2018a; Szu & Osborne, 2012) about these concepts and thus has potential to support learning about probability and uncertainty in science.

We have outlined numerous vantage points for research agendas to explore the potential of Bayes in science education so that future research can identify how students can become better reasoners in light of uncertainty. We have also discussed how we as scientists and science educators can contribute to the public trust in science by becoming better in how we communicate about probability and uncertainty. In all, Bayes is a big idea—one with epistemological and philosophical, statistical, and pedagogical implications. We argue that the time for advancing Bayesian methods has come, particularly in light of distrust in science and science education and the way Bayesian methods can address this state of affairs by offering a different foundation and language around a core but sometimes unsettling tenet and persistent feature, uncertainty.

\printbibliography

\appendix

\section{code}
add JASP links/code/etc.?

\end{document}

%% 
%% Copyright (C) 2019 by Daniel A. Weiss <daniel.weiss.led at gmail.com>
%% 
%% This work may be distributed and/or modified under the
%% conditions of the LaTeX Project Public License (LPPL), either
%% version 1.3c of this license or (at your option) any later
%% version.  The latest version of this license is in the file:
%% 
%% http://www.latex-project.org/lppl.txt
%% 
%% Users may freely modify these files without permission, as long as the
%% copyright line and this statement are maintained intact.
%% 
%% This work is not endorsed by, affiliated with, or probably even known
%% by, the American Psychological Association.
%% 
%% This work is "maintained" (as per LPPL maintenance status) by
%% Daniel A. Weiss.
%% 
%% This work consists of the file  apa7.dtx
%% and the derived files           apa7.ins,
%%                                 apa7.cls,
%%                                 apa7.pdf,
%%                                 README,
%%                                 APA7american.txt,
%%                                 APA7british.txt,
%%                                 APA7dutch.txt,
%%                                 APA7english.txt,
%%                                 APA7german.txt,
%%                                 APA7ngerman.txt,
%%                                 APA7greek.txt,
%%                                 APA7czech.txt,
%%                                 APA7turkish.txt,
%%                                 APA7endfloat.cfg,
%%                                 Figure1.pdf,
%%                                 shortsample.tex,
%%                                 longsample.tex, and
%%                                 bibliography.bib.
%% 
%%
%% End of file `./samples/longsample.tex'.

### 2021-04-29 manuscript.tex
changes from e.j.'s adds

@@
 \usepackage{epigraph}
-% For sorting references from same year in order instead of on title
+\usepackage{amsfonts, amsmath,amssymb}
+
 \newcommand{\SR}[1]{} %for EJ's bibtex file
+\newcommand{\given}{\, | \,}
 \newcommand{\EJ}[1]{\todo[inline, color=green]{  #1 }}
@@
 \addbibresource{bibliography.bib}
+\addbibresource{referenties.bib}
 
-\title{Trust Me, I'm a Bayesian: Building Back Trust in Science Using Bayes}
-\shorttitle{Bayes in Science Education}
+% \setlength{\parindent}{8ex}
 
-\author{Joshua M. Rosenberg}
-\affiliation{University of Tennessee, Knoxville}
+% \usepackage{authblk}
 
-\author{Marcus Kubsch}
-\affiliation{IPN - Leibniz Institute for Science and Mathematics Education}
+\title{Building Trust in Science Through a Bayesian Perspective on Probability and Uncertainty}
+\shorttitle{Building Trust Through a Bayesian Perspective}
 
-\author{Eric-Jan Wagenmakers}
-\affiliation{University of Amsterdam}
+\threeauthors{Joshua M. Rosenberg}{Marcus Kubsch}{Eric-Jan Wagenmakers}
+\threeaffiliations{University of Tennessee, Knoxville}{IPN - Leibniz Institute for Science and Mathematics Education}{University of Amsterdam}
+
+% \authorsnames[1,2,3]{Joshua M. Rosenberg, Marcus Kubsch, Eric-Jan Wagenmakers}
+% \authorsaffiliations{{University of Tennessee, Knoxville},{IPN - Leibniz Institute for Science and Mathematics Education},{University of Amsterdam}}
 
@@
 \abstract{
-
 Uncertainty is ubiquitous in science and this uncertainty takes many forms, but the rhetoric around scientific knowledge and negotiating standards of evidence is often portrayed as immutable.
@@
 Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains.
-In this essay, we a) briefly introduce Bayes theorem and wider ideas around Bayesian epistemology, b) elaborate on the importance of probability and uncertainty in science and science education and review how people learn about these concepts and where they face difficulties, and c) provide conceptual and practical examples within the JASP statistical software of how Bayes could help to address these issues.
+In this essay, we (a) briefly introduce Bayes' theorem and wider ideas around Bayesian epistemology, (b) elaborate on the importance of probability and uncertainty in science and science education and review how people learn about these concepts and where they face difficulties, and (c) provide conceptual and practical examples within the JASP statistical software of how Bayes could help to address these issues.
 We conclude with implications and directions for future research focused around bolstering trust in science outside of classroom settings (for which Bayesian approaches could serve as a crosscutting part of science literacy) and in pre-collegiate science classrooms, where Bayes can provide a coherent approach to engaging in numerous science practices such as arguing from evidence and analyzing and interpreting data in a principled and flexible way.
@@
    \addORCIDlink{Joshua M. Rosenberg}{0000-0003-2170-0447}
-   \addORCIDlink{Marcus Kubsch}{0000-0000-0000-0000}
+   \addORCIDlink{Marcus Kubsch}{0000-0001-5497-8336}
    \addORCIDlink{Eric-Jan Wagenmakers}{0000-0003-1596-1034} 
+   \\
+   Joshua M. Rosenberg and Marcus Kubsch contributed equally to this manuscript.
 }
@@
 
-\EJ{This is a comment.}
+\EJ{I have doubts about the ``Trust me, I'm a Bayesian'' part of the title, also because ``Trust'' now occurs twice.
+Maybe simply ``Building Trust in Science Through a Bayesian Perspective on Probability and Uncertainty''?}
+\MK{Ok with the new title - I liked the sound of the trust me thing but utlimatley it may not be really on point!}
+\EJ{I also uploaded my bibtex file, ``referenties.bib''.
+I could not get it to work, but you could copy the relevant references into your own bibtex file.
+Actually, I also managed to make the initial bibtex setup not work, and I have absolutely no clue as to how this happened.}
 
@@
 For example, measurement uncertainty is present when there are differences (small or large) in the repeated measures of the same object or phenomenon \parencite{pls03} or when the object that is measured, like a plant, is inherently variable in its characteristics \parencite{pls03}.
-Another form of uncertainty is its role within explanatory theories, such as those associated with evolution and quantum physics, for which central models take the form of probablistic statments, i.e., uncertainty is inherent in these models (Gigerenzer, 2000).
+Another form of uncertainty is its role within explanatory theories, such as those associated with evolution and quantum physics, for which central models take the form of probablistic statements, i.e., uncertainty is a core constituent of these models \parencite{g00}.
 
-Negotiating these many forms of uncertainty in argumentation—and building toward consensus at the same time that uncertainty is present—is a key part of the scientific process (Thagard, 2000).
-In consequence, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions about the world (Giere, 2010; Godfrey-Smith, 2006; Nersessian, 2002; Reichenbach, 1977).
+\EJ{Yeah so as a hard-nosed determinist I still have an issue with ``inherent'' uncertainty.
+If we don't sort this out we could resolve it through a footnote (e.g., ``EJ Wagenmakers would like to note that he is a determinist who believes that the concepts of probability and uncertainty belongs wholly to the mind \parencite{Jevons18741913}, with the possible exception of the strange world of quantum mechanics.''}
+\MK{I got rid of the inherent part.}
 
-However, while scientists and philosophers of science may be aware of the uncertainty (and limitations) surrounding scientific knowledge and negotiating standards of evidence interpretation, scientific knowledge is still often portrayed as immutable in educational contexts (Duschl et al., 1990).
+\EJ{It would be good to go over the reference list in detail; in particular, I noticed several ``bayes'' that should be ``Bayes'' (the B needs to be enclosed in brackets so it displays in upper case)}
+Negotiating these many forms of uncertainty in argumentation—and building toward consensus at the same time that uncertainty is present—is a key part of the scientific process \parencite{t00}.
+In consequence, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions about the world  \parencite{g10, g06, n02, r77}.
+
+However, while scientists and philosophers of science may be aware of the uncertainty (and limitations) surrounding scientific knowledge and negotiating standards of evidence interpretation, scientific knowledge is still often portrayed as immutable in educational contexts \parencite{d90}.
 In consequence, trust in science can erode when scientists individually or collectively change their views, such as during the first phase of the COVID-19 crisis in Germany, when scientists faced serious criticism from politicians and the media for changing their opinion based on new information.
@@
 
-Adding to this problem around the limited recognition of the tentative nature of science, when scientists try to communicate their uncertainty, research has shown us that many—including learners and experts alike—struggle to interpret statistics such as confidence intervals or error bars (Gigerenzer et al.
-2004; Sedlmeier, 2007; Tversky & Kahneman, 1977).
-Similarly, difficulties in understanding the language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution of nuclear decay that are modelled in terms of probabilistic statements (e.g., Fiedler et al., 2019).
+Adding to this problem around the limited recognition of the tentative nature of science, when scientists try to communicate their uncertainty, research has shown us that many—including learners and experts alike—struggle to interpret statistics such as confidence intervals or error bars \parencite{gkv04, s07, tk74}.
+Similarly, difficulties in understanding the language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution of nuclear decay that are modelled in terms of probabilistic statements \parencite{fth17}.
 
-Based on accumulating evidence from cognitive psychology, neuroscience, and education, we build on and extend beyond the science and engineering practice of argumentation the foundational work of science education (e.g., Nussbaum, 2011; Szu & Osborne, 2012) and statistics education scholars (e.g., Albert, 2002; Bolstad, 2002; Hoegh et al., 2020; Hu, 2020; Martignon and Erickson, 2014, Sedlmeier, 2007; Witmer, 2017) to argue for a relatively novel approach to uncertainty and probability.
+Based on accumulating evidence from cognitive psychology, neuroscience, and education, we build on and extend beyond the science and engineering practice of argumentation the foundational work of science educational \parencite{n11, s07} and statistics education scholars \parencite{a02, b02, h_a20, h_j20,me14, s07} to argue for a \EJ{I'd say ``underused'', because Bayes itself is pretty old and now pretty hip} approach to uncertainty and probability that is currently underrepresented in science education (research).
 Specifically, we argue that it is helpful to consider uncertainty in terms of a Bayesian epistemological framework to help people to understand uncertainty and probability better and to engage with them in a way that leverages their existing ways of thinking.
-Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains (Gopnik, 2012).
-In this way, we make this conjecture not only on the merits of Bayes Theorem as a mathematical or statistical tool, but based on a body of psychological and educational research that shows that such an approach is an intuitive way for people to think about data and evidence in the world (Gopnik, 2012) and in classroom contexts (Leuders & Loibl, 2020; Nussbaum, 2011; Sedlmeier, 2007; Szu & Osborne, 2012; Witmer, 2017).
+Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains \parencite{g12}.
+In this way, we make this conjecture not only on the merits of Bayes' Theorem as a mathematical or statistical tool, but based on a body of psychological and educational research that shows that such an approach is an intuitive way for people to think about data and evidence in the world \parencite{g12} and in classroom contexts \parencite{ll20, n11, s07, so12, w17}.
+
 The remainder of this paper is structured in the following way.
@@
 
-In essence, Bayes’ theorem provides a framework for updating one’s beliefs about the world from observations in a way that takes prior knowledge into account in a principled manner.
-Bayes theorem (Equation 1) states that our knowledge about the world after making an observation can be expressed as a probability distribution over the possible outcomes of the phenomenon we observed—the posterior distribution.
-To calculate the posterior, we express our knowledge about the world before the observation in terms of a probability distribution, the prior, and multiply that with the likelihood which summarizes the information in our observation.
-These are then divided by the average likelihood to yield a posterior distribution that sums to one.:
+In essence, Bayes' theorem provides a framework for updating one's beliefs about the world from observations in a way that takes prior knowledge into account in a principled manner.
+Bayes' theorem (Equation 1) states that at any point in time, our knowledge about the world after making a new observation (i.e., our posterior knowledge) can be obtained by multiplying our prior knowledge with a predictive updating factor (e.g., \cite{RouderMorey2019,WagenmakersEtAl2016CD} ):
+\begin{equation*}
+        \text{Posterior uncertainty} = \text{Prior uncertainty} \times \text{Predictive updating factor}.
+\end{equation*}
+Mathematically, this can be expressed as  
+\begin{equation}
+\label{eq:BayesRule}
+        p(\theta \mid \text{data}) = p(\theta) \times \frac{p(\text{data} \mid \theta)}{p(\text{data})},
+\end{equation}
+where $\theta$ is a hypothesis, proposition, claim, parameter value, or in general any particular account of the world about which we are uncertain.
+The predictive updating factor consists of two components.
+First, $p(\text{data} \given \theta)$ is the likelihood, that is, the extent to which the observed data were expected under $\theta$.
+Second, $p(\text{data})$ is the marginal likelihood, that is, the extent to which the observed data were expected without conditioning on $\theta$.
+If the observed data are more expected (i.e., better predicted) when conditioning on $\theta$, the predictive updating factor is higher than 1, and $\theta$ becomes more credible.
 
-Posterior = Prior x LikelihoodAverage Likelihood
+To make this more tangible, consider the following example.
+Suppose you are interested in the proportion of students that use Windows vs.
+MacOS as their operating system.
+The proportion of Windows users $\theta$ ranges between 0 and 1.
+Before you start our Bayesian analysis, you need to quantify your background knowledge -- the relative plausibility of the different values for $\theta$.
+Given your previous experiences in class rooms your best guess is that about half of the students use MacOS.
+You would not be surprised if the  proportion of Windows users was between 30\% and 70\% but you would be rather surprised to find either no Windows or exclusive Windows usage.
+This background knowledge motivates the use of the dome-shaped prior distribution shown as the dashed line in Figure~\ref{fig:WindowsMacPriorPosterior}.\footnote{This is the beta$(2,2)$ distribution.} Now suppose you start to collect data by asking students about their operating systems.
+The first ten observations of Windows (W) and MacOS (M) usage that you make are $\{W, W, M, W, M, W, W, M, W, M\}$: six students use Windows, four use MacOS.
+With these data in hand, we then combine our prior knowledge (i.e., the dome-shaped prior distribution) with the predictive updating factor (assuming a binomial likelihood) and obtain the posterior distribution shown as the solid line in Figure~\ref{fig:WindowsMacPriorPosterior}.
+The posterior distribution is more peaked than the prior distribution, indicating that the data have sharpened our knowledge about $\theta$.
+In addition, the posterior distribution is higher than the prior distribution for values of $\theta$ between approximately $.4$ and $.75$: these values of $\theta$ predicted the data relatively well and have therefore gained credibility.
+In contrast, values of $\theta$ lower than $.4$ and higher than $.75$ predicted the data relatively poorly, which is why they have lost credibility compared to the prior.\footnote{Note that the area under a probability distribution equals 1, so that posterior gains for specific regions of $\theta$ cancel out exactly against posterior losses for the other regions.}
 
-\EJ{So I thought about introducing this more elaborately with an example where interest centers on what proportion of students is using
-Windows vs. the Mac. Data from the first 10 students are $\{W, W, M, W, M, W, W, M, W, M\}$. I'm using a beta$(2,2)$ prior. The results are shown in the two figures below.}
+\begin{figure*}[h]
+\begin{center}
+\includegraphics[width = .785\paperwidth]{WindowsMacPriorPosterior.pdf}
+\caption{Simple example of Bayesian learning.
+A dome-shaped prior distribution captures the background knowledge concerning the proportion of students using Windows vs.
+MacOS.
+Observing fictitious data from 10 students (six using Windows, four using MacOS) drives a knowledge update that results in a bell-shaped posterior distribution.
+Figure based on the \emph{Learn Bayes} module in JASP.}
+\label{fig:WindowsMacPriorPosterior}
+\end{center}
+\end{figure*}
 
-These are big ideas, and, fortunately, there are a number of intuitive explanations of Bayesian methods, including Kurt (2019), Kruschke and Liddell (2018a) Wagenmakers et al.
-(2016).
-A recent example from physics can serve to illustrate this principle in the context of science education.
-In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light (Brumfiel, 2011) - and thus defying Einstein’s theory of relativity.
+Although we have demonstrated this updating process in a single step, it could also be executed sequentially, one student at a time.
+For instance, the first observation you make is a $W$ and in light of your prior this makes Windows usage a little more likely than MacOS usage.
+This slight change in knowledge is reflected in the difference between the distributions on the top two lines in Figure~\ref{fig:WindowsMacSequential}; line ``0'' represents the dome-shaped prior distribution and line ``1'' represents the posterior distribution after the first observations.
+Note that observing a $W$ has nudged the distribution a little to the right.
+Crucially, the `nudged' posterior distribution after the first observation now takes the role of the prior distribution ready to be updated by our second observation.
+The second observation is again a $W$, and line ``2'' shows that the resulting posterior distribution is nudged to the right once more.
+At every stage in the sequential updating process, the posterior distribution based on the observations seen so far becomes the prior distribution for incorporating the information from the next observation.
+Specifically, the changing distributions in Figure~\ref{fig:WindowsMacSequential} show our knowledge about the world is subject to constant change, by making observations and integrating incoming information with current knowledge using Bayes' theorem.
+It should be emphasized that although the updating process is conceptually different, the final outcome is identical: the shape of the posterior distribution is unaffected by whether data arrive simultaneously or sequentially.
+
+\begin{figure*}[h]
+\begin{center}
+\includegraphics[width = .65\paperwidth]{WindowsMacSequential.pdf}
+\caption{Sequential Bayesian learning.
+A dome-shaped prior distribution (line ``0'') captures the background knowledge concerning the proportion of students using Windows vs.
+MacOS.
+Each new observation results in an update to a posterior distribution, which then becomes the prior distribution for the analysis of the next observation.
+Figure based on the \emph{Learn Bayes} module in JASP.}
+\label{fig:WindowsMacSequential}
+\end{center}
+\end{figure*}
+
+But how do the principles from this example translate to science (education)?
+A recent case from physics can serve to illustrate this.
+In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light \parencite{b11} - and thus defying Einstein’s theory of relativity.
 However, instead of celebrating a historic discovery, the scientists started to look for what had gone wrong in their observations.
@@
 Thus, rather than changing their belief about the speed of light, scientists looked for things that could have gone wrong with the measurement.
-This example shows how a qualitative take on Bayes’ theorem can explain and predict many aspects of how animals (Okasha et al., 2012), people (Griffiths et al.
-2008, Tenenbaum et al.
-2006), and organisations (Kruschke et al., 2012) learn from observations.
+This example shows how a qualitative take on Bayes’ theorem can explain and predict many aspects of how animals \parencite{o13}, people \parencite{tgk06}, and organisations \parencite{kaj12} learn from observations.
 Recently, advances in computational power and techniques have started the more and more widespread application of this principle in a quantitative way that allows researchers to precisely express their knowledge about the world and then learn from observations.
 
-Expressed in the Bayesian learning cycle (Figure 1), a Bayesian reasoning framework allows scientists to integrate the strengths of both, inductive and deductive research approaches, in a formalized way.
+Expressed in the Bayesian learning cycle (Figure \ref{fig:The Bayesian Learning Cycle}), a Bayesian reasoning framework allows scientists to integrate the strengths of both, inductive and deductive research approaches, in a formalized way.
 Further, when specifying the prior, scientists openly and explicitly need to describe the (un)certainty in the current state of knowledge and the posterior captures how the data has changed the knowledge and the (un)certainty in that knowledge.
@@
 
-Figure 1
-The Bayesian learning cycle adapted from Wagenmakers and Gronau (2018).
-\includegraphics[scale = .75]{learning.png}
+\begin{figure*}[h]
+\begin{center}
+\includegraphics[width = .65\paperwidth]{learning.png}
+\caption{The Bayesian Learning cycle}
+\label{fig:The Bayesian Learning Cycle}
+\end{center}
+\end{figure*}
 
@@
 
-Probability and uncertainty are ubiquitous in science—in scientific methodologies, the scientific concepts, or in how science is communicated (Gougis et al., 2016).
+Probability and uncertainty are ubiquitous in science—in scientific methodologies, the scientific concepts, or in how science is communicated \textcite{gsoobmc17}.
 The practice of science often begins with measurement, and with measurement comes measurement error, which has a random component.
 This random component arises from random variations in the measurement process and limits the certainty that one can have in the measurements.
-Reducing that measurement error has been critical for many discoveries in science, such as in sorting out the periodic table as we know it today (Fontani, 2014).
-However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., the recent study on the relationship between SARS-CoV-2 viral load and patient age by Jones et al.
-(2020) was heavily debated in the scientific community (Frick, 2020).
+Reducing that measurement error has been critical for many discoveries in science, such as in sorting out the periodic table as we know it today \textcite{fco15}.
+However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., the recent study on the relationship between SARS-CoV-2 viral load and patient age by \textcite{jmvbzhd20} was heavily debated in the scientific community \parencite{frick_peer-review_2020}.
 
 Probability and uncertainty are not only key concepts in science practices such as analyzing and interpreting data, but also part of the scientific concepts themselves.
-One example is Heisenberg’s uncertainty principle or quantum mechanics more generally  (Feynman, 1951).
+One example is Heisenberg’s uncertainty principle or quantum mechanics more generally \textcite{feynman1951operator}.
 In quantum mechanics, confidence in measurement is limited not only by measurement error but in principle—even in the perfect experimental setup there is no deterministic outcome.
 Instead, one has to calculate the probability of each possible outcome using the Born rule.
-To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunnelling (e.g., Caroll [2020]).
-Another example of how probability and uncertainty are central to learning scientific concepts is the probabilistic nature of evolution (Fiedler et al., 2017).
-Evolution is a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically.
-Further, the continued survival of these variations is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modelled probabilistically.
-Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers (Millstein, 2014).
-
-Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public (e.g., Gigerenzer 2004, McShane & Gal, 2017, Cumming, 2014).
-For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign higher probability to the center of the interval (Kruschke & Liddell, 2018a) or p-values are interpreted as effect sizes (Gelman & Carlin, 2017; Nuzzo; 2014) although p-values describe how (in)compatible a set of data is with a set of assumptions.
+To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunnelling  \textcite{c19}.
+Another example of how probability and uncertainty are central to learning scientific concepts is  evolution  \textcite{fsnh19}.
+Evolution is modelled as a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically.
+Further, the continued survival of these variations is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modelled probabilistically.\EJ{I agree that we can only describe the process probabilistically.
+But that does not mean it \emph{is} probabilistic.} \MK{tried to address this comment as well.} Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers  \textcite{m16}.
+Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public \parencite[]{gkv04, c14, mg17}.
+For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign higher probability to the center of the interval \parencite{kl18} or \emph{p}-values are interpreted as effect sizes \parencite{gc17, n14} (although p-values describe how (in)compatible a set of data is with a set of assumptions.
 In sum, misinterpretations of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work.
@@
 Given the above-noted roles of probability in scientific concepts and theories, it is unsurprising that educational researchers have documented how students learn about probability.
-As one example, probability and chance events play an important role in the life sciences (Garfield, 2003; Garvin-Doxas & Klymkowski, 2017), particularly in evolutionary processes and learning about them (Tibell & Harms, 2017).
-Recently, Fiedler et al.
-(2019) investigated the relation between statistical reasoning and acceptance and knowledge about evolution in a large sample of nearly 500 university students in the US.
+As one example, probability and chance events play an important role in the life sciences \parencite{g03, gk08}, particularly in evolutionary processes and learning about them \parencite{th17}.
+Recently, \textcite{fsnh19} investigated the relation between statistical reasoning and acceptance and knowledge about evolution in a large sample of nearly 500 university students in the US.
 They found that students’ statistical reasoning capabilities are a strong predictor of both acceptance of evolution and knowledge about evolution.
-Thus, this study suggests (but does not directly show) that instructional strategies for supporting students’ statistical reasoning could improve their learning about specific evolutariony concepts.
+Thus, this study suggests (but does not directly show) that instructional strategies for supporting students’ statistical reasoning could improve their learning about specific evolutionary concepts.
 
-In the physical sciences, students’ struggles with the probabilistic nature of quantum physics (Bao & Redish, 2002) and nuclear decay (Santostasi et al., 2017) are known, but the underlying reasons and mechanisms remain less researched than in evolution.
-Research into students’ misconceptions inspired by conceptual change research  (e.g., Marshman & Singh, 2015, Aguilar et al., 2014, Stefani & Tsarparlis, 2009) has started to address this gap.
-Findings seem to mirror the position of Batanero et al.
-(1994) in that the concepts of probability are taught too abstractly and in a way that does not build upon and connect with the ideas that students’ already hold, the experiences that students make, and the language that students use.
+In the physical sciences, students' struggles with the probabilistic nature of quantum physics \parencite{br02} and nuclear decay \parencite{smmv17} are known, but the underlying reasons and mechanisms remain less researched than in evolution.
+Research into students’ misconceptions inspired by conceptual change research has started to address this gap \parencite{ms15, alra14, st09}.
+Findings seem to mirror the position of \textcite{bgv94} in that the concepts of probability are taught too abstractly and in a way that does not build upon and connect with the ideas that students’ already hold, the experiences that students make, and the language that students use.
 Thus, these studies affirm students' struggles with scientific concepts that are probabilistic but hint at the underlying issue being foundational, conceptual struggles with ideas about probability not limited to science learning.
 Indeed, the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners.
-The work of Tversky and Kahneman (1974) points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit.
-A prominent example is the neglect of base rate information in the (timely) context of medical testing, that is, the probability of having a medical condition on the basis of a positive test result is often hugely overestimated because information about the base rate of people affected from the condition is neglected (Kahneman, 2012).
+The work of \textcite{tk74} points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit.
+A prominent example is the neglect of base rate information in the (timely) context of medical testing, that is, the probability of having a medical condition on the basis of a positive test result is often hugely overestimated because information about the base rate of people affected from the condition is neglected \parencite{kahneman_thinking_2012}.
 
-More recent work by Gigerenzer and colleagues has provided important insights into the underlying mechanisms for the occurence of these biases (e.g., Gigerenzer & Hoffrange, 1995; Jenny et al., 2018).
+More recent work by Gigerenzer and colleagues has provided important insights into the underlying mechanisms for the occurrence of these biases \parencite{gh95, jkg18}.
 Their research suggests that the biases are at least partly the result of information about probability and uncertainty being provided in a way that is incompatible with the heuristics that people develop from their everyday experiences.
-Based on this assumption, Jenny et al.
-(2018) were able to demonstrate that students judge probability significantly better, when language is adopted that better aligns with the heuristics that most learners develop from their everyday experiences.
-In sum, changing representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that these support instead of inhibit learning seems feasible.
+Based on this assumption, \textcite{jkg18} were able to demonstrate that students judge probability significantly better, when language is adopted that better aligns with the heuristics that most learners develop from their everyday experiences.
+In sum, it seems feasible to change representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that these support instead of inhibit learning.
 
@@
 
-Recent work in cognitive and psychological science on how children learn argues that children understand and learn about the world through the intuitive application of Bayesian ideas (Gopnik, 2012; Tenenbaum et al., 2006; Griffiths et al., 2008).
+Recent work in cognitive and psychological science on how children learn argues that children understand and learn about the world through the intuitive application of Bayesian ideas \parencite{g12, tgk06, tgk06}.
 This idea has led current research to conceptualize human learning in terms of Bayesian probabilistic models of learning.
@@
 
-This view of how children learn from data has become especially prominent in developmental science and is successfully applied there (Gopnik and Tenenbaum, 2007; Gopnik & Wellman, 2012).
+This view of how children learn from data has become especially prominent in developmental science and is successfully applied there \parencite{gt07, gw12}.
 One might assume that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education.
 Developmentalists have pointed to the potential and opportunity here.
-Gopnik (2012) wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p.
+\textcite{g12} wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p.
 1627).
 
-Despite arguments for integrating developmental and educational research, it has largely not been the case that—with respect to advances in Bayesian methods—“science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” (Gopnik, 2012, p.
-1627), as some science education scholars have bemoaned (Lehrer & Schauble, 2006).
-Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas (Piaget & Inhelder, 1969) that represent the greatest contribution of developmental ideas to education (Lehrer & Schauble, 2007; Metz, 1995).
+Despite arguments for integrating developmental and educational research, it has largely not been the case that—with respect to advances in Bayesian methods—“science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” \textcite[p.
+1627]{g12}, as some science education scholars have bemoaned \parencite{ls15}).
+Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas \textcite{pi69} that represent the greatest contribution of developmental ideas to education \textcite{ls07}.
 We argue that this is still largely the case in the present, despite advances in both developmental research as well as statistical methodology related to Bayesian ideas—and that it is time to start exploring how bringing out the inherent Bayesian ways of thinking in education may provide support for learners’ efforts to make sense of the world.
@@
 Related science education research emphasizes the role of prior knowledge that students bring with them to the analysis of data and how that prior knowledge impacts the inferences they make about data.
-For example, Masnick et al.
-(2007) summarized findings across a number of studies showing that when elementary grade students reasoned with data in a context about which they had strong and valid prior understandings that informed their interpretation of data (about force and motion in the context of ramps), and they were able to explain sources of error in the data (Masnick & Klahr, 2003), but when the data was in a context they did not understand well, they struggled to distinguish between random variation due to sources of error and sources of variation due to something that was causing a difference in the data (generated by measuring the motion of pendulums).
-Later design- and laboratory-based research has shown that students can develop an understanding of scientific ideas through the analysis of data (Lehrer & Schauble, 2004; Masnick et al., 2017).
+For example, \textcite{mkm07} summarized findings across a number of studies showing that when elementary grade students reasoned with data in a context about which they had strong and valid prior understandings that informed their interpretation of data (about force and motion in the context of ramps), and they were able to explain sources of error in the data \parencite{mk03}, but when the data was in a context they did not understand well, they struggled to distinguish between random variation due to sources of error and sources of variation due to something that was causing a difference in the data (generated by measuring the motion of pendulums).
+Later design- and laboratory-based research has shown that students can develop an understanding of scientific ideas through the analysis of data \parencite{ls04,mkk17}.
 
-Laboratory-based research, especially, has examined how learners’ prior understanding may have a bearing on how they make sense of how likely different outcomes are (Klahr & Dunbar, 1998; Masnick et al., 2007; Schwartz et al., 2007).
+Laboratory-based research, especially, has examined how learners’ prior understanding may have a bearing on how they make sense of how likely different outcomes are \parencite{kd88, mkm07, ssc07}.
 While these studies focused on how students’ prior ideas and data interact, Bayesian frameworks were not used to formalize this process of integrating data in light of prior beliefs about scientific phenomena.
-In short, this work aligns closely with the Bayesian models of how children learn from data described by Gopnik (2012), but this research does not explicitly interpret their findings in light of these ideas about how children learn from data in a Bayesian way.
+In short, this work aligns closely with the Bayesian models of how children learn from data described by \textcite{g12}, but this research does not explicitly interpret their findings in light of these ideas about how children learn from data in a Bayesian way.
 
-Overall, developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in science education to design instruction based around eliciting and understanding students’ ideas (Gotwals & Birmingham, 2012; Haverly et al., 2019; Windschitl et al., 2012).
+Overall, developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in science education to design instruction based around eliciting and understanding students’ ideas \parencite{gb16, hbavb20, wtbs12}.
 There is an opportunity to use Bayesian ideas as a framework for formalizing these ideas about how prior ideas affect students’ reasoning about data and how teachers can leverage students’ ideas to design instruction, yet Bayes has mostly been absent from research in science education.
@@
 Having described the role of probability and uncertainty in science and science education, we now describe what these Bayesian ideas can mean for science literacy.
-We see two primary ways in which Bayesian methods can enhance science literacy (National Academies of Sciences, Engineering, and Medicine, 2016), or a range of capabilities that people may possess which have value (individually or collectively) to them.
-These capabilities range from foundational literacies, including numeracy and textual literacy (what is most thought of concerning literacy in common parlance), content knowledge, and epistemic knowledge and other ways of thinking and habits of mind (National Academies of Sciences, Engineering, and Medicine, 2016).
+We see two primary ways in which Bayesian methods can enhance science literacy, or a range of capabilities that people may possess which have value (individually or collectively) to them \parencite{council2016science}.
+These capabilities range from foundational literacies, including numeracy and textual literacy (what is most thought of concerning literacy in common parlance), content knowledge, and epistemic knowledge and other ways of thinking and habits of mind \parencite{council2016science}.
 One way that Bayesian ideas can inform discussions of science literacy is to provide individuals with a technique for combining their prior understanding with new evidence.
@@
 
-The history of science is, in many ways, a story of established ideas being built upon or supplanted (Fara, 2010), and the contemporary emphasis on viewing science as a practice was sparked by Kuhn’s (1962) historical-epistemological account on what this suggests about the nature and philosophy of science (Godfrey-Smith, 2009).
-Yet, our ways of interpreting scientific evidence do rarely recognize this, perhaps for good reason: Most of the time, new data or information do not contradict foundational scientific ideas (Kuhn, 1962).
-The same phenomenon, though, is present on a more micro-level: While scientists may be familiar with the data they collect not eliding with ideas that are a part of the scientific consensus, in media and the public discourse scientific ideas are often presented as fixed and certain—even in cases for which consensus has not been reached (Collins, 1987).
+The history of science is, in many ways, a story of established ideas being built upon or supplanted \parencite{fara2010science}, and the contemporary emphasis on viewing science as a practice was sparked by \textcite{k62} historical-epistemological account on what this suggests about the nature and philosophy of science \parencite{g09}.
+Yet, our ways of interpreting scientific evidence do rarely recognize this, perhaps for good reason: Most of the time, new data or information do not contradict foundational scientific ideas \parencite{k62}.
+The same phenomenon, though, is present on a more micro-level: While scientists may be familiar with the data they collect not eliding with ideas that are a part of the scientific consensus, in media and the public discourse scientific ideas are often presented as fixed and certain—even in cases for which consensus has not been reached \parencite{c87}
 
@@
 We argue that there are three ways that Bayesian methods could support science literacy.
-First, the result is a non-binary inference, one that doesn’t speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty.
+
+\JR{may cut these depending on what E.J. adds below}
+\EJ{OK I added a bunch of stuff. Feel free to ignore, cut, rewrite, integrate with other materials etc. Here is where my stuff starts...}
+
+The famous French astronomer and mathematician Pierre-Simon Laplace argued that Bayes' rule quantifies how rational observers ought to change their opinion in light of new evidence; in this sense, Laplace believed that Bayes' rule was \emph{common sense expressed in numbers}.
+We now showcase several insights that follow directly from Bayes' rule.
+In our earlier exposition of Bayes' rule we already emphasized the idea that hypotheses gain credibility when they predict the data well, and lose credibility when they predict new data poorly \parencite{WagenmakersEtAl2016CD}.
+This idea generalizes what the mathematician George Polya termed the ``Fundamental Inductive Pattern'':
+\begin{quotation}
+``This inductive pattern says nothing surprising.
+On the contrary, it expresses a belief which no reasonable person seems to doubt: \emph{The verification of a consequence renders a conjecture more credible}.
+With a little attention, we can observe countless reasonings in everyday life, in the law courts, in science, etc., which appear to confirm to our pattern.'' \parencite[pp.
+4-5]{Polya1954Vol2}
+\end{quotation}
+
+The rule from Equation~\ref{eq:BayesRule} includes the constant term $p(\text{data})$, which does not involve $\theta$. Hence, we can also write Bayes' rule as:
+\begin{equation}
+\label{eq:BayesRulePropto}
+p(\theta \mid \text{data}) \propto  p(\text{data} \mid \theta) \times p(\theta),
+\end{equation}
+where `$\propto$' stands for `is proportional to'.
+Thus, Bayes' rule states that our posterior knowledge $p(\theta \given \text{data})$ is proportional to the likelihood $p(\text{data} \given \theta)$ (i.e., predictive adequacy, unsurprise, or the extent to which the observed data are expected given $\theta$) multiplied with our prior knowledge $p(\theta)$.
+
+Equation~\ref{eq:BayesRulePropto} emphasizes the fact that the posterior uncertainty about $\theta$ is a \emph{compromise} between our prior uncertainty about $\theta$ and the predictive performance of $\theta$.
+But the posterior uncertainty after having observing the first datum becomes the prior uncertainty for the to-be-observed second datum (cf.
+Figure~\ref{fig:WindowsMacSequential}).
+Consequently, after having observed the second datum, the posterior uncertainty represents a compromise of a compromise.
+As the data accumulate, the posterior uncertainty is more and more determined by predictive performance, and the impact of the initial uncertainty about $\theta$ is increasingly watered down: `the data overwhelm the prior' \parencite{WrinchJeffreys1919}.
+
+This implies that a constant stream of accumulating data ought to bring any two rational people into an arbitrarily close agreement, no matter how divergent their opinions may have been at the outset.
+There are two important caveats here.
+First and foremost, the data only overwhelm the prior if that prior is neither 0 (denoting an impossibility) nor 1 (denoting absolute certainty).
+If you already know for certain that a hypothesis is true or false, then you cannot adjust your opinion in light of the data: your initial opinion will be your final opinion, no matter what the data may indicate.
+Philosophically, adopting such priors is a dangerous practice; for instance, in antiquity the adage of the \emph{New Academy} --a school of skeptics headed by Carneades-- was ``never assert absolutely''.
+In modern times, Lindley popularized this idea in statistics and coined it ``Cromwell's rule'':
+\begin{quotation}
+``Consequently, an uncertain event of zero probability remains so whatever information is provided.
+In other words, if a decision-maker thinks something cannot be true and interprets this to mean it has zero probability, he will never be influenced by \emph{any} data, which is surely absurd.
+So leave a little probability for the moon being made of green cheese; it can be as small as 1 in a million, but have it there since otherwise an army of astronauts returning with samples of the said cheese will leave you unmoved.
+A probability of one is equally dangerous because then the probability of $\bar{E}$ will be zero.
+So never believe in anything absolutely, leave some room for doubt: as Oliver Cromwell told the Church of Scotland, `I beseech you, in the bowels of Christ, think it possible you may be mistaken'.'' \parencite[p.
+104]{Lindley1985}
+\end{quotation}
+
+The second caveat is that, in real life, the data are sometimes relatively slow to overwhelm the prior, as people can be reluctant to change their beliefs (for a Bayesian model see \cite{Gershman2019}).
+A more extreme case is \emph{belief polarization}: confronted with the same stream of information, two people who hold different opinions may drift further apart instead of moving closer together (but see \cite{Anglin2019}).
+This appears irrational, but several Bayesian accounts have been offered to explain the phenomenon (e.g., \cite{CookLewandowsky2016,JernEtAl2014}).
+
+Now consider two specific values for $\theta$, $\theta_1$ and $\theta_2$. We can use Bayes' rule to obtain the posterior probability for each one:
+\begin{align}
+p(\theta_1 \mid \text{data})& =  \frac{p(\text{data} \mid \theta_1) \times p(\theta_1)}{p(\text{data})}\\
+p(\theta_2 \mid \text{data})& =  \frac{p(\text{data} \mid \theta_2) \times p(\theta_2)}{p(\text{data})}.
+\end{align}
+When we divide the posterior probabilities the common term $p(\text{data})$ drop out and we have:
+\begin{equation}
+    \frac{p(\theta_1 \mid \text{data})}{p(\theta_2 \mid \text{data})} = \frac{p(\theta_1)}{p(\theta_2)} \, \times \, \frac{p(\text{data} \mid \theta_1)}{p(\text{data} \mid \theta_2)}.
+\end{equation}
+We replace $\theta_2$ with $\mathcal{H}_1$ (i.e., the alternative hypothesis, in which a test-relevant parameter $\xi$ is free to vary) and $\theta_1$ with $\mathcal{H}_0$ (i.e., the null hypothesis in which $\xi$ takes on a fixed value, for instance $\xi=0$):
+\begin{equation}
+    \frac{p(\mathcal{H}_1 \mid \text{data})}{p(\mathcal{H}_0 \mid \text{data})} = \frac{p(\mathcal{H}_1)}{p(\mathcal{H}_0)} \, \times \, \frac{p(\text{data} \mid \mathcal{H}_1)}{p(\text{data} \mid \mathcal{H}_0)}.
+\end{equation}
+This way of writing Bayes' rule highlights that extraordinary claims require extraordinary evidence -- if a particular hypothesis $\mathcal{H}_1$ is extremely unlikely a priori, the prior odds $\frac{p(\mathcal{H}_1)}{p(\mathcal{H}_0)}$ are stacked against it, and in order for the posterior odds to favor $\mathcal{H}_1$ the support from the data (i.e, the degree to which $\mathcal{H}_1$ outpredicts $\mathcal{H}_0$) needs to be overwhelmingly strong.
+
+Bayes' rule can also be shown to embody the \emph{principle of parsimony}: the rule implicitly contains a preference for the simplest model that explains the data well.
+To see this, consider a coin taken from a magician's coffin, and let parameter $\xi$ indicate the chance that the coin lands heads on any throw.
+The null hypothesis holds that the coin is fair, $\mathcal{H}_0: \xi = \nicefrac{1}{2}$.
+The alternative hypothesis that we entertain here specifies that the coin may be double-tails, fair, or double-heads, $\mathcal{H}_1: \xi \in \{0, \nicefrac{1}{2}, 1\}$, with each of the three options deemed equally likely a priori.
+
+The coin is tossed and we observe heads.
+The probability of this datum is \nicefrac{1}{2} under $\mathcal{H}_0$; under $\mathcal{H}_1$, it is $p(\text{heads} \given \xi=0) \cdot p(\xi=0 \given \mathcal{H}_1) + p(\text{heads} \given \xi=\nicefrac{1}{2}) \cdot p(\xi=\nicefrac{1}{2} \given \mathcal{H}_1) + p(\text{heads} \given \xi=1) \cdot p(\xi=1 \given \mathcal{H}_1) = 0 \cdot \nicefrac{1}{3}  + \nicefrac{1}{2} \cdot \nicefrac{1}{3} + 1 \cdot \nicefrac{1}{3} = \nicefrac{1}{2}$.
+So the datum is equally likely under $\mathcal{H}_0$ and $\mathcal{H}_1$: both models receive an equal amount of support.
+This means the first outcome does not change our conviction concerning $\mathcal{H}_0$ versus $\mathcal{H}_1$.
+The datum did, however, change our beliefs about $\xi$ under $\mathcal{H}_1$.
+Specifically, we now know that $\xi$ cannot be 0; moreover, the datum was twice as likely under $\xi=1$ as under $\xi = \nicefrac{1}{2}$, so that our posterior distribution for $\xi$ under $\mathcal{H}_1$ is now $p(\xi = \nicefrac{1}{2}) = \nicefrac{1}{3}, p(\xi = 1) = \nicefrac{2}{3}$.
+
+The coin is tossed a second time and it lands tails.
+Under $\mathcal{H}_0$, the probability of this happening is again $\nicefrac{1}{2}$, so the total probability for the data sequence $\{\text{heads}, \text{tails}\}$ under $\mathcal{H}_0$ equals $\nicefrac{1}{2} \cdot \nicefrac{1}{2} = \nicefrac{1}{4}$.
+Under $\mathcal{H}_1$, the probability of the second toss landing heads is computed under the posterior distribution obtained after the first toss, and this yields: $p(\text{tails} \given \xi = \nicefrac{1}{2}) \cdot p(\xi=\nicefrac{1}{2} \given \mathcal{H}_1) + p(\text{tails} \given \xi = 1) \cdot p(\xi=1 \given \mathcal{H}_1) = \nicefrac{1}{2} \cdot \nicefrac{1}{3} + 0 \cdot \nicefrac{2}{3} = \nicefrac{1}{6}$.
+The total probability for the data sequence $\{\text{heads}, \text{tails}\}$ under $\mathcal{H}_1$ equals $\nicefrac{1}{2} \cdot \nicefrac{1}{6} = \nicefrac{1}{12}$.
+This means that the observed data provided $(\nicefrac{1}{4}) / (\nicefrac{1}{12}) = 3$ times more support for $\mathcal{H}_0$ than for $\mathcal{H}_1$.
+This happens because $\mathcal{H}_1$ spreads out its predictions, hedging its bets.
+In contrast, the simple model $\mathcal{H}_0$ made a precise prediction.
+
+Instead of learning about the predictive performance of $\mathcal{H}_0$ and $\mathcal{H}_1$ one observation at a time, we could also have considered the probability that the models assign to the entire sequence $\{\text{heads}, \text{tails}\}$.
+Under $\mathcal{H}_0$ we again have $\nicefrac{1}{2} \cdot \nicefrac{1}{2} = \nicefrac{1}{4}$.
+Under $\mathcal{H}_1$, we notice that the data falsify both $\xi=0$ and $\xi=1$.
+This leaves $\xi=\nicefrac{1}{2}$, which suggests the same answer as under $\mathcal{H}_0$, that is $\nicefrac{1}{2} \cdot \nicefrac{1}{2} = \nicefrac{1}{4}$.
+However, we need to multiply this probability with $\nicefrac{1}{3}$, the prior probability that $\xi=\nicefrac{1}{2}$.
+This is the penalty for complexity that $\mathcal{H}_1$ pays for entertaining three different values of $\xi$ from the outset.
+
+Thus, Bayes rule includes a automatic ``Ockham's razor'' \parencite{Jeffreys1939,JefferysBerger1992} in the sense that daring predictions are rewarded when they come true.
+
+Finally, we can also write Bayes' rule as follows:
+\begin{equation}
+\begin{split}
+    p(\mathcal{H}_1 \mid \text{data})& = \frac{p(\text{data} \mid \mathcal{H}_1) p(\mathcal{H}_1)}{p(\text{data})}\\
+    & = \frac{p(\text{data} \mid \mathcal{H}_1) p(\mathcal{H}_1)}{p(\text{data} \mid \mathcal{H}_1) p(\mathcal{H}_1) + p(\text{data} \mid \mathcal{H}_0) p(\mathcal{H}_0)}.
+\end{split}
+\end{equation}
+This equation shows that the posterior plausibility is dictated by the predictive performance for $\mathcal{H}_1$ and $\mathcal{H}_0$, weighted by the prior plausibility of each hypothesis.
+In other words, ``The Bayesian world is a comparative world in which there are no absolutes.'' \parencite[p.
+308]{Lindley2000}.
+This is in stark contrast to $p$-value statistical hypothesis testing, in which a statistical model (the null hypothesis) is judged in isolation.
+\EJ{...and here it ends.}
+
+First, the result is a non-binary inference, one that doesn't speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty.
 Though a small difference, discussing the safety of vaccines on these terms can provide a more solid foundation for principled disagreements and arguments.
@@
 
-I think there are several very general ideas that could be illustrated with Bayes’ rule. These include:
+%I think there are several very general ideas that could be illustrated with Bayes’ rule. These include:
 
-\begin{enumerate}
-    \item When we gain new information, our updated knowledge of the world is a compromise between our old knowledge and the new information.
-    \item Today’s posterior is tomorrow’s prior 
-    \item Hypotheses/claims that predict new data well receive a boost in credibility, whereas those that predict the data poorly suffer a decline.
-    \item All knowledge is relative
-    \item Avoid absolute beliefs (Cromwell’s rule; these are not amenable to learning)
-    \item Surprise lost is credibility gained
-    \item Polya’s fundamental inductive pattern: “This inductive pattern says nothing surprising.
+%\begin{enumerate}
+%    \item When we gain new information, our updated knowledge of the world is a compromise between our old knowledge and the new information.
+%    \item Today’s posterior is tomorrow’s prior 
+%    \item Hypotheses/claims that predict new data well receive a boost in credibility, whereas those that predict the data poorly suffer a decline.
+%    \item All knowledge is relative
+%    \item Avoid absolute beliefs (Cromwell’s rule; these are not amenable to learning)
+%    \item Surprise lost is credibility gained
+%    \item Polya’s fundamental inductive pattern: ``This inductive pattern says nothing surprising.
 On the contrary, it expresses a belief which no reasonable person seems to doubt: The verification of a consequence renders a conjecture more credible.
-With a little attention, we can observe countless reasonings in everyday life, in the law courts, in science, etc., which appear to confirm to our pattern.” (Polya, 1954, pp.
+With a little attention, we can observe countless reasonings in everyday life, in the law courts, in science, etc., which appear to confirm to our pattern.'' (Polya, 1954, pp.
 4-5)
-    \item Extraordinary claims require extraordinary evidence
-\end{enumerate}
+%    \item Extraordinary claims require extraordinary evidence
+%\end{enumerate}
 
@@
 
-Research on the role of Bayes in science education has mostly been carried out in the context of undergraduate statistics education, in which there have been a number of calls to action (Gould et al., 2018; Hoegh et al., 2020), examples and design options (Albert, 2002; Bolstad, 2002; Hu, 2020; Gelman, 2008; Witmer, 2017) and debate (Johnson et al., 2020) over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs.
-The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power (Gould et al., 2018) as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers (Albert & Hu, 2020).
-As evidenced by the recent special issue of the Journal of Statistics Education of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.
+Research on the role of Bayes in science education has mostly been carried out in the context of undergraduate statistics education, in which there have been a number of calls to action \parencite{gpkpwc18, h_a20}, examples and design options \parencite{a02, b02,g08, w17, h_j} and debate \parencite{jrhrr20} over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs.
+The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power \parencite{gpkpwc18} as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers \parencite{ah20} .As evidenced by the recent special issue of the \emph{Journal of Statistics Education} of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.
 
@@
 Two publications merit mention though.
-The work of Szu and Osborne (2012) and Nussbaum (2011) both applied Bayesian perspectives to the science practice of argumentation.
+The work of \textcite{so12} and \textcite{n11} both applied Bayesian perspectives to the science practice of argumentation.
 Szu and Osborne presented the case for how and why Bayes Theorem can apply to research and practice on scientific reasoning, broadly, and argumentation, particularly.
 This paper represents the most comprehensive account of the relevance of Bayesian methods for science education.
-Szu and Osborne argue that Bayes is useful as both a formal mathematical tool, and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes Theorem that follow more or less deterministically from “external, objectively probabilistic systems” (p.
-61)—is “the key leap that characterizes the debate about the value of Bayesian inference as a model of scientific reasoning” (p.
+Szu and Osborne argue that Bayes is useful as both a formal mathematical tool, and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes Theorem that follow more or less deterministically from ``external, objectively probabilistic systems'' (p.
+61)—is ``the key leap that characterizes the debate about the value of Bayesian inference as a model of scientific reasoning'' (p.
 61).
-In this way, Szu and Osborne argue that the greatest use of Bayes theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community (e.g., Batanero et al., 2016; Makar & Rubin, 2018).
+In this way, Szu and Osborne argue that the greatest use of Bayes theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community \parencite{batanero2016research, mr18}.
 They offer some research-related backing for the use of Bayes Theorem (e.g., noting how conceptual change research, particularly, and constructivist research, broadly, can both be explained coherently within a Bayesian approach) as well as some practical, instructional recommendations, some of which we detail later in this section.
 
-Nussbaum’s (2011) work on Bayesian approaches to argumentation complement the work of Szu and Osborne in that Nussbaum describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation.
+\textcite{n11} work on Bayesian approaches to argumentation complement the work of Szu and Osborne in that Nussbaum describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation.
 Concerning the latter, Nussbaum describes how the social issue of raising taxes to provide resources to homeless individuals was a rich context for students to engage in forms of Bayesian reasoning.
-In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what Szu and Osborne (2012) characterize as the more externally objective use of Bayes Theorem—and how the estimates that result from applying Bayes theorem led students to re-evaluate their initial arguments.
+In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what \textcite{so12} characterize as the more externally objective use of Bayes Theorem—and how the estimates that result from applying Bayes theorem led students to re-evaluate their initial arguments.
 
-In summary, the work of Szu and Osborne (2012) and Nussbaum begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process.
+In summary, the work of \textcite{so12} and \textcite{n11} begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process.
 Curiously, no work has applied Bayesian methods to another science practice, that of analyzing and interpreting data; one for which Bayesian methods have clear conceptual applicability and utility, but one for which the absence of suitable tools may provide to be a barrier.
-In this way, Bayesian methods could serve as a cross-cutting concept (National Research Council, 2012) that unites the practice of argumentation with analyzing and interpreting data, and using mathematics and computational thinking.
+In this way, Bayesian methods could serve as a cross-cutting concept \parencite{nrc12} that unites the practice of argumentation with analyzing and interpreting data, and using mathematics and computational thinking.
 
@@
 
-First, we should consider “unplugged” versions of Bayesian analyses, akin to how computer science education scholars have emphasized that computational thinking does not necessarily require the use of a computer (Yadav et al., 2018).
-These could interface with the science and engineering practice of modeling; could students use diagrammatic models in conjunction with Bayesian “heuristics” to revise their models in a similar manner to how this practice can unfold in classrooms (see Schwarz et al., 2009).
+First, we should consider ``unplugged'' versions of Bayesian analyses, akin to how computer science education scholars have emphasized that computational thinking does not necessarily require the use of a computer (Yadav et al., 2018).
+These could interface with the science and engineering practice of modeling; could students use diagrammatic models in conjunction with Bayesian “heuristics” to revise their models in a similar manner to how this practice can unfold in classrooms \parencite{schwarz2009developing}.
 
-Another advance should be to build out curricular strategies, elaborating on the ideas of Szu and Osborne (2011) and Nussbaum (2011).
-At a high-level, Szu and Osborne (2011) write that in order to reason about data in light of their initial beliefs (and to construct a kind of informal likelihood statistic), students “need to see judgments about data and evidence being an assessment not only of the probability of the hypothesis being correct but also of it being wrong” (p.
+Another advance should be to build out curricular strategies, elaborating on the ideas of \textcite{so12} and \textcite{n11}.
+At a high-level, Szu and Osborne write that in order to reason about data in light of their initial beliefs (and to construct a kind of informal likelihood statistic), students “need to see judgments about data and evidence being an assessment not only of the probability of the hypothesis being correct but also of it being wrong” (p.
 67).
@@
 These are helpful suggestions, and they complement finer-grained instructional techniques, especially those around how students can reason about uncertainty probabilistically.
-Particularly, Binder et al.
-(2018), for instance, suggest replacing probabilities with natural frequencies (in alignment with a great deal of work on how humans understand probability; Gigerenzer & Hoffrange, 1995) as well as using and interpreting tree diagrams to depict natural frequencies.
-Martignon and Erickson (2014)—building on much earlier experimental work by Wason and Shapiro (1966)—provide a simple example of how even fourth graders, with the right tools to represent probabilities, can support students to transition from logical to probabilistic reasoning.
+Particularly, \textcite{bkbm18} for instance, suggest replacing probabilities with natural frequencies in alignment with a great deal of work on how humans understand probability) \parencite{gh95} as well as using and interpreting tree diagrams to depict natural frequencies.
+\textcite{me14}—building on much earlier experimental work by \textcite{wason1971natural}—provide a simple example of how even fourth graders, with the right tools to represent probabilities, can support students to transition from logical to probabilistic reasoning.
 These could also support the aim of developing unplugged versions of Bayesian analyses.
@@
 Third, we should, of course, consider teachers and how teachers can (or already) support students to reason about uncertainty, and, possibly, to use Bayesian ideas.
-While some scholars have contributed foundational work demonstrating how this uncertainty can be productive as a part of teaching practice (Manz & Suarez, 2018), a broad and formal perspective on how uncertainty can be approached in science education is needed.
+While some scholars have contributed foundational work demonstrating how this uncertainty can be productive as a part of teaching practice \parencite{manz2018supporting}, a broad and formal perspective on how uncertainty can be approached in science education is needed.
 There is the need to explore the potential of a principled approach to uncertainty, one that may be intuitive to learners and useful to teachers, but is as yet under-represented in the science education literature.
@@
 JASP allows for both conventional frequentist as well as Bayesian versions of commonly used statistical tests, such as t-tests and regression analyses.
-There are a number of accessible guides available for specifying Bayesian versions of analyses (van Doorn et al., 2020; Wagenmakers et al., 2018), and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms.
-Another possibly useful tool could be Bayesian networks (see Bayes Box; Bayes Box, 2021), which could be used to model complex system in a way that is similar to other tools used in science education for modeling complex systems, such as SageModeler (2020).
+There are a number of accessible guides available for specifying Bayesian versions of analyses \parencite{v20, wagenmakers2018bayesian}, and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms.
+Another possibly useful tool could be Bayesian networks (see Bayes Box; Bayes Box, 2021), which could be used to model complex system in a way that is similar to other tools used in science education for modeling complex systems, such as \textcite{20}.
 
@@
 
-While some educational researchers may state that such theoretical, abstract ideas about the philosophical and statistical underpinnings of working with data are irrelevant to how students work with data, we make the case, here, that these ideas can inform instructional practice, assessments, and the design of activities that involve students in working with data, especially given some of the clear problems that have been articulated around the present frequentist Null Hypothesis Statistical Testing (NHST)-based approaches (Cumming, 2014; Gigerenzer, 2004).
+While some educational researchers may state that such theoretical, abstract ideas about the philosophical and statistical underpinnings of working with data are irrelevant to how students work with data, we make the case, here, that these ideas can inform instructional practice, assessments, and the design of activities that involve students in working with data, especially given some of the clear problems that have been articulated around the present frequentist Null Hypothesis Statistical Testing (NHST)-based approaches \parencite{c14, gkv04}.
 
-Science education researchers may note that these ideas are conceptually-focused, rather than aligning with the present emphasis in science education upon engaging students in science and engineering practice in order to make sense of core ideas and to engage with cross-cutting concepts (National Research Council, 2012).
-We argue that Bayes is unique in that it, perhaps akin to a cross-cutting concept in the sense of a style of reasoning (Szu & Osborne, 2012; Osborne et al.
-2018) that blends conceptual ideas with practices.
+Science education researchers may note that these ideas are conceptually-focused, rather than aligning with the present emphasis in science education upon engaging students in science and engineering practice in order to make sense of core ideas and to engage with cross-cutting concepts \parencite{nrc12}.
+We argue that Bayes is unique in that it, perhaps akin to a cross-cutting concept in the sense of a style of reasoning \parencite{so12, ork18} that blends conceptual ideas with practices.
 Bayes is both a philosophical and psychological perspective on how to learn from data as well as a statistical framework.
@@
 In this paper, we have introduced Bayes as a powerful way of conceptualizing probability and uncertainty.
-These are two concepts that are ubiquitous in the learning, doining, and communicating of science and therefore critical aspects of science literacy (National Academies of Sciences, Engineering, and Medicine, 2016).
-However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike (Gigerenzer et al.
-2004; Sedlmeier, 2007; Tversky & Kahneman, 1977).
+These are two concepts that are ubiquitous in the learning, doing, and communicating of science and therefore critical aspects of science literacy \parencite{}.
+However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike \parencite{gkv04, s07, tk74}.
 Thus, it is not surprising when the public misinterprets statements about probability and uncertainty but the consequences may be dire when trust in science is eroded, and, sadly, examples of this are plentiful during current the COVID-19, e.g., the efficiency of masks in preventing the spread of COVID and the efficiency of the AstraZeneca vaccine were strongly questioned, delaying efforts to comb at the pandemic.
 The public, however, can hardly be blamed for this in light of the prevalent findings that the frequentist conceptualizations of probability and uncertainty in school are challenging and uncommonly align with or build on peoples’ intuitions.
-As research from cognitive and psychological science as well as emerging work from statistics and statistics data science education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions (Kruschke & Liddell, 2018a; Szu & Osborne, 2012) about these concepts and thus has potential to support learning about probability and uncertainty in science.
+As research from cognitive and psychological science as well as emerging work from statistics and statistics data science education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions \parencite{kl18, so12} about these concepts and thus has potential to support learning about probability and uncertainty in science.
 

### 2021-04-29 manuscript.tex
update

@@
 
-\EJ{I have doubts about the ``Trust me, I'm a Bayesian'' part of the title, also because ``Trust'' now occurs twice.
-Maybe simply ``Building Trust in Science Through a Bayesian Perspective on Probability and Uncertainty''?}
-\MK{Ok with the new title - I liked the sound of the trust me thing but utlimatley it may not be really on point!}
-\EJ{I also uploaded my bibtex file, ``referenties.bib''.
-I could not get it to work, but you could copy the relevant references into your own bibtex file.
-Actually, I also managed to make the initial bibtex setup not work, and I have absolutely no clue as to how this happened.}
-
 Uncertainty is ubiquitous in science and this uncertainty takes many forms.
@@
 
-\EJ{Yeah so as a hard-nosed determinist I still have an issue with ``inherent'' uncertainty.
-If we don't sort this out we could resolve it through a footnote (e.g., ``EJ Wagenmakers would like to note that he is a determinist who believes that the concepts of probability and uncertainty belongs wholly to the mind \parencite{Jevons18741913}, with the possible exception of the strange world of quantum mechanics.''}
-\MK{I got rid of the inherent part.}
-
-\EJ{It would be good to go over the reference list in detail; in particular, I noticed several ``bayes'' that should be ``Bayes'' (the B needs to be enclosed in brackets so it displays in upper case)}
 Negotiating these many forms of uncertainty in argumentation—and building toward consensus at the same time that uncertainty is present—is a key part of the scientific process \parencite{t00}.
@@
 
-The remainder of this paper is structured in the following way.
-We will first briefly introduce Bayes theorem and wider ideas around Bayesian epistemology.
+The remainder of this paper is structured as follows.
+First, we will introduce Bayes theorem and wider ideas around Bayesian epistemology.
 Next, we will elaborate on the importance of probability and uncertainty in science (education) and review how people learn about these concepts and where they face difficulties.
@@
 To make this more tangible, consider the following example.
-Suppose you are interested in the proportion of students that use Windows vs.
-MacOS as their operating system.
+Suppose an educator is interested in the proportion of students that use Windows or MacOS as their operating system.
 The proportion of Windows users $\theta$ ranges between 0 and 1.
 Before you start our Bayesian analysis, you need to quantify your background knowledge -- the relative plausibility of the different values for $\theta$.
-Given your previous experiences in class rooms your best guess is that about half of the students use MacOS.
-You would not be surprised if the  proportion of Windows users was between 30\% and 70\% but you would be rather surprised to find either no Windows or exclusive Windows usage.
+Given their previous experiences as a teacher, their best guess is that about half of the students use MacOS.
+They would not be surprised if the proportion of Windows users was between 30\% and 70\% but you would be rather surprised to find either no Windows or exclusive Windows usage.
+
 This background knowledge motivates the use of the dome-shaped prior distribution shown as the dashed line in Figure~\ref{fig:WindowsMacPriorPosterior}.\footnote{This is the beta$(2,2)$ distribution.} Now suppose you start to collect data by asking students about their operating systems.
 The first ten observations of Windows (W) and MacOS (M) usage that you make are $\{W, W, M, W, M, W, W, M, W, M\}$: six students use Windows, four use MacOS.
+
 With these data in hand, we then combine our prior knowledge (i.e., the dome-shaped prior distribution) with the predictive updating factor (assuming a binomial likelihood) and obtain the posterior distribution shown as the solid line in Figure~\ref{fig:WindowsMacPriorPosterior}.
@@
 
-But how do the principles from this example translate to science (education)?
-A recent case from physics can serve to illustrate this.
+How do the principles from this example translate to students' work with data and evidence -- not only educators?
+A recent case from physics can serve to illustrate this approach in the context of students' activity in science classrooms.
 In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light \parencite{b11} - and thus defying Einstein’s theory of relativity.
@@
 
-\JR{may cut these depending on what E.J. adds below}
-\EJ{OK I added a bunch of stuff. Feel free to ignore, cut, rewrite, integrate with other materials etc. Here is where my stuff starts...}
-
-The famous French astronomer and mathematician Pierre-Simon Laplace argued that Bayes' rule quantifies how rational observers ought to change their opinion in light of new evidence; in this sense, Laplace believed that Bayes' rule was \emph{common sense expressed in numbers}.
+The famous French astronomer and mathematician Laplace argued that Bayes' rule quantifies how rational observers ought to change their opinion in light of new evidence; in this sense, Laplace believed that Bayes' rule was \emph{common sense expressed in numbers}.
 We now showcase several insights that follow directly from Bayes' rule.
 In our earlier exposition of Bayes' rule we already emphasized the idea that hypotheses gain credibility when they predict the data well, and lose credibility when they predict new data poorly \parencite{WagenmakersEtAl2016CD}.
-This idea generalizes what the mathematician George Polya termed the ``Fundamental Inductive Pattern'':
+This idea generalizes what the mathematician Polya termed the ``Fundamental Inductive Pattern'':
 \begin{quotation}
@@
 This is in stark contrast to $p$-value statistical hypothesis testing, in which a statistical model (the null hypothesis) is judged in isolation.
-\EJ{...and here it ends.}
 
-First, the result is a non-binary inference, one that doesn't speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty.
+Thus, in summary, the result of applying Bayes rule is a non-binary inference, one that doesn't speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty.
 Though a small difference, discussing the safety of vaccines on these terms can provide a more solid foundation for principled disagreements and arguments.
-
-Second, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in-line with how most of us likely already interpret data.
+In addition, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in-line with how most of us likely already interpret data.
 Specifically Bayes provides people with a way to talk about weighing between ideas and their prior beliefs and new information and data.
 In short, Bayes gives people a chance to deal with uncertainty in a principled but accessible way.
-
-Last, his process could be formalized mathematically.
-
-%I think there are several very general ideas that could be illustrated with Bayes’ rule. These include:
-
-%\begin{enumerate}
-%    \item When we gain new information, our updated knowledge of the world is a compromise between our old knowledge and the new information.
-%    \item Today’s posterior is tomorrow’s prior 
-%    \item Hypotheses/claims that predict new data well receive a boost in credibility, whereas those that predict the data poorly suffer a decline.
-%    \item All knowledge is relative
-%    \item Avoid absolute beliefs (Cromwell’s rule; these are not amenable to learning)
-%    \item Surprise lost is credibility gained
-%    \item Polya’s fundamental inductive pattern: ``This inductive pattern says nothing surprising.
-On the contrary, it expresses a belief which no reasonable person seems to doubt: The verification of a consequence renders a conjecture more credible.
-With a little attention, we can observe countless reasonings in everyday life, in the law courts, in science, etc., which appear to confirm to our pattern.'' (Polya, 1954, pp.
-4-5)
-%    \item Extraordinary claims require extraordinary evidence
-%\end{enumerate}
 
@@
 
-\appendix
-
-\section{code}
-add JASP links/code/etc.?
-
 \end{document}

### 2021-05-01 manuscript.tex
final version to submit

@@
 \documentclass[man]{apa7}
+% \documentclass[man, mask, floatsintext]{apa7}
 
@@
 \abstract{
-Uncertainty is ubiquitous in science and this uncertainty takes many forms, but the rhetoric around scientific knowledge and negotiating standards of evidence is often portrayed as immutable.
-Adding to this problem around the limited recognition of the tentative nature of science, when scientists try to communicate their uncertainty, research has shown us that many—including both learners and experts—struggle to interpret the statistics that are often used to make inferences from scientific data.
-We argue that a Bayesian approach can offer a different foundation and language around a core but sometimes unsettling tenet and persistent feature, uncertainty.
-Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains.
-In this essay, we (a) briefly introduce Bayes' theorem and wider ideas around Bayesian epistemology, (b) elaborate on the importance of probability and uncertainty in science and science education and review how people learn about these concepts and where they face difficulties, and (c) provide conceptual and practical examples within the JASP statistical software of how Bayes could help to address these issues.
-We conclude with implications and directions for future research focused around bolstering trust in science outside of classroom settings (for which Bayesian approaches could serve as a crosscutting part of science literacy) and in pre-collegiate science classrooms, where Bayes can provide a coherent approach to engaging in numerous science practices such as arguing from evidence and analyzing and interpreting data in a principled and flexible way.
+Uncertainty is ubiquitous in science, but scientific understanding is often represented to the public and in educational contexts as certain and immutable.
+Adding to this problem is that when scientists try to communicate their uncertainty, research has demonstrated that many people--learners and experts alike--struggle to interpret the statistics that are often used to make inferences about evidence from scientific data.
+We conjecture that a Bayesian perspective can offer a foundation for teaching and learning about a persistent but sometimes unsettling tension for science, making sense of and communicating about uncertainty.
+Central to this approach is viewing uncertainty \ital{probabilistically}, as doing so provides a language for expressing the degree of uncertainty in knowledge across a range of scientific domains and socio-scientific issues.
+In this paper, we elaborate on the importance of probability and uncertainty in science (and science education) and review how people learn about these concepts---and when and why they face difficulties when doing so.
+Then, we provide an accessible primer on Bayes Theorem and more general principles following from the application of Bayes Theorem.
+Next, we provide examples of how a Bayesian perspective could inform efforts around both science literacy and science education with an emphasis upon building trust in science.
+We conclude with directions for future research focused on bolstering trust in science outside of and within pre-collegiate science classrooms, where a Bayesian perspective can provide a coherent approach to engaging in science to make sense of scientific ideas in a principled yet flexible manner.
 }
 
-\keywords{APA style, demonstration}
+\keywords{Bayesian, uncertainty, probability, trustworthiness, statistics}
 
@@
 
-Uncertainty is ubiquitous in science and this uncertainty takes many forms.
-For example, measurement uncertainty is present when there are differences (small or large) in the repeated measures of the same object or phenomenon \parencite{pls03} or when the object that is measured, like a plant, is inherently variable in its characteristics \parencite{pls03}.
-Another form of uncertainty is its role within explanatory theories, such as those associated with evolution and quantum physics, for which central models take the form of probablistic statements, i.e., uncertainty is a core constituent of these models \parencite{g00}.
+Uncertainty is ubiquitous in science and creating ways to account for this uncertainty has helped science to progress.
+Consider, for instance, uncertainty due to the measurements or instruments used to study phenomena: In such cases, error due to measurement can be incorporated into models through the repeated measures of the same object or phenomenon \parencite{pls03}.
+Measurement is also present when the object that is measured--like a plant--is variable in its characteristics because of environmental and genetic variation (and interactions among these sources of variation) between the individual plants \parencite{pls03}.
+Another form of uncertainty is accounted for in terms of its central role within explanatory theories--such as those associated with evolution and quantum physics--for which models take the form of probabilistic statements about the world \parencite{g00}.
 
-Negotiating these many forms of uncertainty in argumentation—and building toward consensus at the same time that uncertainty is present—is a key part of the scientific process \parencite{t00}.
-In consequence, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions about the world  \parencite{g10, g06, n02, r77}.
+Negotiating the many forms of uncertainty through (constructively) arguing—-and building toward consensus at the same time that uncertainty is present—-is a key part of the scientific process \parencite{duschl2008science, t00, nrc12}.
+Concomitantly, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions \parencite{g10, g06, n02, r77}.
+However, while scientists and historians and philosophers of science are well aware of the uncertainty (and limitations) around scientific knowledge and the practice of negotiating and arguing from evidence, scientific knowledge is often represented to the public and in educational contexts as certain and immutable \parencite{d90}.
+Trust in science can erode and has eroded when scientists individually or collectively change their views, such as during the first phase of the COVID-19 crisis, when scientists faced serious criticism from the public, the media, and elected representatives for changing their recommendations based upon new scientific data and findings.
+A similar tension is now present concerning vaccines for COVID-19--a tension that involves reconciling robust but still initial clinical trial data and new data from population-level vaccination efforts and making policy-level recommendations (and individual decisions) amid this uncertainty.
+Adding to this difficult state of affairs regarding the role of uncertainty in science is that when scientists try to communicate their uncertainty, prior research has demonstrated that many people–learners and experts alike–struggle to interpret the statistics and derivations of statistics (such as standard errors/margins of error and confidence intervals) and that are often used to make inferences from scientific data \parencite{gkv04, s07, tk74}.
+Similarly, difficulties in understanding the technical language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution and nuclear decay that are modeled in terms of probabilistic statements \parencite{fth17}.
 
-However, while scientists and philosophers of science may be aware of the uncertainty (and limitations) surrounding scientific knowledge and negotiating standards of evidence interpretation, scientific knowledge is still often portrayed as immutable in educational contexts \parencite{d90}.
-In consequence, trust in science can erode when scientists individually or collectively change their views, such as during the first phase of the COVID-19 crisis in Germany, when scientists faced serious criticism from politicians and the media for changing their opinion based on new information.
-A similar tension was present in the United States and around the world regarding potential COVID-19 medication, when initial scientific data and changing interpretations of its utility collided with citizens’ skepticism (and, partisan political views).
+In this paper, we argue that these challenges concerning trust in scientific knowledge that not only can but must change can be addressed in a principled, flexible way, but that doing so requires making use of a Bayesian perspective that heretofore has been under-used outside of particular scientific and engineering domains---domains that often make little contact with science educators and researchers with an interest in building trust in science.
+Yet, the opportunity to do so now exists for technical and social reasons.
+Based on accumulating evidence from psychologists using Bayesian models of cognition \parencite{g12, tgk06}, statisticians and statistics educators advancing tools and practices for Bayesian methods \parencite{a02, b02, h_a20, h_j20,me14, s07} and science education scholars who have advanced a Bayesian perspective on argumentation \parencite{n11, s07}, we conjecture that a Bayesian perspective can help people--and science learners--to understand and reason about uncertainty in a way that leverages their existing ways of thinking.
+Central to this conceptualization is viewing uncertainty \ital{probabilistically}, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains \parencite{g12}.
+In drawing on research from psychologists, learning scientists, and educational researchers, we make this conjecture not only on the merits of Bayes' Theorem as a mathematical or statistical tool, but primarily on evidence that the perspective we are advocating can provide an intuitive way for people to think about data and evidence in the world \parencite{g12} and in classroom contexts \parencite{ll20, n11, s07, so12, w17}.
 
-Adding to this problem around the limited recognition of the tentative nature of science, when scientists try to communicate their uncertainty, research has shown us that many—including learners and experts alike—struggle to interpret statistics such as confidence intervals or error bars \parencite{gkv04, s07, tk74}.
-Similarly, difficulties in understanding the language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution of nuclear decay that are modelled in terms of probabilistic statements \parencite{fth17}.
+\section{Uncertainty in Science and the Central Role of Probability}
 
-Based on accumulating evidence from cognitive psychology, neuroscience, and education, we build on and extend beyond the science and engineering practice of argumentation the foundational work of science educational \parencite{n11, s07} and statistics education scholars \parencite{a02, b02, h_a20, h_j20,me14, s07} to argue for a \EJ{I'd say ``underused'', because Bayes itself is pretty old and now pretty hip} approach to uncertainty and probability that is currently underrepresented in science education (research).
-Specifically, we argue that it is helpful to consider uncertainty in terms of a Bayesian epistemological framework to help people to understand uncertainty and probability better and to engage with them in a way that leverages their existing ways of thinking.
-Central to this conceptualization is viewing uncertainty probabilistically, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains \parencite{g12}.
-In this way, we make this conjecture not only on the merits of Bayes' Theorem as a mathematical or statistical tool, but based on a body of psychological and educational research that shows that such an approach is an intuitive way for people to think about data and evidence in the world \parencite{g12} and in classroom contexts \parencite{ll20, n11, s07, so12, w17}.
+Probability and uncertainty are ubiquitous in scientific methodologies, scientific concepts, and in how science is communicated \textcite{gsoobmc17}.
+The practice of science often begins with measurement for which measurement error--which has a random component--is endemic.
+This random component arises from random variations in the measurement process and limits the certainty that one can have in the measurements.
+Reducing that measurement error has been critical for many discoveries in science, such as in sorting out the periodic table as we know it today \parencite{fco15}.
+However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., the recent study on the relationship between SARS-CoV-2 viral load and patient age by \textcite{jmvbzhd20} was heavily debated in the scientific community \parencite{frick_peer-review_2020}.
 
-The remainder of this paper is structured as follows.
-First, we will introduce Bayes theorem and wider ideas around Bayesian epistemology.
-Next, we will elaborate on the importance of probability and uncertainty in science (education) and review how people learn about these concepts and where they face difficulties.
-Then, we will provide examples of how Bayes could help to address these issues before concluding with implications and directions for future research.
+Probability and uncertainty are not only key concepts in science practices (such as analyzing and interpreting data) but are also an integral part of explanatory models of and theories for scientific concepts.
+For example, in the context of Heisenberg’s uncertainty principle and--more generally--quantum mechanics \parencite{feynman1951operator}, confidence is limited not only by measurement error but in principle---even in the perfect experimental setup there is no deterministic outcome.
+Instead, one has to calculate the probability of each possible outcome using the Born rule.
+To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunneling \parencite{c19}.
+Another example of how probability and uncertainty are central to learning scientific concepts is evolution \parencite{fsnh19}.
+Evolution is modeled as a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically.
+Further, the continued survival of these variations is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modeled probabilistically.
+Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers \parencite{m16}.
+Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public \parencite[]{gkv04, c14, mg17}.
+For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign a higher probability to the center of the interval \parencite{kl18} or \emph{p}-values are interpreted as effect sizes \parencite{gc17, n14} (although p-values describe how (in)compatible a set of data is with a set of assumptions.
+In sum, misinterpretations of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work.
 
-\section{Primer on Bayes Theorem and Bayesian Views of Epistemology}
+\section{Uncertainty in Science Education}
 
-In essence, Bayes' theorem provides a framework for updating one's beliefs about the world from observations in a way that takes prior knowledge into account in a principled manner.
-Bayes' theorem (Equation 1) states that at any point in time, our knowledge about the world after making a new observation (i.e., our posterior knowledge) can be obtained by multiplying our prior knowledge with a predictive updating factor (e.g., \cite{RouderMorey2019,WagenmakersEtAl2016CD} ):
+\subsection{The Roles of Probability Across Scientific Disciplines}
+
+Given the above-noted roles of probability in scientific concepts and theories, it is unsurprising that educational researchers have documented how students learn about probability.
+As one example, probability and chance events play an important role in the life sciences \parencite{g03, gk08}, particularly in evolutionary processes and learning about them \parencite{th17}.
+Recently, \textcite{fsnh19} investigated the relation between statistical reasoning and acceptance and knowledge about evolution in a large sample of nearly 500 university students in the US.
+They found that students’ statistical reasoning capabilities are a strong predictor of both acceptance of evolution and knowledge about evolution.
+
+In the physical sciences, students' struggles with the probabilistic nature of quantum physics \parencite{br02} and nuclear decay \parencite{smmv17} have been established, but the underlying reasons and mechanisms remain less researched than in evolution.
+Research into students’ misconceptions inspired by conceptual change research has started to address this gap \parencite{ms15, alra14, st09}.
+Findings mirror the position of \textcite{bgv94}.
+Specifically, the concepts of probability are taught abstractly and in such a way that does not build upon the ideas that students already hold, the experiences that students have, and the language that students use.
+Thus, these studies affirm students' struggles with scientific concepts that are probabilistic but hint at an underlying issue of more foundational challenges with reasoning about probability beyond science education contexts.
+
+Indeed, past research has shown that the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners.
+The work of \textcite{tk74}, especially, points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit.
+A prominent example is the neglect of base rate information in the (timely) context of medical testing, that is, the probability of having a medical condition based on a positive test result is often hugely overestimated because information about the base rate of people affected from the condition is neglected \parencite{kahneman_thinking_2012}.
+
+Responding to and building upon Kahneman and Tversky's work, Gigerenzer and colleagues provided important insights into the underlying mechanisms for the occurrence of these biases.
+Their research suggests that the biases are at least partly the result of information about probability and uncertainty being provided in a way that is incompatible with the heuristics that people develop from their everyday experiences \parencite{gh95, jkg18}.
+Based upon this premise, \textcite{jkg18} were able to demonstrate that students judge probability significantly better when language is adopted that better aligns with the heuristics that most learners develop from their everyday experiences.
+In sum, it seems feasible to change representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that these support, rather than hinder, learning.
+
+\subsection{Uncertainty in the Context of Analyzing and Interpreting Data}
+
+Recent work--primarily by developmental psychologists and scientists--has applied Bayesian ideas to a new context: how children make sense of and world \parencite{g12, tgk06, tgk06}.
+This notion that children reason about the world in a way that is analogous to Bayes Theorem is in some ways surprising, but it also builds upon much earlier work that also documented the surprisingly complex ways that children could reason about the world---as well as the assumed limitations of children's reasoning, the work of \textcite{pi69}.
+Different from earlier work on the stages through which children develop, recent work using Bayesian models considers the differences in the reasoning ability of children and adults to be a matter of degree, rather than kind.
+In short, children hold initial ideas and understandings that they change (or update) in ways that are concordant with how Bayes Theorem represents the updating of an initial belief in light of data (and how the Bayesian Learning Cycle described earlier allows individuals to integrate inductive and deductive reasoning).
+
+In addition to developmental research that uses Bayes Theorem as an analog for how children make sense of data and information in light of their initial ideas and beliefs, other work bears a striking semblance to developmental research that does.
+In particular, research on how learners’ prior understanding influence how they make sense of how likely different patterns in data are centers the same tension around interpreting data in light of initial beliefs that a Bayesian perspective highlights \parencite[e.g., ][]{kd88, mkm07, ssc07}.
+While these studies have focused on how students’ prior ideas and data interact, Bayesian frameworks were not used to formalize this process of integrating data.
+
+A Bayesian view of how children learn from their experiences and update their beliefs about the world has become especially prominent over the past 10-15 years \parencite{gt07, gw12}.
+Because of this, one might assume that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education and developmental scientists have pointed to this opportunity.
+For instance, \textcite{g12} wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p.
+1627).
+
+However, despite arguments such as Gopnik's (2012) for integrating developmental and educational research, it has largely not been the case that “science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” \textcite[p.
+1627]{g12}, as some science education scholars have bemoaned \parencite{ls15}).
+Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas \parencite[e.g., ][]{pi69} that represent the greatest contribution of developmental ideas to education \textcite{ls07}.
+We argue that this is still largely the case in the present, despite advances in both developmental research as well as statistical methodology related to Bayesian ideas—and that it is time to start exploring how bringing out the inherent Bayesian ways of thinking in education may provide support for learners’ efforts to make sense of the world.
+
+At the same time that opportunities to connect developmental science with education (especially science education), science education research has emphasized how important students' prior knowledge and understandings are to their understanding of scientific concepts and processes \parencite[e.g., ][]{ssc07}.
+For example, \textcite{mkm07} summarized findings across several studies showing that when elementary grade students reasoned with data in a context about which they had strong and valid prior understandings that informed their interpretation of data (about force and motion in the context of ramps), and they were able to explain sources of error in the data \parencite{mk03}, but when the data was in a context they did not understand well, they struggled to distinguish between random variation due to sources of error and sources of variation due to something that was causing a difference in the data (generated by measuring the motion of pendulums).
+Later design- and laboratory-based research has shown that students can develop an understanding of scientific ideas through the analysis of data \parencite{ls04,mkk17}.
+
+In summary, developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in education--especially science education--to design instruction based around eliciting and understanding students’ ideas \parencite{gb16, hbavb20, wtbs12}.
+There is an opportunity to use Bayesian ideas as a framework for formalizing these ideas about how prior ideas affect students’ reasoning about data and how teachers can leverage students’ ideas to design instruction, but a Bayesian perspective has--with a few exceptions focusing on how Bayes Theorem can provide a model for the science practice of argumentation \parencite[e.g., ][]{n11, so12}--has been absent from research in science education.
+
+After having reviewed how people learn about probability and uncertainty and where they face difficulties and research on uncertainty in the context of analyzing data, we next formally introduce Bayes Theorem and describe the broader process through which one's beliefs can be revised, or updated, in a way that leverages both inductive reasoning from data and deductive reasoning from beliefs or principles--what we refer to as the \ital{Bayesian Learning Cycle}.
+We also describe some of the ways of knowing, or epistemological considerations, that follow from Bayes Theorem and which comprise a core part of the Bayesian perspective that we argue has a role in building trust in science.
+
+\section{A Primer on Bayes Theorem and a Bayesian Epistemology}
+
+\subsection{An Accessible Introduction to Bayes Theorem}
+
+At its core, Bayes' theorem provides a means to update one's beliefs about the world from observations in such a way that prior knowledge can be used in a principled manner.
+As represented in Equation 1, Bayes' Theorem states that at any point in time, our knowledge about the world after making a new observation (i.e., our posterior knowledge) can be obtained by multiplying our prior knowledge with a predictive updating factor (e.g., \cite{RouderMorey2019,WagenmakersEtAl2016CD} ):
 \begin{equation*}
@@
 First, $p(\text{data} \given \theta)$ is the likelihood, that is, the extent to which the observed data were expected under $\theta$.
-Second, $p(\text{data})$ is the marginal likelihood, that is, the extent to which the observed data were expected without conditioning on $\theta$.
+Second, $p(\text{data}$ is the marginal likelihood, that is, the extent to which the observed data were expected without conditioning on $\theta$.
 If the observed data are more expected (i.e., better predicted) when conditioning on $\theta$, the predictive updating factor is higher than 1, and $\theta$ becomes more credible.
 
-To make this more tangible, consider the following example.
-Suppose an educator is interested in the proportion of students that use Windows or MacOS as their operating system.
+To illustrate the above equation, consider the following: An educator is interested in the proportion of students that use Windows or MacOS as their operating system.
 The proportion of Windows users $\theta$ ranges between 0 and 1.
 Before you start our Bayesian analysis, you need to quantify your background knowledge -- the relative plausibility of the different values for $\theta$.
-Given their previous experiences as a teacher, their best guess is that about half of the students use MacOS.
+Given their previous experiences as a teacher, their best guess is that about half of the students use Windows.
 They would not be surprised if the proportion of Windows users was between 30\% and 70\% but you would be rather surprised to find either no Windows or exclusive Windows usage.
 
-This background knowledge motivates the use of the dome-shaped prior distribution shown as the dashed line in Figure~\ref{fig:WindowsMacPriorPosterior}.\footnote{This is the beta$(2,2)$ distribution.} Now suppose you start to collect data by asking students about their operating systems.
-The first ten observations of Windows (W) and MacOS (M) usage that you make are $\{W, W, M, W, M, W, W, M, W, M\}$: six students use Windows, four use MacOS.
+This background knowledge motivates the use of the dome-shaped prior distribution shown as the dashed line in Figure~\ref{fig:WindowsMacPriorPosterior}.\footnote{This is the beta$(2,2)$ distribution.} Now suppose you start to collect data by asking students about the operating systems of their computers.
+The first ten observations of Windows (W) and MacOS (M) usage that you make are $\{W, W, M, W, M, W, W, M, W, M\}$: six students use Windows and four use MacOS.
 
@@
 
-Although we have demonstrated this updating process in a single step, it could also be executed sequentially, one student at a time.
+Although we demonstrated this updating process in a single step, it could also be executed sequentially, one student at a time.
 For instance, the first observation you make is a $W$ and in light of your prior this makes Windows usage a little more likely than MacOS usage.
@@
 
-How do the principles from this example translate to students' work with data and evidence -- not only educators?
-A recent case from physics can serve to illustrate this approach in the context of students' activity in science classrooms.
+\subsection{Integrating Inductive and Deductive Reasoning With the Bayesian Learning Cycle}
+
+How do the principles from the above example translate to scientific data?
+A recent case from the physical sciences can serve to illustrate this approach.
 In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light \parencite{b11} - and thus defying Einstein’s theory of relativity.
 However, instead of celebrating a historic discovery, the scientists started to look for what had gone wrong in their observations.
-
-How can Bayes’ theorem help to make sense of this?
-With nearly a century of accumulated evidence that the theory of relativity holds, scientists had a very strong prior belief about the speed of light.
+With nearly a century of accumulated evidence that the theory of relativity holds, from a Bayesian perspective, scientists had a very strong prior belief about the speed of light.
 To change this belief, extraordinary evidence would have been required.
-Thus, a single instance of apparently faster-than light elementary particles does not give a likelihood strong enough to change the prior in a meaningful way and in consequence the posterior, the knowledge about the world after making an observation, does not change in a meaningful way.
-
+Thus, a single instance of apparently faster-than-light elementary particles does not give a likelihood strong enough to change the prior in a meaningful way, and in consequence, the posterior--the knowledge about the world after making an observation--does not change in a meaningful way.
 Thus, rather than changing their belief about the speed of light, scientists looked for things that could have gone wrong with the measurement.
-This example shows how a qualitative take on Bayes’ theorem can explain and predict many aspects of how animals \parencite{o13}, people \parencite{tgk06}, and organisations \parencite{kaj12} learn from observations.
+This example shows how a qualitative take on Bayes’ theorem can explain and predict many aspects of how animals \parencite{o13}, people \parencite{tgk06}, and organizations \parencite{kaj12} learn from observations.
 Recently, advances in computational power and techniques have started the more and more widespread application of this principle in a quantitative way that allows researchers to precisely express their knowledge about the world and then learn from observations.
-
 Expressed in the Bayesian learning cycle (Figure \ref{fig:The Bayesian Learning Cycle}), a Bayesian reasoning framework allows scientists to integrate the strengths of both, inductive and deductive research approaches, in a formalized way.
-Further, when specifying the prior, scientists openly and explicitly need to describe the (un)certainty in the current state of knowledge and the posterior captures how the data has changed the knowledge and the (un)certainty in that knowledge.
+Further, when specifying the prior, scientists openly and explicitly need to describe the uncertainty in the current state of knowledge, and the posterior captures how the data has changed the knowledge and the uncertainty in that knowledge.
 Thus, Bayesian reasoning is, at least in principle, open and transparent by design.
-
 \begin{figure*}[h]
@@
 \includegraphics[width = .65\paperwidth]{learning.png}
-\caption{The Bayesian Learning cycle}
+\caption{The Bayesian learning cycle is a reasoning framework that allows scientists, citizens, and learners to integrate the strengths of inductive and deductive reasoning while making sense of data in light of prior beliefs.}
 \label{fig:The Bayesian Learning Cycle}
@@
 \end{figure*}
+Having introduced Bayes Theorem and an application of Bayes Theorem as a cycle through which individuals can learn, we show how Bayes Theorem can be used to demonstrate several principles that are relevant to building trust in science.
 
-Having introduced Bayes theorem and an application of Bayes theorem as a cycle through which individuals can learn, we describe the role of uncertainty in science, with a focus on the role of probability.
+\section{The Epistemological Principles that Follow from Bayes Theorem}
 
-\section{Uncertainty in Science and the Role of Probability}
-
-Probability and uncertainty are ubiquitous in science—in scientific methodologies, the scientific concepts, or in how science is communicated \textcite{gsoobmc17}.
-The practice of science often begins with measurement, and with measurement comes measurement error, which has a random component.
-This random component arises from random variations in the measurement process and limits the certainty that one can have in the measurements.
-Reducing that measurement error has been critical for many discoveries in science, such as in sorting out the periodic table as we know it today \textcite{fco15}.
-However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., the recent study on the relationship between SARS-CoV-2 viral load and patient age by \textcite{jmvbzhd20} was heavily debated in the scientific community \parencite{frick_peer-review_2020}.
-
-Probability and uncertainty are not only key concepts in science practices such as analyzing and interpreting data, but also part of the scientific concepts themselves.
-One example is Heisenberg’s uncertainty principle or quantum mechanics more generally \textcite{feynman1951operator}.
-In quantum mechanics, confidence in measurement is limited not only by measurement error but in principle—even in the perfect experimental setup there is no deterministic outcome.
-Instead, one has to calculate the probability of each possible outcome using the Born rule.
-To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunnelling  \textcite{c19}.
-Another example of how probability and uncertainty are central to learning scientific concepts is  evolution  \textcite{fsnh19}.
-Evolution is modelled as a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically.
-Further, the continued survival of these variations is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modelled probabilistically.\EJ{I agree that we can only describe the process probabilistically.
-But that does not mean it \emph{is} probabilistic.} \MK{tried to address this comment as well.} Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers  \textcite{m16}.
-Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public \parencite[]{gkv04, c14, mg17}.
-For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign higher probability to the center of the interval \parencite{kl18} or \emph{p}-values are interpreted as effect sizes \parencite{gc17, n14} (although p-values describe how (in)compatible a set of data is with a set of assumptions.
-In sum, misinterpretations of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work.
-
-\section{Uncertainty in Science Education}
-
-Given the above-noted roles of probability in scientific concepts and theories, it is unsurprising that educational researchers have documented how students learn about probability.
-As one example, probability and chance events play an important role in the life sciences \parencite{g03, gk08}, particularly in evolutionary processes and learning about them \parencite{th17}.
-Recently, \textcite{fsnh19} investigated the relation between statistical reasoning and acceptance and knowledge about evolution in a large sample of nearly 500 university students in the US.
-They found that students’ statistical reasoning capabilities are a strong predictor of both acceptance of evolution and knowledge about evolution.
-Thus, this study suggests (but does not directly show) that instructional strategies for supporting students’ statistical reasoning could improve their learning about specific evolutionary concepts.
-
-In the physical sciences, students' struggles with the probabilistic nature of quantum physics \parencite{br02} and nuclear decay \parencite{smmv17} are known, but the underlying reasons and mechanisms remain less researched than in evolution.
-Research into students’ misconceptions inspired by conceptual change research has started to address this gap \parencite{ms15, alra14, st09}.
-Findings seem to mirror the position of \textcite{bgv94} in that the concepts of probability are taught too abstractly and in a way that does not build upon and connect with the ideas that students’ already hold, the experiences that students make, and the language that students use.
-Thus, these studies affirm students' struggles with scientific concepts that are probabilistic but hint at the underlying issue being foundational, conceptual struggles with ideas about probability not limited to science learning.
-Indeed, the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners.
-The work of \textcite{tk74} points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit.
-A prominent example is the neglect of base rate information in the (timely) context of medical testing, that is, the probability of having a medical condition on the basis of a positive test result is often hugely overestimated because information about the base rate of people affected from the condition is neglected \parencite{kahneman_thinking_2012}.
-
-More recent work by Gigerenzer and colleagues has provided important insights into the underlying mechanisms for the occurrence of these biases \parencite{gh95, jkg18}.
-Their research suggests that the biases are at least partly the result of information about probability and uncertainty being provided in a way that is incompatible with the heuristics that people develop from their everyday experiences.
-Based on this assumption, \textcite{jkg18} were able to demonstrate that students judge probability significantly better, when language is adopted that better aligns with the heuristics that most learners develop from their everyday experiences.
-In sum, it seems feasible to change representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that these support instead of inhibit learning.
-
-\section{How Learners and People Make Sense of Data in Light of Uncertainty}
-
-Recent work in cognitive and psychological science on how children learn argues that children understand and learn about the world through the intuitive application of Bayesian ideas \parencite{g12, tgk06, tgk06}.
-This idea has led current research to conceptualize human learning in terms of Bayesian probabilistic models of learning.
-In other words, people (and children) may view (or come to view) the world probabilistically, and in ways that happen to align with Bayesian interpretations of probability.
-
-This view of how children learn from data has become especially prominent in developmental science and is successfully applied there \parencite{gt07, gw12}.
-One might assume that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education.
-Developmentalists have pointed to the potential and opportunity here.
-\textcite{g12} wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p.
-1627).
-
-Despite arguments for integrating developmental and educational research, it has largely not been the case that—with respect to advances in Bayesian methods—“science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” \textcite[p.
-1627]{g12}, as some science education scholars have bemoaned \parencite{ls15}).
-Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas \textcite{pi69} that represent the greatest contribution of developmental ideas to education \textcite{ls07}.
-We argue that this is still largely the case in the present, despite advances in both developmental research as well as statistical methodology related to Bayesian ideas—and that it is time to start exploring how bringing out the inherent Bayesian ways of thinking in education may provide support for learners’ efforts to make sense of the world.
-
-Related science education research emphasizes the role of prior knowledge that students bring with them to the analysis of data and how that prior knowledge impacts the inferences they make about data.
-For example, \textcite{mkm07} summarized findings across a number of studies showing that when elementary grade students reasoned with data in a context about which they had strong and valid prior understandings that informed their interpretation of data (about force and motion in the context of ramps), and they were able to explain sources of error in the data \parencite{mk03}, but when the data was in a context they did not understand well, they struggled to distinguish between random variation due to sources of error and sources of variation due to something that was causing a difference in the data (generated by measuring the motion of pendulums).
-Later design- and laboratory-based research has shown that students can develop an understanding of scientific ideas through the analysis of data \parencite{ls04,mkk17}.
-
-Laboratory-based research, especially, has examined how learners’ prior understanding may have a bearing on how they make sense of how likely different outcomes are \parencite{kd88, mkm07, ssc07}.
-While these studies focused on how students’ prior ideas and data interact, Bayesian frameworks were not used to formalize this process of integrating data in light of prior beliefs about scientific phenomena.
-In short, this work aligns closely with the Bayesian models of how children learn from data described by \textcite{g12}, but this research does not explicitly interpret their findings in light of these ideas about how children learn from data in a Bayesian way.
-
-Overall, developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in science education to design instruction based around eliciting and understanding students’ ideas \parencite{gb16, hbavb20, wtbs12}.
-There is an opportunity to use Bayesian ideas as a framework for formalizing these ideas about how prior ideas affect students’ reasoning about data and how teachers can leverage students’ ideas to design instruction, yet Bayes has mostly been absent from research in science education.
-After having reviewed how people learn about probability and uncertainty and where they face difficulties, we want to present how Bayes could be part of a solution.
-First, we will make the case how Bayes could support science literacy at large before delving into more concrete topics.
-
-\section{Bayes and Science Literacy}
-
-\subsection{Overview of Bayes and Science Literacy}
-
-Having described the role of probability and uncertainty in science and science education, we now describe what these Bayesian ideas can mean for science literacy.
-We see two primary ways in which Bayesian methods can enhance science literacy, or a range of capabilities that people may possess which have value (individually or collectively) to them \parencite{council2016science}.
-These capabilities range from foundational literacies, including numeracy and textual literacy (what is most thought of concerning literacy in common parlance), content knowledge, and epistemic knowledge and other ways of thinking and habits of mind \parencite{council2016science}.
-One way that Bayesian ideas can inform discussions of science literacy is to provide individuals with a technique for combining their prior understanding with new evidence.
-Along this line, Bayes and the Bayesian learning cycle more specifically can provide a conceptual tool for people to combine their understanding with new evidence (and data) in a rigorous and coherent way, as we discuss next.
-
-The history of science is, in many ways, a story of established ideas being built upon or supplanted \parencite{fara2010science}, and the contemporary emphasis on viewing science as a practice was sparked by \textcite{k62} historical-epistemological account on what this suggests about the nature and philosophy of science \parencite{g09}.
-Yet, our ways of interpreting scientific evidence do rarely recognize this, perhaps for good reason: Most of the time, new data or information do not contradict foundational scientific ideas \parencite{k62}.
-The same phenomenon, though, is present on a more micro-level: While scientists may be familiar with the data they collect not eliding with ideas that are a part of the scientific consensus, in media and the public discourse scientific ideas are often presented as fixed and certain—even in cases for which consensus has not been reached \parencite{c87}
-
-These historical and philosophical ideas align with a Bayesian perspective on science literacy.
-Established scientific ideas can be considered to be those that individual people - psychologically - should be more confident about, whereas newer ideas, or data suggesting new ideas, can be weighted less.
-In this way, Bayes can provide a simple heuristic for people to evaluate new information or evidence; by asking the question of how informative is the evidence in light of the prior.
-Thus, individuals possess a tool, one that is non-dichotomous, in that the evidence is not true or false, and also probabilistic, in that what individuals can think, going forward, is about the extent to which a particular belief has merit and is informative.
-
-A concrete example can help to illustrate how a Bayesian perspective on a topic relevant to science literacy may offer a different way forward.
-Consider a salient topic, medical vaccines for diseases, such as Polio and the disease caused by the coronavirus SARS-CoV-2, COVID-19.
-Most evidence suggests that vaccines have been safe, historically, and are generally safe, at present (Centers for Disease Control and Prevention, 2021).
-Nevertheless, they are not perfectly safe, as unlikely, adverse outcomes do happen, and it is possible that there are adverse side effects from some vaccines that research has not yet revealed.
-In light of this evidence, consider the development of a new vaccine, such as those available at the time of this writing for COVID-19.
-Some use long-established methods of vaccine development (namely, deactivating SARS-CoV-2 particles but retaining cell signalling proteins to which humans develop an immune response—and, after, a degree of immunity) and injecting these deactivated virus particles into people), whereas others use a new mechanism (namely, injecting messenger RNA into people; this messenger RNA is then transcribed by cells in people’s bodies to produce a particular protein common to SARS-CoV-2 particles, which human’s bodies than learn to recognize to develop immunity.
-
-How should we consider evidence about the vaccines?
-Prior to any data from medical clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and on the basis of similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are.
-After initial trials were completed, the degree of safety that individuals associate with these vaccines—in light of the prior and the limited evidence—could increase, but still not be near the level of safety that many would consider acceptable.
-Following the completion of clinical trials that show minimal side effects (holding aside their efficacy), one’s estimates of the safety of these vaccines could increase by a substantial amount, but concerns could still remain, particularly for messenger RNA vaccines, which are newer, and may, plausibly, have unintended side effects.
-
-\subsection{Strategies for Supporting Science Literacy}
-
-What is the result of this entire process of weighting new evidence in light of a generic prior that we discussed above?
-We argue that there are three ways that Bayesian methods could support science literacy.
-
-The famous French astronomer and mathematician Laplace argued that Bayes' rule quantifies how rational observers ought to change their opinion in light of new evidence; in this sense, Laplace believed that Bayes' rule was \emph{common sense expressed in numbers}.
-We now showcase several insights that follow directly from Bayes' rule.
-In our earlier exposition of Bayes' rule we already emphasized the idea that hypotheses gain credibility when they predict the data well, and lose credibility when they predict new data poorly \parencite{WagenmakersEtAl2016CD}.
-This idea generalizes what the mathematician Polya termed the ``Fundamental Inductive Pattern'':
+In the previous section, we emphasized the idea that hypotheses, or ideas, gain credibility when they predict the data well, and lose credibility when they predict new data poorly \parencite{WagenmakersEtAl2016CD}.
+This idea generalizes to the ``Fundamental Inductive Pattern'':
 \begin{quotation}
@@
 
-The rule from Equation~\ref{eq:BayesRule} includes the constant term $p(\text{data})$, which does not involve $\theta$. Hence, we can also write Bayes' rule as:
+The rule from Equation~\ref{eq:BayesRule} includes the constant term $p(\text{data})$, which does not involve $\theta$. As a result, Bayes' rule can be re-written as:
 \begin{equation}
@@
 where `$\propto$' stands for `is proportional to'.
-Thus, Bayes' rule states that our posterior knowledge $p(\theta \given \text{data})$ is proportional to the likelihood $p(\text{data} \given \theta)$ (i.e., predictive adequacy, unsurprise, or the extent to which the observed data are expected given $\theta$) multiplied with our prior knowledge $p(\theta)$.
+Thus, Bayes' rule states that our posterior knowledge $p(\theta \given \text{data})$ is proportional to the likelihood $p(\text{data} \given \theta)$ (or the extent to which the observed data are expected given $\theta$) multiplied by our prior knowledge $p(\theta)$.
 
-Equation~\ref{eq:BayesRulePropto} emphasizes the fact that the posterior uncertainty about $\theta$ is a \emph{compromise} between our prior uncertainty about $\theta$ and the predictive performance of $\theta$.
-But the posterior uncertainty after having observing the first datum becomes the prior uncertainty for the to-be-observed second datum (cf.
+Equation~\ref{eq:BayesRulePropto} emphasizes that the posterior uncertainty about $\theta$ is a \emph{compromise} between our prior uncertainty about $\theta$ and the predictive performance of $\theta$.
+But, the posterior uncertainty after having observed the first datum becomes the prior uncertainty for the next datum---and any other data (cf.
 Figure~\ref{fig:WindowsMacSequential}).
-Consequently, after having observed the second datum, the posterior uncertainty represents a compromise of a compromise.
+Consequently, after having observed another datum, the posterior uncertainty represents a compromise of a compromise.
 As the data accumulate, the posterior uncertainty is more and more determined by predictive performance, and the impact of the initial uncertainty about $\theta$ is increasingly watered down: `the data overwhelm the prior' \parencite{WrinchJeffreys1919}.
 
-This implies that a constant stream of accumulating data ought to bring any two rational people into an arbitrarily close agreement, no matter how divergent their opinions may have been at the outset.
+This implies (but does not dictate, as we discuss next) that a constant stream of accumulating data ought to bring any two people into an arbitrarily close agreement, no matter how divergent their opinions may have been at the outset.
 There are two important caveats here.
-First and foremost, the data only overwhelm the prior if that prior is neither 0 (denoting an impossibility) nor 1 (denoting absolute certainty).
-If you already know for certain that a hypothesis is true or false, then you cannot adjust your opinion in light of the data: your initial opinion will be your final opinion, no matter what the data may indicate.
+
+First and foremost, the data only overwhelm the prior if that prior is neither equal to zero (denoting an impossibility) nor one (denoting absolute certainty).
+If someone already knows for certain that a hypothesis is true or false, then you cannot adjust your opinion in light of the data: your initial opinion will be your final opinion, no matter what the data may indicate.
 Philosophically, adopting such priors is a dangerous practice; for instance, in antiquity the adage of the \emph{New Academy} --a school of skeptics headed by Carneades-- was ``never assert absolutely''.
-In modern times, Lindley popularized this idea in statistics and coined it ``Cromwell's rule'':
-\begin{quotation}
-``Consequently, an uncertain event of zero probability remains so whatever information is provided.
-In other words, if a decision-maker thinks something cannot be true and interprets this to mean it has zero probability, he will never be influenced by \emph{any} data, which is surely absurd.
-So leave a little probability for the moon being made of green cheese; it can be as small as 1 in a million, but have it there since otherwise an army of astronauts returning with samples of the said cheese will leave you unmoved.
-A probability of one is equally dangerous because then the probability of $\bar{E}$ will be zero.
-So never believe in anything absolutely, leave some room for doubt: as Oliver Cromwell told the Church of Scotland, `I beseech you, in the bowels of Christ, think it possible you may be mistaken'.'' \parencite[p.
-104]{Lindley1985}
-\end{quotation}
+In modern times, Lindley popularized this idea in statistics and coined it ``Cromwell's rule'' \parencite[p.
+104]{Lindley1985}.
 
@@
 
-Now consider two specific values for $\theta$, $\theta_1$ and $\theta_2$. We can use Bayes' rule to obtain the posterior probability for each one:
+Next, consider two specific values for $\theta$, $\theta_1$ and $\theta_2$. We can use Bayes' rule to obtain the posterior probability for each one:
 \begin{align}
@@
 \end{align}
-When we divide the posterior probabilities the common term $p(\text{data})$ drop out and we have:
+When we divide the posterior probabilities the common term, $p(\text{data})$, drop out, and we have:
 \begin{equation}
@@
 \end{equation}
-This way of writing Bayes' rule highlights that extraordinary claims require extraordinary evidence -- if a particular hypothesis $\mathcal{H}_1$ is extremely unlikely a priori, the prior odds $\frac{p(\mathcal{H}_1)}{p(\mathcal{H}_0)}$ are stacked against it, and in order for the posterior odds to favor $\mathcal{H}_1$ the support from the data (i.e, the degree to which $\mathcal{H}_1$ outpredicts $\mathcal{H}_0$) needs to be overwhelmingly strong.
+This way of writing Bayes' rule highlights that extraordinary claims require extraordinary evidence -- if a particular hypothesis $\mathcal{H}_1$ is extremely unlikely a priori, the prior odds $\frac{p(\mathcal{H}_1)}{p(\mathcal{H}_0)}$ are stacked against it, and for the posterior odds to favor $\mathcal{H}_1$ the support from the data (i.e, the degree to which $\mathcal{H}_1$ predicts $\mathcal{H}_0$) needs to be overwhelmingly strong.
 
 Bayes' rule can also be shown to embody the \emph{principle of parsimony}: the rule implicitly contains a preference for the simplest model that explains the data well.
-To see this, consider a coin taken from a magician's coffin, and let parameter $\xi$ indicate the chance that the coin lands heads on any throw.
-The null hypothesis holds that the coin is fair, $\mathcal{H}_0: \xi = \nicefrac{1}{2}$.
-The alternative hypothesis that we entertain here specifies that the coin may be double-tails, fair, or double-heads, $\mathcal{H}_1: \xi \in \{0, \nicefrac{1}{2}, 1\}$, with each of the three options deemed equally likely a priori.
+To see this, consider a coin with two sides and let parameter $\xi$ indicate the chance that the coin lands heads on any throw.
+The null hypothesis holds that the coin is fair, with heads and tails equally likely: $\mathcal{H}_0: \xi = \nicefrac{1}{2}$.
+The alternative hypothesis that we entertain here specifies that the coin may be any of the following options with equivalent likelihood: double-tails, fair, or double-heads, $\mathcal{H}_1: \xi \in \{0, \nicefrac{1}{2}, 1\}$.
 
-The coin is tossed and we observe heads.
+The coin is tossed and heads is observed.
 The probability of this datum is \nicefrac{1}{2} under $\mathcal{H}_0$; under $\mathcal{H}_1$, it is $p(\text{heads} \given \xi=0) \cdot p(\xi=0 \given \mathcal{H}_1) + p(\text{heads} \given \xi=\nicefrac{1}{2}) \cdot p(\xi=\nicefrac{1}{2} \given \mathcal{H}_1) + p(\text{heads} \given \xi=1) \cdot p(\xi=1 \given \mathcal{H}_1) = 0 \cdot \nicefrac{1}{3}  + \nicefrac{1}{2} \cdot \nicefrac{1}{3} + 1 \cdot \nicefrac{1}{3} = \nicefrac{1}{2}$.
-So the datum is equally likely under $\mathcal{H}_0$ and $\mathcal{H}_1$: both models receive an equal amount of support.
+So, the datum is equally likely under $\mathcal{H}_0$ and $\mathcal{H}_1$: both models receive an equal amount of support.
 This means the first outcome does not change our conviction concerning $\mathcal{H}_0$ versus $\mathcal{H}_1$.
 The datum did, however, change our beliefs about $\xi$ under $\mathcal{H}_1$.
-Specifically, we now know that $\xi$ cannot be 0; moreover, the datum was twice as likely under $\xi=1$ as under $\xi = \nicefrac{1}{2}$, so that our posterior distribution for $\xi$ under $\mathcal{H}_1$ is now $p(\xi = \nicefrac{1}{2}) = \nicefrac{1}{3}, p(\xi = 1) = \nicefrac{2}{3}$.
+Specifically, we now know that $\xi$ cannot be zero; moreover, the datum was twice as likely under $\xi=1$ as under $\xi = \nicefrac{1}{2}$, so that our posterior distribution for $\xi$ under $\mathcal{H}_1$ is now $p(\xi = \nicefrac{1}{2}) = \nicefrac{1}{3}, p(\xi = 1) = \nicefrac{2}{3}$.
 
@@
 This means that the observed data provided $(\nicefrac{1}{4}) / (\nicefrac{1}{12}) = 3$ times more support for $\mathcal{H}_0$ than for $\mathcal{H}_1$.
-This happens because $\mathcal{H}_1$ spreads out its predictions, hedging its bets.
+This happens because $\mathcal{H}_1$ "spreads out" its predictions, hedging its bets.
 In contrast, the simple model $\mathcal{H}_0$ made a precise prediction.
 
-Instead of learning about the predictive performance of $\mathcal{H}_0$ and $\mathcal{H}_1$ one observation at a time, we could also have considered the probability that the models assign to the entire sequence $\{\text{heads}, \text{tails}\}$.
+Instead of learning about the predictive performance one observation at a time of $\mathcal{H}_0$ and $\mathcal{H}_1$, we could also have considered the probability that the models assign to the entire sequence $\{\text{heads}, \text{tails}\}$.
 Under $\mathcal{H}_0$ we again have $\nicefrac{1}{2} \cdot \nicefrac{1}{2} = \nicefrac{1}{4}$.
@@
 This is the penalty for complexity that $\mathcal{H}_1$ pays for entertaining three different values of $\xi$ from the outset.
-
-Thus, Bayes rule includes a automatic ``Ockham's razor'' \parencite{Jeffreys1939,JefferysBerger1992} in the sense that daring predictions are rewarded when they come true.
+Thus, in addition to reflecting the principle of parsimony and highlighting the utility of parsimonious explanations, Bayes rule includes a automatic ``Ockham's razor'' \parencite{Jeffreys1939,JefferysBerger1992} in the sense that daring predictions are rewarded when they come true.
 
@@
 This is in stark contrast to $p$-value statistical hypothesis testing, in which a statistical model (the null hypothesis) is judged in isolation.
+Having introduced Bayes Theorem and some principles that may follow from it, we next make the case how Bayes could support science literacy at large before delving into implications for science teaching and learning.
 
-Thus, in summary, the result of applying Bayes rule is a non-binary inference, one that doesn't speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty.
+\section{A Bayesian Perspective and Science Literacy}
+
+We conceive that Bayesian methods can enhance science literacy or a range of capabilities that people may possess which have value (individually or collectively) to them \parencite{council2016science}.
+These capabilities range from core \ital{literacies}, including numeracy and textual literacy (what is mostly thought of in common parlance concerning literacy), content knowledge, and epistemic knowledge \parencite{council2016science}.
+A potentially powerful way in which Bayesian ideas can inform discussions of science literacy is to provide individuals with a principle for combining their prior understanding with new evidence.
+Along this line, Bayes and the Bayesian Learning Cycle can provide a tool for people (and learners) to combine their understanding with new evidence (and data) rigorously and coherently.
+
+A timely example can help to illustrate how a Bayesian perspective on a topic relevant to science literacy may offer a different way forward.
+Consider a salient topic, medical vaccines for diseases, such as Polio and the disease caused by the coronavirus SARS-CoV-2, COVID-19.
+Most evidence suggests that vaccines have been safe, historically, and are generally safe, at present (Centers for Disease Control and Prevention, 2021).
+Nevertheless, they are not perfectly safe, as unlikely, adverse outcomes do happen, and there may be adverse side effects from some vaccines that research has not yet revealed.
+In light of this evidence, consider the development of a new vaccine, such as those available at the time of this writing for COVID-19.
+Some use long-established methods of vaccine development (namely, deactivating SARS-CoV-2 particles but retaining cell signaling proteins to which humans develop an immune response—and, after, a degree of immunity) and injecting these deactivated virus particles into people), whereas others use a new mechanism (namely, injecting messenger RNA into people; this messenger RNA is then transcribed by cells in people’s bodies to produce a particular protein common to SARS-CoV-2 particles, which human’s bodies then learn to recognize to develop immunity.
+
+How should we consider evidence about the vaccines?
+Before any data from medical clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and based on similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are.
+After initial trials were completed, the degree of safety that individuals associate with these vaccines—in light of the prior and the limited evidence—could increase, but still not be near the level of safety that many would consider acceptable.
+Following the completion of clinical trials that show minimal side effects (holding aside their efficacy), one’s estimates of the safety of these vaccines could increase by a substantial amount, but concerns could remain, particularly for messenger RNA vaccines, which are newer, and may, plausibly, have unintended side effects.
+
+While the complex weighting of the evidence above may seem natural to a scientist, it is challenging for many.
+In the following, we showcase how several important insights for weighting evidence follow directly from Bayes' rule.
+Thus, we argue that emphasizing Bayesian reasoning can support science literacy by providing people with a principled way to weighting evidence that is accessible because it builds on peoples' prior ideas.
+
+In the context of the above example on SARS-CoV-2, the result of applying Bayes Theorem in a heuristic mode is a non-binary inference, one that doesn't speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty.
 Though a small difference, discussing the safety of vaccines on these terms can provide a more solid foundation for principled disagreements and arguments.
-In addition, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in-line with how most of us likely already interpret data.
-Specifically Bayes provides people with a way to talk about weighing between ideas and their prior beliefs and new information and data.
-In short, Bayes gives people a chance to deal with uncertainty in a principled but accessible way.
+In addition, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in line with how most of us likely already interpret data \parencite{tgk06, gw12, gh95}; A Bayesian perspective can provide people with a way to talk about weighing between ideas and their prior beliefs and new information and data.
+In short, Bayes gives people a chance to deal with uncertainty in a principled but accessible manner.
 
-\section{Bayes and Science Education}
+In addition, many themes introduced in the section on the epistemological principles that follow from Bayes Theorem are relevant to the aim of supporting science literacy.
+First, \ital{even individuals with different initial beliefs may be brought closer to an agreement in light of data} that serves as evidence for a particular belief or hypothesis.
+In this way, Bayes Theorem illustrates--at the conceptual level as well as statistically--how individuals may be unlikely to agree based on the limited or weak information, but that this can shift in light of accumulating data.
+As we noted earlier, this is far from a deterministic process, and \ital{holding initial beliefs that a hypothesis has a probability of zero or one means that data does not affect one's beliefs}.
+This suggests that a cultural norm and principle for science literacy can be to always reserve a degree of uncertainty such that one could be influenced by (new) data or evidence: From a Bayesian perspective, \ital{our knowledge is uncertain and open to change and revision} so long as we remain open to new evidence.
 
-\subsection{Overview of Bayes and Science Education}
+Related to the importance of not holding scientific beliefs with a probability of zero or one, another principle that follows from Bayes Theorem is that \ital{extraordinary claims require extraordinary evidence}, and that unlikely or implausible hypotheses or claims are likely to necessitate strong empirical data to be convincing to others.
+A corollary of this principle is that we should strong initial or prior beliefs can serve as a tool through which we can discuss extraordinary claims: Except in light of strong evidence, "ordinary" claims can serve as the default for citizens and scientists making sense of unusual and potentially biased claims.
+Similarly, Bayes Theorem demonstrates that the \ital{simplest possible explanation--the most parsimonious explanation--is often the best explanation}.
 
-Research on the role of Bayes in science education has mostly been carried out in the context of undergraduate statistics education, in which there have been a number of calls to action \parencite{gpkpwc18, h_a20}, examples and design options \parencite{a02, b02,g08, w17, h_j} and debate \parencite{jrhrr20} over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs.
-The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power \parencite{gpkpwc18} as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers \parencite{ah20} .As evidenced by the recent special issue of the \emph{Journal of Statistics Education} of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.
+In sum, a Bayesian perspective offers a general approach to science literacy that can account for uncertainty in a rigorous and principled manner.
+In addition, a number of principles that follow from Bayes Theorem, such as being receptive to new evidence and weighting that evidence per its strength in light of the strength of one's initial beliefs or understanding could bolster science literacy---and how the public comes to understand and gain trust in science.
+
+\section{A Bayesian Perspective and Science Education}
+
+Research on the role of Bayes in science education has mostly been carried out in the context of undergraduate statistics education, in which there have been several calls to action \parencite{gpkpwc18, h_a20}, examples and design options \parencite{a02, b02,g08, w17, h_j20} and debate \parencite{jrhrr20} over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs.
+The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power \parencite{gpkpwc18} as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers \parencite{ah20}.
+As evidenced by the recent special issue of the \emph{Journal of Statistics Education} of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.
 
@@
 This paper represents the most comprehensive account of the relevance of Bayesian methods for science education.
-Szu and Osborne argue that Bayes is useful as both a formal mathematical tool, and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes Theorem that follow more or less deterministically from ``external, objectively probabilistic systems'' (p.
+Szu and Osborne argue that Bayes is useful as both a formal mathematical tool and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes Theorem that follow more or less deterministically from ``external, objectively probabilistic systems'' (p.
 61)—is ``the key leap that characterizes the debate about the value of Bayesian inference as a model of scientific reasoning'' (p.
 61).
-In this way, Szu and Osborne argue that the greatest use of Bayes theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community \parencite{batanero2016research, mr18}.
+In this way, Szu and Osborne argue that the greatest use of Bayes Theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community \parencite{batanero2016research, mr18}.
 They offer some research-related backing for the use of Bayes Theorem (e.g., noting how conceptual change research, particularly, and constructivist research, broadly, can both be explained coherently within a Bayesian approach) as well as some practical, instructional recommendations, some of which we detail later in this section.
@@
 Concerning the latter, Nussbaum describes how the social issue of raising taxes to provide resources to homeless individuals was a rich context for students to engage in forms of Bayesian reasoning.
-In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what \textcite{so12} characterize as the more externally objective use of Bayes Theorem—and how the estimates that result from applying Bayes theorem led students to re-evaluate their initial arguments.
+In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what \textcite{so12} characterize as the more externally objective use of Bayes Theorem—and how the estimates that result from applying Bayes Theorem led students to re-evaluate their initial arguments.
 
@@
 Curiously, no work has applied Bayesian methods to another science practice, that of analyzing and interpreting data; one for which Bayesian methods have clear conceptual applicability and utility, but one for which the absence of suitable tools may provide to be a barrier.
-In this way, Bayesian methods could serve as a cross-cutting concept \parencite{nrc12} that unites the practice of argumentation with analyzing and interpreting data, and using mathematics and computational thinking.
+In this way, Bayesian methods could serve as a cross-cutting concept \parencite{nrc12} that unites the practice of argumentation with analyzing and interpreting data and using mathematics and computational thinking.
 
-\subsection{Strategies for Supporting Science Education}
-
-Bayesian methods clearly have some utility in science education, even if that utility is under-realized at present.
+Bayesian methods have some utility in science education, even if that utility is under-realized at present.
 Next, we present some of the specific ways Bayesian methods could enhance science education—and could do so in a way that makes Bayes more tractable.
 
-First, we should consider ``unplugged'' versions of Bayesian analyses, akin to how computer science education scholars have emphasized that computational thinking does not necessarily require the use of a computer (Yadav et al., 2018).
-These could interface with the science and engineering practice of modeling; could students use diagrammatic models in conjunction with Bayesian “heuristics” to revise their models in a similar manner to how this practice can unfold in classrooms \parencite{schwarz2009developing}.
-
-Another advance should be to build out curricular strategies, elaborating on the ideas of \textcite{so12} and \textcite{n11}.
-At a high-level, Szu and Osborne write that in order to reason about data in light of their initial beliefs (and to construct a kind of informal likelihood statistic), students “need to see judgments about data and evidence being an assessment not only of the probability of the hypothesis being correct but also of it being wrong” (p.
+First, curricular resources and pedagogical approaches that take a Bayesian perspective can be developed.
+These could elaborate on the ideas of \textcite{so12} and \textcite{n11}.
+At a high-level, Szu and Osborne write that to reason about data in light of their initial beliefs (and to construct a kind of informal likelihood statistic), students “need to see judgments about data and evidence being an assessment not only of the probability of the hypothesis being correct but also of it being wrong” (p.
 67).
-Filling in more details, speaking from both the perspective of a researcher (of argumentation) and an educator (or, a designer of classroom activities), Nussbaum summarizes that using Bayes “forces one to (a) think about an issue from different perspectives, (b) reflect on one’s range of certainty and uncertainty, and (c) conduct a simple com- puter simulation (e.g., calculating the product) that integrates various facets of the issue together and affords further reflection” (p.
+Filling in more details, speaking from both the perspective of a researcher (of argumentation) and an educator (or, a designer of classroom activities), Nussbaum summarizes that using Bayes “forces one to (a) think about an issue from different perspectives, (b) reflect on one’s range of certainty and uncertainty and (c) conduct a simple computer simulation (e.g., calculating the product) that integrates various facets of the issue together and affords further reflection” (p.
 99).
 
-These are helpful suggestions, and they complement finer-grained instructional techniques, especially those around how students can reason about uncertainty probabilistically.
-Particularly, \textcite{bkbm18} for instance, suggest replacing probabilities with natural frequencies in alignment with a great deal of work on how humans understand probability) \parencite{gh95} as well as using and interpreting tree diagrams to depict natural frequencies.
-\textcite{me14}—building on much earlier experimental work by \textcite{wason1971natural}—provide a simple example of how even fourth graders, with the right tools to represent probabilities, can support students to transition from logical to probabilistic reasoning.
+The work of \textcite{so12} and \textcite{n11} can be complemented by finer-grained instructional techniques, especially those advanced by statistics educators to support students to reason about uncertainty probabilistically.
+A benefit of these strategies is that they can help to address some of the ingrained challenges that people and learners face when attempting to reason probabilistically \parencite[e.g., ][]{tk74}.
+Particularly, \textcite{bkbm18} for instance, suggest replacing probabilities with natural frequencies.
+This suggestion is in alignment with prior work by \textcite{gh95} on how different representations of probabilities can explain some of the challenges that people and learners face.
+In addition, \textcite{me14}—building on much earlier experimental work by \textcite{wason1971natural}—provide a simple example of how even fourth-graders, with the right tools to represent probabilities, can support students to transition from logical to probabilistic reasoning.
 These could also support the aim of developing unplugged versions of Bayesian analyses.
 
-Third, we should, of course, consider teachers and how teachers can (or already) support students to reason about uncertainty, and, possibly, to use Bayesian ideas.
+Science education researchers may consider how teachers can support students to reason about uncertainty in a way that aligns with a Bayesian perspective.
 While some scholars have contributed foundational work demonstrating how this uncertainty can be productive as a part of teaching practice \parencite{manz2018supporting}, a broad and formal perspective on how uncertainty can be approached in science education is needed.
-There is the need to explore the potential of a principled approach to uncertainty, one that may be intuitive to learners and useful to teachers, but is as yet under-represented in the science education literature.
-If learners and people better understand uncertainty, they can better understand the scientific process which helps to build trust in science.
 Further, a better understanding of uncertainty can help to learn about scientific--and other--concepts where probability plays a key role; it may be that learning about how Bayes rule can be used in legal or medical contexts could impress upon students the cross-cutting nature of probability and uncertainty.
-This is especially helpful in light of the teacher's practice of noticing and responding to students' ideas and the process through which students change their understanding.
 
-Finally, recently developed tools, particularly the JASP statistical software, may be helpful.
+Finally, recently developed tools, particularly the JASP statistical software \parencite{JASP2020}, may be helpful.
 JASP allows for both conventional frequentist as well as Bayesian versions of commonly used statistical tests, such as t-tests and regression analyses.
-There are a number of accessible guides available for specifying Bayesian versions of analyses \parencite{v20, wagenmakers2018bayesian}, and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms.
-Another possibly useful tool could be Bayesian networks (see Bayes Box; Bayes Box, 2021), which could be used to model complex system in a way that is similar to other tools used in science education for modeling complex systems, such as \textcite{20}.
+There are many accessible guides available for specifying Bayesian versions of analyses \parencite{v20, wagenmakers2018bayesian}, and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms.
+Another possibly useful tool could be Bayesian networks (see Bayes Box; Bayes Box, 2021), which could be used to model complex systems in a way that is similar to other tools used in science education for modeling complex systems, such as \textcite{20}.
 
-\section{Summary and Conclusion}
+\section{Conclusion}
 
-While some educational researchers may state that such theoretical, abstract ideas about the philosophical and statistical underpinnings of working with data are irrelevant to how students work with data, we make the case, here, that these ideas can inform instructional practice, assessments, and the design of activities that involve students in working with data, especially given some of the clear problems that have been articulated around the present frequentist Null Hypothesis Statistical Testing (NHST)-based approaches \parencite{c14, gkv04}.
-
-Science education researchers may note that these ideas are conceptually-focused, rather than aligning with the present emphasis in science education upon engaging students in science and engineering practice in order to make sense of core ideas and to engage with cross-cutting concepts \parencite{nrc12}.
-We argue that Bayes is unique in that it, perhaps akin to a cross-cutting concept in the sense of a style of reasoning \parencite{so12, ork18} that blends conceptual ideas with practices.
-Bayes is both a philosophical and psychological perspective on how to learn from data as well as a statistical framework.
+In summary--to this point--we have introduced Bayes as a powerful way of conceptualizing probability and uncertainty.
+These are two concepts that are ubiquitous in the learning, doing, and communicating of science and therefore critical aspects of science literacy.
+However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike \parencite{gkv04, s07, tk74}.
+Thus, it is not surprising when the public misinterprets statements about probability and uncertainty but the consequences may be dire when trust in science is eroded, and, sadly, examples of this are plentiful during current the COVID-19, e.g., the efficiency of masks in preventing the spread of COVID and the efficiency of the AstraZeneca vaccine were strongly questioned, delaying efforts to combat the pandemic.
+The public, however, can hardly be blamed for this in light of the prevalent findings that the frequentist conceptualizations of probability and uncertainty in school are challenging and uncommonly align with or build on peoples’ intuitions.
+As research from cognitive and psychological science as well as emerging work from statistics and statistics data science education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions \parencite{kl18, so12} about these concepts and thus has the potential to support learning about probability and uncertainty in science.
+Moreover, a Bayesian perspective is unique in that it is akin to a cross-cutting concept in the sense of a style of reasoning \parencite{so12, ork18} that blends conceptual ideas with practices: Bayes is both a philosophical and psychological perspective on how to learn from data as well as a statistical framework.
 Indeed, a key potential benefit of Bayes may be that it provides a framework for helping students to transition from their conceptual understanding of scientific ideas to quantitatively analyzing data.
 
-In this paper, we have introduced Bayes as a powerful way of conceptualizing probability and uncertainty.
-These are two concepts that are ubiquitous in the learning, doing, and communicating of science and therefore critical aspects of science literacy \parencite{}.
-However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike \parencite{gkv04, s07, tk74}.
-Thus, it is not surprising when the public misinterprets statements about probability and uncertainty but the consequences may be dire when trust in science is eroded, and, sadly, examples of this are plentiful during current the COVID-19, e.g., the efficiency of masks in preventing the spread of COVID and the efficiency of the AstraZeneca vaccine were strongly questioned, delaying efforts to comb at the pandemic.
-The public, however, can hardly be blamed for this in light of the prevalent findings that the frequentist conceptualizations of probability and uncertainty in school are challenging and uncommonly align with or build on peoples’ intuitions.
-As research from cognitive and psychological science as well as emerging work from statistics and statistics data science education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions \parencite{kl18, so12} about these concepts and thus has potential to support learning about probability and uncertainty in science.
-
-We have outlined numerous vantage points for research agendas to explore the potential of Bayes in science education so that future research can identify how students can become better reasoners in light of uncertainty.
-We have also discussed how we as scientists and science educators can contribute to the public trust in science by becoming better in how we communicate about probability and uncertainty.
-In all, Bayes is a big idea—one with epistemological and philosophical, statistical, and pedagogical implications.
-We argue that the time for advancing Bayesian methods has come, particularly in light of distrust in science and science education and the way Bayesian methods can address this state of affairs by offering a different foundation and language around a core but sometimes unsettling tenet and persistent feature, uncertainty.
+Even at a time in which science has made and continues to make great advances, the history of science suggests that uncertainty will not be removed (or even reduced) given scientific progress \parencite{fara2010science}; looking ahead, being able to effectively navigate, counsel, and act in the midst of uncertainty may be as important as it has been at other uncertain points in history.
+Like the surprising utility of a Bayesian perspective in domains that are similar to education (e.g., developmental science), might an unexpected feature of a Bayesian perspective be that embracing uncertainty--equipped with the tools that Bayesian methods provide--makes individuals and societies more confident?
+In short, can we build trust in science by embracing uncertainty?
+We think that these are questions that are \ital{likely} to yield positive results for science educators and those advocating for a more rationale and scientific citizenry and for science learners.
 
@@
 \end{document}
-
-%% 
-%% Copyright (C) 2019 by Daniel A. Weiss <daniel.weiss.led at gmail.com>
-%% 
-%% This work may be distributed and/or modified under the
-%% conditions of the LaTeX Project Public License (LPPL), either
-%% version 1.3c of this license or (at your option) any later
-%% version.  The latest version of this license is in the file:
-%% 
-%% http://www.latex-project.org/lppl.txt
-%% 
-%% Users may freely modify these files without permission, as long as the
-%% copyright line and this statement are maintained intact.
-%% 
-%% This work is not endorsed by, affiliated with, or probably even known
-%% by, the American Psychological Association.
-%% 
-%% This work is "maintained" (as per LPPL maintenance status) by
-%% Daniel A. Weiss.
-%% 
-%% This work consists of the file  apa7.dtx
-%% and the derived files           apa7.ins,
-%%                                 apa7.cls,
-%%                                 apa7.pdf,
-%%                                 README,
-%%                                 APA7american.txt,
-%%                                 APA7british.txt,
-%%                                 APA7dutch.txt,
-%%                                 APA7english.txt,
-%%                                 APA7german.txt,
-%%                                 APA7ngerman.txt,
-%%                                 APA7greek.txt,
-%%                                 APA7czech.txt,
-%%                                 APA7turkish.txt,
-%%                                 APA7endfloat.cfg,
-%%                                 Figure1.pdf,
-%%                                 shortsample.tex,
-%%                                 longsample.tex, and
-%%                                 bibliography.bib.
-%% 
-%%
-%% End of file `./samples/longsample.tex'.

### 2021-05-06 manuscript.tex
Updates from Overleaf

@@
 %% apa7.dtx  (with options: `longsample')
-%% ----------------------------------------------------------------------
+%% -----------------------------------------------
 %% 
@@
 %% 
-%% ----------------------------------------------------------------------
+%% -----------------------------------------------
 %% 
@@
 \usepackage{csquotes}
-\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
+\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber, hyperref=false]{biblatex}
+
+\usepackage[options]{nohyperref}  % This makes hyperref commands do nothing without errors
+\usepackage{url}  % This makes \url work
+
 \DeclareLanguageMapping{american}{american-apa}
@@
 
-\title{Building Trust in Science Through a Bayesian Perspective on Probability and Uncertainty}
+\title{Building Trust in Science by Advancing a Bayesian Perspective on Probability and Uncertainty in Science Education}
 \shorttitle{Building Trust Through a Bayesian Perspective}
@@
 \abstract{
-Uncertainty is ubiquitous in science, but scientific understanding is often represented to the public and in educational contexts as certain and immutable.
-Adding to this problem is that when scientists try to communicate their uncertainty, research has demonstrated that many people--learners and experts alike--struggle to interpret the statistics that are often used to make inferences about evidence from scientific data.
-We conjecture that a Bayesian perspective can offer a foundation for teaching and learning about a persistent but sometimes unsettling tension for science, making sense of and communicating about uncertainty.
+Uncertainty is ubiquitous in science, but scientific knowledge is often represented to the public and in educational contexts as certain and immutable.
+Adding to this issue is that when scientists try to communicate their uncertainty, research has demonstrated that many people--learners and experts alike--struggle to interpret the statistics that are often used to make inferences about evidence from scientific data.
+We conjecture that a Bayesian perspective can offer a coherent, holistic approach to teaching, learning, and sense-making about a persistent but sometimes unsettling tension for science, dealing with uncertainty.
 Central to this approach is viewing uncertainty \ital{probabilistically}, as doing so provides a language for expressing the degree of uncertainty in knowledge across a range of scientific domains and socio-scientific issues.
-In this paper, we elaborate on the importance of probability and uncertainty in science (and science education) and review how people learn about these concepts---and when and why they face difficulties when doing so.
-Then, we provide an accessible primer on Bayes Theorem and more general principles following from the application of Bayes Theorem.
-Next, we provide examples of how a Bayesian perspective could inform efforts around both science literacy and science education with an emphasis upon building trust in science.
-We conclude with directions for future research focused on bolstering trust in science outside of and within pre-collegiate science classrooms, where a Bayesian perspective can provide a coherent approach to engaging in science to make sense of scientific ideas in a principled yet flexible manner.
+To frame the problem and opportunity at hand, we describe prior research on uncertainty and probability in science (and science education) and review how people learn about these ideas.
+Then, we provide an accessible primer on Bayes Theorem and discuss general principles relevant to science education that follow from Bayes Theorem.
+Finally, we provide examples of how a Bayesian perspective could inform efforts around both science literacy and science education with an emphasis upon building trust in science.
+We conclude with a call for future research to use a Bayesian perspective to bolster trust in science and to provide a coherent frame for considering variability and uncertainty in data and evidence both outside of and within pre-collegiate science classrooms.
 }
@@
 A similar tension is now present concerning vaccines for COVID-19--a tension that involves reconciling robust but still initial clinical trial data and new data from population-level vaccination efforts and making policy-level recommendations (and individual decisions) amid this uncertainty.
+\\
+
 Adding to this difficult state of affairs regarding the role of uncertainty in science is that when scientists try to communicate their uncertainty, prior research has demonstrated that many people–learners and experts alike–struggle to interpret the statistics and derivations of statistics (such as standard errors/margins of error and confidence intervals) and that are often used to make inferences from scientific data \parencite{gkv04, s07, tk74}.
@@
 
-In this paper, we argue that these challenges concerning trust in scientific knowledge that not only can but must change can be addressed in a principled, flexible way, but that doing so requires making use of a Bayesian perspective that heretofore has been under-used outside of particular scientific and engineering domains---domains that often make little contact with science educators and researchers with an interest in building trust in science.
-Yet, the opportunity to do so now exists for technical and social reasons.
-Based on accumulating evidence from psychologists using Bayesian models of cognition \parencite{g12, tgk06}, statisticians and statistics educators advancing tools and practices for Bayesian methods \parencite{a02, b02, h_a20, h_j20,me14, s07} and science education scholars who have advanced a Bayesian perspective on argumentation \parencite{n11, s07}, we conjecture that a Bayesian perspective can help people--and science learners--to understand and reason about uncertainty in a way that leverages their existing ways of thinking.
-Central to this conceptualization is viewing uncertainty \ital{probabilistically}, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains \parencite{g12}.
+In this paper, we argue that these challenges concerning trust in scientific knowledge that not only can but must change can be addressed in a principled, flexible way, but that doing so requires making use of a Bayesian perspective that heretofore has been under-used outside of particular scientific and engineering domains--domains that often make little contact with science educators and researchers with an interest in building trust in science.
+We do so by drawing on research from the work of individuals from diverse disciplines: psychologists using Bayesian models of cognition \parencite{g12, tgk06}, statisticians and statistics educators advancing tools and practices for Bayesian methods \parencite{a02, b02, h_a20, h_j20,me14, s07} and science education scholars who have advanced a Bayesian perspective on argumentation \parencite{n11, s07}.
 In drawing on research from psychologists, learning scientists, and educational researchers, we make this conjecture not only on the merits of Bayes' Theorem as a mathematical or statistical tool, but primarily on evidence that the perspective we are advocating can provide an intuitive way for people to think about data and evidence in the world \parencite{g12} and in classroom contexts \parencite{ll20, n11, s07, so12, w17}.
+The primary contribution of this paper lies in drawing together different areas of past research to argue for the use of Bayesian methods in a manner and context in which Bayes has not yet become well-known or widely-used: teaching and learning in pre-collegiate (or K-12) settings.
+
+Central to the contributions of Bayesian methods is viewing uncertainty \ital{probabilistically}, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains \parencite{g12}, as we discuss next.
 
@@
 
-Probability and uncertainty are ubiquitous in scientific methodologies, scientific concepts, and in how science is communicated \textcite{gsoobmc17}.
+Probability and uncertainty are ubiquitous in scientific methodologies, scientific concepts, and in how science is communicated \parencite{gsoobmc17}.
 The practice of science often begins with measurement for which measurement error--which has a random component--is endemic.
 This random component arises from random variations in the measurement process and limits the certainty that one can have in the measurements.
-Reducing that measurement error has been critical for many discoveries in science, such as in sorting out the periodic table as we know it today \parencite{fco15}.
-However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., the recent study on the relationship between SARS-CoV-2 viral load and patient age by \textcite{jmvbzhd20} was heavily debated in the scientific community \parencite{frick_peer-review_2020}.
+Reducing that measurement error has been critical for many discoveries in science, such as sorting out the periodic table as it is known today \parencite{fco15}.
+However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., a recent study on the relationship between SARS-CoV-2 viral load and patient age by \textcite{jmvbzhd20} that was heavily debated in the scientific community \parencite{frick_peer-review_2020}.
 
 Probability and uncertainty are not only key concepts in science practices (such as analyzing and interpreting data) but are also an integral part of explanatory models of and theories for scientific concepts.
-For example, in the context of Heisenberg’s uncertainty principle and--more generally--quantum mechanics \parencite{feynman1951operator}, confidence is limited not only by measurement error but in principle---even in the perfect experimental setup there is no deterministic outcome.
+For example, in the context of Heisenberg’s uncertainty principle and--more generally--quantum mechanics \parencite{feynman1951operator}, confidence is limited not only by measurement error but in principle--even in the perfect experimental setup, there is no deterministic outcome.
 Instead, one has to calculate the probability of each possible outcome using the Born rule.
+\
+
 To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunneling \parencite{c19}.
@@
 Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers \parencite{m16}.
+\\
+
 Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public \parencite[]{gkv04, c14, mg17}.
-For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign a higher probability to the center of the interval \parencite{kl18} or \emph{p}-values are interpreted as effect sizes \parencite{gc17, n14} (although p-values describe how (in)compatible a set of data is with a set of assumptions.
-In sum, misinterpretations of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work.
+For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign a higher probability to the center of the interval \parencite{kl18} or \emph{p}-values are interpreted as effect sizes \parencite{gc17, n14} (although \ital{p}-values describe how \ital{incompatible} a set of data is with a set of assumptions).
+As \textcite[][]{jeffreys1961} describes "what the use of P implies, therefore, is that a hypothesis that may be true may be rejected because it has not predicted observable results that have not occurred" (p.
+385).
+In short, one reason why learners and people may have difficulty understanding \ital{p}-values is that they are used in a way that runs counter to how many intuitively think of them.
+In sum, misunderstandings of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work--and these can lead to misinterpretations that can diminish trust in science and the process through which scientific knowledge is constructed.
 
@@
 They found that students’ statistical reasoning capabilities are a strong predictor of both acceptance of evolution and knowledge about evolution.
+Research on teachers' conceptions of evolution has suggested that exposure to curricular materials that emphasized the role of randomness in evolution bolstered teachers' \ital{acceptance} of evolution, but not their understanding of it; the researchers conjectured that the difficulties learners face when working to understand probability and uncertainty may require extensive instruction \parencite{nadelson2010shifting}.
 
@@
 
-Indeed, past research has shown that the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners.
+Independent of the topics of quantum physics and nuclear decay, probability may generally be difficult for people to understand.
+Past research has shown that the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners.
 The work of \textcite{tk74}, especially, points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit.
@@
 Recent work--primarily by developmental psychologists and scientists--has applied Bayesian ideas to a new context: how children make sense of and world \parencite{g12, tgk06, tgk06}.
-This notion that children reason about the world in a way that is analogous to Bayes Theorem is in some ways surprising, but it also builds upon much earlier work that also documented the surprisingly complex ways that children could reason about the world---as well as the assumed limitations of children's reasoning, the work of \textcite{pi69}.
+This notion that children reason about the world in a way that is analogous to Bayes Theorem is in some ways surprising because Bayesian methods are associated with highly advanced statistical analyses, such as deciphering the Enigma machine during World War II \parencite{mcgrayne2011theory}.
+However, the notion that children (and people) weigh evidence about the world in a Bayesian way also builds upon, refines, and in cases refutes aspects of much earlier work that \ital{also} documented the surprisingly complex ways that children could reason about the world \textcite{pi69}.
 Different from earlier work on the stages through which children develop, recent work using Bayesian models considers the differences in the reasoning ability of children and adults to be a matter of degree, rather than kind.
-In short, children hold initial ideas and understandings that they change (or update) in ways that are concordant with how Bayes Theorem represents the updating of an initial belief in light of data (and how the Bayesian Learning Cycle described earlier allows individuals to integrate inductive and deductive reasoning).
+In sum, children hold initial ideas and understandings that they change (or update) in ways that are concordant with how Bayes Theorem represents the updating of an initial belief in light of data (and how the Bayesian Learning Cycle described earlier allows individuals to integrate inductive and deductive reasoning).
 
-In addition to developmental research that uses Bayes Theorem as an analog for how children make sense of data and information in light of their initial ideas and beliefs, other work bears a striking semblance to developmental research that does.
+In addition to developmental research that uses Bayes Theorem as an analog for how children make sense of data and information in light of their initial ideas and beliefs, other work bears a striking semblance to educational and developmental research that explicitly refers to and draws upon a Bayesian perspective.
 In particular, research on how learners’ prior understanding influence how they make sense of how likely different patterns in data are centers the same tension around interpreting data in light of initial beliefs that a Bayesian perspective highlights \parencite[e.g., ][]{kd88, mkm07, ssc07}.
-While these studies have focused on how students’ prior ideas and data interact, Bayesian frameworks were not used to formalize this process of integrating data.
+In addition, prior research on learners' perceptions of the plausibility of a scientific explanation in light of alternative explanations highlights the importance of the degrees of belief (and re-assessing these (probabilistic, rather than dichotomous) degrees of belief in light of new information or experiences) that learners hold \parencite[e.g., ][]{lombardi2013plausibility, lombardi2016plausibility}.
+This work suggests the merit of a Bayesian perspective as a potential frame or theoretical account for (in these key instances) analyzing patterns in data and conceptual change; it also has other benefits for teaching and learning.
+\\
 
-A Bayesian view of how children learn from their experiences and update their beliefs about the world has become especially prominent over the past 10-15 years \parencite{gt07, gw12}.
-Because of this, one might assume that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education and developmental scientists have pointed to this opportunity.
+Because of the advances made my research using a Bayesian perspective to understand human \ital{development} \parencite{gt07, gw12}, one might believe that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education and developmental scientists have pointed to this opportunity.
 For instance, \textcite{g12} wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p.
 1627).
-
-However, despite arguments such as Gopnik's (2012) for integrating developmental and educational research, it has largely not been the case that “science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” \textcite[p.
+However, it has largely not been the case that “science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” \textcite[p.
 1627]{g12}, as some science education scholars have bemoaned \parencite{ls15}).
 Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas \parencite[e.g., ][]{pi69} that represent the greatest contribution of developmental ideas to education \textcite{ls07}.
-We argue that this is still largely the case in the present, despite advances in both developmental research as well as statistical methodology related to Bayesian ideas—and that it is time to start exploring how bringing out the inherent Bayesian ways of thinking in education may provide support for learners’ efforts to make sense of the world.
+\\
 
-At the same time that opportunities to connect developmental science with education (especially science education), science education research has emphasized how important students' prior knowledge and understandings are to their understanding of scientific concepts and processes \parencite[e.g., ][]{ssc07}.
-For example, \textcite{mkm07} summarized findings across several studies showing that when elementary grade students reasoned with data in a context about which they had strong and valid prior understandings that informed their interpretation of data (about force and motion in the context of ramps), and they were able to explain sources of error in the data \parencite{mk03}, but when the data was in a context they did not understand well, they struggled to distinguish between random variation due to sources of error and sources of variation due to something that was causing a difference in the data (generated by measuring the motion of pendulums).
-Later design- and laboratory-based research has shown that students can develop an understanding of scientific ideas through the analysis of data \parencite{ls04,mkk17}.
-
-In summary, developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in education--especially science education--to design instruction based around eliciting and understanding students’ ideas \parencite{gb16, hbavb20, wtbs12}.
+At the same time that opportunities to connect developmental science with science education, scholars have emphasized how important students' prior knowledge and understandings are to their understanding of scientific concepts and processes \parencite[e.g., ][]{ssc07, mkm07, mkk17, ls04}--highlight a potentially powerful connection between Bayesian methods and the priorities of science educators and science education researchers.
+Developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in education--especially science education--to design instruction based around eliciting and understanding students’ ideas \parencite{gb16, hbavb20, wtbs12}.
 There is an opportunity to use Bayesian ideas as a framework for formalizing these ideas about how prior ideas affect students’ reasoning about data and how teachers can leverage students’ ideas to design instruction, but a Bayesian perspective has--with a few exceptions focusing on how Bayes Theorem can provide a model for the science practice of argumentation \parencite[e.g., ][]{n11, so12}--has been absent from research in science education.
-
-After having reviewed how people learn about probability and uncertainty and where they face difficulties and research on uncertainty in the context of analyzing data, we next formally introduce Bayes Theorem and describe the broader process through which one's beliefs can be revised, or updated, in a way that leverages both inductive reasoning from data and deductive reasoning from beliefs or principles--what we refer to as the \ital{Bayesian Learning Cycle}.
-We also describe some of the ways of knowing, or epistemological considerations, that follow from Bayes Theorem and which comprise a core part of the Bayesian perspective that we argue has a role in building trust in science.
+A Bayesian perspective may also have a role in building trust in science by representing uncertainty in a principled yet flexible way.
+\\
 
@@
 \begin{equation*}
-        \text{Posterior uncertainty} = \text{Prior uncertainty} \times \text{Predictive updating factor}.
+        \text{Posterior uncertainty} = \text{Prior uncertainty} \times \text{Predictive updating factor}
 \end{equation*}
@@
 First, $p(\text{data} \given \theta)$ is the likelihood, that is, the extent to which the observed data were expected under $\theta$.
-Second, $p(\text{data}$ is the marginal likelihood, that is, the extent to which the observed data were expected without conditioning on $\theta$.
+Second, $p(\text{data})$ is the marginal likelihood, that is, the extent to which the observed data were expected without conditioning on $\theta$.
 If the observed data are more expected (i.e., better predicted) when conditioning on $\theta$, the predictive updating factor is higher than 1, and $\theta$ becomes more credible.
@@
 
-Although we demonstrated this updating process in a single step, it could also be executed sequentially, one student at a time.
-For instance, the first observation you make is a $W$ and in light of your prior this makes Windows usage a little more likely than MacOS usage.
+Although we have demonstrated this updating process in a single step, it could also be executed sequentially, one student at a time.
+For instance, the first observation you make is a $W$, and in light of your prior, this makes Windows usage a little more likely than MacOS usage.
 This slight change in knowledge is reflected in the difference between the distributions on the top two lines in Figure~\ref{fig:WindowsMacSequential}; line ``0'' represents the dome-shaped prior distribution and line ``1'' represents the posterior distribution after the first observations.
@@
 It should be emphasized that although the updating process is conceptually different, the final outcome is identical: the shape of the posterior distribution is unaffected by whether data arrive simultaneously or sequentially.
+\\
 
@@
 
-How do the principles from the above example translate to scientific data?
-A recent case from the physical sciences can serve to illustrate this approach.
+A recent case from the physical sciences can serve to illustrate how the principles from the example in the previous section apply to science and science education.
 In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light \parencite{b11} - and thus defying Einstein’s theory of relativity.
@@
 Thus, a single instance of apparently faster-than-light elementary particles does not give a likelihood strong enough to change the prior in a meaningful way, and in consequence, the posterior--the knowledge about the world after making an observation--does not change in a meaningful way.
+\\
+
 Thus, rather than changing their belief about the speed of light, scientists looked for things that could have gone wrong with the measurement.
@@
 
+In this section, we build on the idea of the Bayesian Learning Cycle--that a Bayesian perspective aligns with a way of systematically taking in new information to improve what one knows (and how certain one is about what one knows--to elaborate on some general principles related to the nature of knowledge and knowing, or epistemological principles.
+We illustrate these to later show how they can have a role in the science classroom and how they can be used by individuals outside of classrooms and schools to understand and act upon scientific information.
+
 In the previous section, we emphasized the idea that hypotheses, or ideas, gain credibility when they predict the data well, and lose credibility when they predict new data poorly \parencite{WagenmakersEtAl2016CD}.
@@
 \begin{quotation}
-``This inductive pattern says nothing surprising.
+
+\noindent ``This inductive pattern says nothing surprising.
 On the contrary, it expresses a belief which no reasonable person seems to doubt: \emph{The verification of a consequence renders a conjecture more credible}.
@@
 Equation~\ref{eq:BayesRulePropto} emphasizes that the posterior uncertainty about $\theta$ is a \emph{compromise} between our prior uncertainty about $\theta$ and the predictive performance of $\theta$.
-But, the posterior uncertainty after having observed the first datum becomes the prior uncertainty for the next datum---and any other data (cf.
+But, the posterior uncertainty after having observed the first datum becomes the prior uncertainty for the next datum--and any other data (cf.
 Figure~\ref{fig:WindowsMacSequential}).
@@
 If someone already knows for certain that a hypothesis is true or false, then you cannot adjust your opinion in light of the data: your initial opinion will be your final opinion, no matter what the data may indicate.
-Philosophically, adopting such priors is a dangerous practice; for instance, in antiquity the adage of the \emph{New Academy} --a school of skeptics headed by Carneades-- was ``never assert absolutely''.
+Philosophically, adopting such priors is a dangerous practice; for instance, in antiquity, the adage of the \emph{New Academy} --a school of skeptics headed by Carneades-- was ``never assert absolutely''.
 In modern times, Lindley popularized this idea in statistics and coined it ``Cromwell's rule'' \parencite[p.
@@
 \end{align}
-When we divide the posterior probabilities the common term, $p(\text{data})$, drop out, and we have:
+When we divide the posterior probabilities, the common term, $p(\text{data})$, can be dropped from each side of the equation, and we have:
 \begin{equation}
@@
 This is in stark contrast to $p$-value statistical hypothesis testing, in which a statistical model (the null hypothesis) is judged in isolation.
+\\
+
 Having introduced Bayes Theorem and some principles that may follow from it, we next make the case how Bayes could support science literacy at large before delving into implications for science teaching and learning.
 
-\section{A Bayesian Perspective and Science Literacy}
+\section{How Bayesian Principles Could Bolster Science Literacy}
 
-We conceive that Bayesian methods can enhance science literacy or a range of capabilities that people may possess which have value (individually or collectively) to them \parencite{council2016science}.
-These capabilities range from core \ital{literacies}, including numeracy and textual literacy (what is mostly thought of in common parlance concerning literacy), content knowledge, and epistemic knowledge \parencite{council2016science}.
-A potentially powerful way in which Bayesian ideas can inform discussions of science literacy is to provide individuals with a principle for combining their prior understanding with new evidence.
-Along this line, Bayes and the Bayesian Learning Cycle can provide a tool for people (and learners) to combine their understanding with new evidence (and data) rigorously and coherently.
+We conceive that Bayesian methods can enhance \ital{science literacy}, capabilities range from core \ital{literacies}, including numeracy and textual literacy (what is mostly thought of in common parlance concerning literacy), content knowledge, and epistemic knowledge \parencite{council2016science}.
+The specific ways we envision a Bayesian perspective bolstering science literacy are also techniques that teachers can appropriate in their planning and teaching; they may already align with how they teach about science through engaging in scientific practices \parencite{nrc12} or how they articulate the nature of science \parencite{lederman2020avoiding}.
 
-A timely example can help to illustrate how a Bayesian perspective on a topic relevant to science literacy may offer a different way forward.
-Consider a salient topic, medical vaccines for diseases, such as Polio and the disease caused by the coronavirus SARS-CoV-2, COVID-19.
+Consider a salient topic as a context to apply these principles, medical vaccines for diseases, such as the disease caused by the coronavirus SARS-CoV-2, COVID-19.
 Most evidence suggests that vaccines have been safe, historically, and are generally safe, at present (Centers for Disease Control and Prevention, 2021).
@@
 Some use long-established methods of vaccine development (namely, deactivating SARS-CoV-2 particles but retaining cell signaling proteins to which humans develop an immune response—and, after, a degree of immunity) and injecting these deactivated virus particles into people), whereas others use a new mechanism (namely, injecting messenger RNA into people; this messenger RNA is then transcribed by cells in people’s bodies to produce a particular protein common to SARS-CoV-2 particles, which human’s bodies then learn to recognize to develop immunity.
+We note that--at present--diseases for which it has been difficult to develop effective vaccines, including Malaria, may have viable mRNA vaccines in the near future \parencite{datoo2021high}.
 
-How should we consider evidence about the vaccines?
-Before any data from medical clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and based on similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are.
+Before any data from clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and based on similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are.
 After initial trials were completed, the degree of safety that individuals associate with these vaccines—in light of the prior and the limited evidence—could increase, but still not be near the level of safety that many would consider acceptable.
@@
 
-While the complex weighting of the evidence above may seem natural to a scientist, it is challenging for many.
+While the weighting of the evidence in light of prior research and theoretical accounts above may be natural to a scientist, it is challenging for many.
 In the following, we showcase how several important insights for weighting evidence follow directly from Bayes' rule.
-Thus, we argue that emphasizing Bayesian reasoning can support science literacy by providing people with a principled way to weighting evidence that is accessible because it builds on peoples' prior ideas.
+We do not think or expect that members of the public can do this, but members of the public (and learners) do have viable ideas and hypotheses about how the world works.
+Thus, we argue that emphasizing Bayesian reasoning can support science literacy by providing people with a principled way to weight evidence that is accessible because it builds on peoples' prior ideas.
 
@@
 In sum, a Bayesian perspective offers a general approach to science literacy that can account for uncertainty in a rigorous and principled manner.
-In addition, a number of principles that follow from Bayes Theorem, such as being receptive to new evidence and weighting that evidence per its strength in light of the strength of one's initial beliefs or understanding could bolster science literacy---and how the public comes to understand and gain trust in science.
+In addition, a number of principles that follow from Bayes Theorem, such as being receptive to new evidence and weighting that evidence per its strength in light of the strength of one's initial beliefs or understanding could bolster science literacy--and how the public comes to understand and gain trust in science.
 
-\section{A Bayesian Perspective and Science Education}
+\section{Toward a Bayesian Perspective in Science Education}
 
@@
 
-\textcite{n11} work on Bayesian approaches to argumentation complement the work of Szu and Osborne in that Nussbaum describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation.
+\textcite{n11} work on Bayesian approaches to argumentation describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation.
 Concerning the latter, Nussbaum describes how the social issue of raising taxes to provide resources to homeless individuals was a rich context for students to engage in forms of Bayesian reasoning.
@@
 
-In summary, the work of \textcite{so12} and \textcite{n11} begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process.
-Curiously, no work has applied Bayesian methods to another science practice, that of analyzing and interpreting data; one for which Bayesian methods have clear conceptual applicability and utility, but one for which the absence of suitable tools may provide to be a barrier.
-In this way, Bayesian methods could serve as a cross-cutting concept \parencite{nrc12} that unites the practice of argumentation with analyzing and interpreting data and using mathematics and computational thinking.
-
-Bayesian methods have some utility in science education, even if that utility is under-realized at present.
-Next, we present some of the specific ways Bayesian methods could enhance science education—and could do so in a way that makes Bayes more tractable.
+The work of \textcite{so12} and \textcite{n11} begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process.
+Neither, though, speaks to how a Bayesian perspective could have a role in bolstering understanding or trust in science by learners and others.
+More generally, these two contributions are steps toward conceiving what a Bayesian perspective may offer science education, but there remains room to build further.
+For instance, despite their association with statistical methods that are growing in use in post-secondary contexts and in industry \parencite{mcgrayne2011theory}, no science education scholarship has applied Bayesian methods to another science practice, that of analyzing and interpreting data.
+In this way, being clear about what we mean by a \ital{Bayesian perspective} is important: this perspective could be associated not only with a \ital{practice} (analyzing data) but also as a cross-cutting concept \parencite{nrc12} that can, in the context of learning about particular scientific ideas, unite the practice of argumentation with analyzing and interpreting data with the notions of probability and (un)certainty.
+Bayesian methods can also provide a frame that includes a set of epistemological principles and a partial explanation for the process of how and why people change their ideas in light of new information or data--two key components of (and potential barriers to) how members of the public understand and act upon scientific knowledge \parencite{sinatra2014addressing}.
 
@@
 
-The work of \textcite{so12} and \textcite{n11} can be complemented by finer-grained instructional techniques, especially those advanced by statistics educators to support students to reason about uncertainty probabilistically.
+If a Bayesian perspective is to be realized in classroom contexts, then specific, fine-grained instructional techniques will need to be developed.
+Such instructional techniques do not necessarily need to include technology or statistical software; \ital{unplugged} activities may be useful, and science education scholars could draw from research in other disciplines \parencite[e.g., ][]{caeli2020unplugged} as well as research on the role of epistemic tools in science education \parencite[e.g., ][]{stroupe2019introduction} to design these.
+
+Activities could also be adapted from those designed by statistics educators to support students to reason about uncertainty probabilistically.
 A benefit of these strategies is that they can help to address some of the ingrained challenges that people and learners face when attempting to reason probabilistically \parencite[e.g., ][]{tk74}.
@@
 These could also support the aim of developing unplugged versions of Bayesian analyses.
+\\
 
@@
 Further, a better understanding of uncertainty can help to learn about scientific--and other--concepts where probability plays a key role; it may be that learning about how Bayes rule can be used in legal or medical contexts could impress upon students the cross-cutting nature of probability and uncertainty.
+\\
 
 Finally, recently developed tools, particularly the JASP statistical software \parencite{JASP2020}, may be helpful.
-JASP allows for both conventional frequentist as well as Bayesian versions of commonly used statistical tests, such as t-tests and regression analyses.
+JASP allows for both conventional frequentist as well as Bayesian versions of commonly used statistical tests, such as \ital{t}-tests and regression analyses.
 There are many accessible guides available for specifying Bayesian versions of analyses \parencite{v20, wagenmakers2018bayesian}, and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms.
-Another possibly useful tool could be Bayesian networks (see Bayes Box; Bayes Box, 2021), which could be used to model complex systems in a way that is similar to other tools used in science education for modeling complex systems, such as \textcite{20}.
+Another possibly useful tool could be Bayesian networks \parencite[e.g., ][]{bayesbox}, which could be used to model complex systems in a way that is similar to other tools used in science education for modeling complex systems, such as \textcite{20}.
 
@@
 
-In summary--to this point--we have introduced Bayes as a powerful way of conceptualizing probability and uncertainty.
-These are two concepts that are ubiquitous in the learning, doing, and communicating of science and therefore critical aspects of science literacy.
+Probability and uncertainty are two concepts that are ubiquitous in the learning, doing, and communicating of science and therefore critical aspects of science literacy.
 However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike \parencite{gkv04, s07, tk74}.
 Thus, it is not surprising when the public misinterprets statements about probability and uncertainty but the consequences may be dire when trust in science is eroded, and, sadly, examples of this are plentiful during current the COVID-19, e.g., the efficiency of masks in preventing the spread of COVID and the efficiency of the AstraZeneca vaccine were strongly questioned, delaying efforts to combat the pandemic.
-The public, however, can hardly be blamed for this in light of the prevalent findings that the frequentist conceptualizations of probability and uncertainty in school are challenging and uncommonly align with or build on peoples’ intuitions.
-As research from cognitive and psychological science as well as emerging work from statistics and statistics data science education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions \parencite{kl18, so12} about these concepts and thus has the potential to support learning about probability and uncertainty in science.
+\\
+
+We think the blame does not squarely lie with the public in light of findings that the traditional representations of probability and uncertainty are challenging for learners and people alike to comprehend.
+As research from cognitive and psychological science as well as emerging work from statistics and statistics education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions \parencite{kl18, so12, jeffreys1961} about these concepts and thus has the potential to support learning about probability and uncertainty in science.
 Moreover, a Bayesian perspective is unique in that it is akin to a cross-cutting concept in the sense of a style of reasoning \parencite{so12, ork18} that blends conceptual ideas with practices: Bayes is both a philosophical and psychological perspective on how to learn from data as well as a statistical framework.
+
 Indeed, a key potential benefit of Bayes may be that it provides a framework for helping students to transition from their conceptual understanding of scientific ideas to quantitatively analyzing data.
+In this way, we argue for the use of a Bayesian \ital{perspective}, one that could influence research about (and teaching to support) learners' epistemic considerations and how learners' ideas change over time--factors that may have a bearing on how they understand science outside of school, or later as adults \parencite{sinatra2014addressing}.
+Thus, we see there is a number of ways in which a Bayesian perspective can influence science education--and, therefore, we see  a number of mechanisms through which Bayesian methods could bolster trust in science.
+Research and theory suggest that Bayesian ideas could have an important role in science education, but the number of ways that it could make this prospect daunting.
+Different from past research focused on the connection between argumentation and Bayes Theorem, our aim with this paper was to highlight the broad potential of a Bayesian perspective for considering variability and uncertainty in data and evidence and to offer a few directions for how the potential of Bayes in future research.
 
@@
 In short, can we build trust in science by embracing uncertainty?
-We think that these are questions that are \ital{likely} to yield positive results for science educators and those advocating for a more rationale and scientific citizenry and for science learners.
+We think that these are questions that are \ital{likely} to yield positive results for science educators and those advocating for a more scientifically-informed citizenry--and for science learners.
 

### 2021-09-22 manuscript.tex
updates for revision

@@
 %%
-\documentclass[man]{apa7}
-% \documentclass[man, mask, floatsintext]{apa7}
+
+% response letter for Science & Education paper: https://docs.google.com/document/d/11knIlTG488LM4kw5y8irPS8CqqBbTbdJJyGeXEFmhqY/edit
+
+% will remove the below for submission; for reference now
+% https://kubsch.shinyapps.io/Confidence_Updater
+% https://github.com/MKubsch/confidence_updater
+
+\documentclass[man, floatsintext]{apa7} % JR: for some reason, the figures aren't floating in text
 
@@
 \usepackage{amsfonts, amsmath,amssymb}
+\usepackage{comment}
 
@@
 \newcommand{\given}{\, | \,}
-\newcommand{\EJ}[1]{\todo[inline, color=green]{  #1 }}
+\newcommand{\EJ}[1]{\todo[inline, color=yellow]{  #1 }}
 \newcommand{\JR}[1]{\todo[inline, color=gray]{  #1 }}
 \newcommand{\MK}[1]{\todo[inline, color=red]{  #1 }}
+\newcommand{\MD}[1]{\todo[inline, color=green]{  #1 }}
 
@@
 
-\usepackage[options]{nohyperref}  % This makes hyperref commands do nothing without errors
 \usepackage{url} % This makes \url work
@@
 
-% \setlength{\parindent}{8ex}
+\title{Why and How a Bayesian Approach Supports Science Educators and Learners to Reason Under Scientific Uncertainty}
+\shorttitle{Reasoning Under Scientific Uncertainty}
 
-% \usepackage{authblk}
-
-\title{Building Trust in Science by Advancing a Bayesian Perspective on Probability and Uncertainty in Science Education}
-\shorttitle{Building Trust Through a Bayesian Perspective}
-
-\threeauthors{Joshua M. Rosenberg}{Marcus Kubsch}{Eric-Jan Wagenmakers}
-\threeaffiliations{University of Tennessee, Knoxville}{IPN - Leibniz Institute for Science and Mathematics Education}{University of Amsterdam}
-
-% \authorsnames[1,2,3]{Joshua M. Rosenberg, Marcus Kubsch, Eric-Jan Wagenmakers}
-% \authorsaffiliations{{University of Tennessee, Knoxville},{IPN - Leibniz Institute for Science and Mathematics Education},{University of Amsterdam}}
+\fourauthors{Joshua M. Rosenberg}{Marcus Kubsch}{Eric-Jan Wagenmakers}{Mine Dogucu}
+\fouraffiliations{University of Tennessee, Knoxville}{IPN - Leibniz Institute for Science and Mathematics Education}{University of Amsterdam}{University of California, Irvine}
 
@@
 Uncertainty is ubiquitous in science, but scientific knowledge is often represented to the public and in educational contexts as certain and immutable.
-Adding to this issue is that when scientists try to communicate their uncertainty, research has demonstrated that many people--learners and experts alike--struggle to interpret the statistics that are often used to make inferences about evidence from scientific data.
-We conjecture that a Bayesian perspective can offer a coherent, holistic approach to teaching, learning, and sense-making about a persistent but sometimes unsettling tension for science, dealing with uncertainty.
-Central to this approach is viewing uncertainty \ital{probabilistically}, as doing so provides a language for expressing the degree of uncertainty in knowledge across a range of scientific domains and socio-scientific issues.
-To frame the problem and opportunity at hand, we describe prior research on uncertainty and probability in science (and science education) and review how people learn about these ideas.
-Then, we provide an accessible primer on Bayes Theorem and discuss general principles relevant to science education that follow from Bayes Theorem.
-Finally, we provide examples of how a Bayesian perspective could inform efforts around both science literacy and science education with an emphasis upon building trust in science.
-We conclude with a call for future research to use a Bayesian perspective to bolster trust in science and to provide a coherent frame for considering variability and uncertainty in data and evidence both outside of and within pre-collegiate science classrooms.
+This contrast can foster distrust when scientific knowledge develops in ways that can be perceived as reversals, e.g., consider changing recommendations regarding mask wearing during the COVID pandemic.
+Adding to this issue is that when scientists try to communicate their uncertainty, research has demonstrated that many people--learners and experts alike--struggle to interpret the statistics that are used to make draw inferences from scientific data.
+We argue that a Bayesian perspective can offer a coherent, holistic solution to these problems.
+Central to a Bayesian approach to teaching, learning, and sense-making is viewing uncertainty \ital{probabilistically}, as doing so provides a language for expressing the degree of beliefs about what is known about the world.
+In this paper, we describe prior research on uncertainty and probability in science (and science education) and review how people learn about these ideas.
+Then, we provide a primer on Bayes' Theorem and discuss general principles relevant to science education that follow from Bayes' Theorem.
+Finally, we focus on \ital{how} to make Bayesian approaches to reasoning about uncertainty practical for science educators through the introduction of an interactive, accessible tool.
+We conclude with a call for future research to use a Bayesian perspective to bolster trust in science and to provide a coherent frame for considering  uncertainty in data and evidence within and beyond science classrooms.
 }
@@
    \addORCIDlink{Eric-Jan Wagenmakers}{0000-0003-1596-1034} 
+   \addORCIDlink{Mine Dogucu}{0000-0002-8007-934X}
    \\
@@
 \begin{document}
+
 \maketitle
 
-\setlength{\epigraphwidth}{1.0\textwidth} % set for individual epigraph
-\epigraph{It is seen in this essay that the theory of probabilities is at bottom only common sense reduced to calculus; it makes us appreciate with exactitute that which exact minds feel by a sort of instinct without being able ofttimes to give a reason for it .
-.
-.
-we shall see that there is no science more worthy of our meditations, and that no more useful one could be incorporated in the system of public instruction.}{Pierre-Simon Laplace, 1829}
-\setlength{\epigraphwidth}{0.4\textwidth} % return to default for the next epigraphs
-%\nocite{Laplace18291902} %[p. 196]
+Uncertainty is ubiquitous in science.
+Consider, for instance, uncertainty due to measurement error or temporal variability in the characteristics under study.
+\parencite{fuller2009measurement}.
+Another form of uncertainty is its central role within explanatory theories--such as those associated with evolution and quantum physics--for which models take the form of probabilistic statements about the world \parencite{g00}.
 
-Uncertainty is ubiquitous in science and creating ways to account for this uncertainty has helped science to progress.
-Consider, for instance, uncertainty due to the measurements or instruments used to study phenomena: In such cases, error due to measurement can be incorporated into models through the repeated measures of the same object or phenomenon \parencite{pls03}.
-Measurement is also present when the object that is measured--like a plant--is variable in its characteristics because of environmental and genetic variation (and interactions among these sources of variation) between the individual plants \parencite{pls03}.
-Another form of uncertainty is accounted for in terms of its central role within explanatory theories--such as those associated with evolution and quantum physics--for which models take the form of probabilistic statements about the world \parencite{g00}.
-
-Negotiating the many forms of uncertainty through (constructively) arguing—-and building toward consensus at the same time that uncertainty is present—-is a key part of the scientific process \parencite{duschl2008science, t00, nrc12}.
-Concomitantly, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions \parencite{g10, g06, n02, r77}.
-However, while scientists and historians and philosophers of science are well aware of the uncertainty (and limitations) around scientific knowledge and the practice of negotiating and arguing from evidence, scientific knowledge is often represented to the public and in educational contexts as certain and immutable \parencite{d90}.
-Trust in science can erode and has eroded when scientists individually or collectively change their views, such as during the first phase of the COVID-19 crisis, when scientists faced serious criticism from the public, the media, and elected representatives for changing their recommendations based upon new scientific data and findings.
-A similar tension is now present concerning vaccines for COVID-19--a tension that involves reconciling robust but still initial clinical trial data and new data from population-level vaccination efforts and making policy-level recommendations (and individual decisions) amid this uncertainty.
+Negotiating the many forms of uncertainty through constructively arguing--and building toward consensus at the same time that uncertainty is present--is a key part of the scientific process \parencite{g00}.
+It is also a key part of science education \parencite{duschl2008science, t00, nrc12, manz2018supporting, so12, n11}.
+Concomitantly, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions \parencite{g10,  n02, r77, lakatos1976falsification, feyerabend1993against, carnap1935philosophy, Kuhn1962}.
 \\
 
-Adding to this difficult state of affairs regarding the role of uncertainty in science is that when scientists try to communicate their uncertainty, prior research has demonstrated that many people–learners and experts alike–struggle to interpret the statistics and derivations of statistics (such as standard errors/margins of error and confidence intervals) and that are often used to make inferences from scientific data \parencite{gkv04, s07, tk74}.
-Similarly, difficulties in understanding the technical language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution and nuclear decay that are modeled in terms of probabilistic statements \parencite{fth17}.
+However, while scientists, historians, and philosophers of science are well aware of the uncertainty and limitations around uncertain scientific knowledge and negotiating and arguing from evidence \parencite{polanyi1962tacit, polanyi1966logic}, scientific knowledge is often represented to the public and in educational contexts as certain and immutable \parencite{d90, manz2018supporting, carey1993understanding}.
+Trust in science can erode and has eroded when scientists individually or collectively change their views, such as during the first phase of the COVID-19 crisis, when scientists faced serious criticism from the public, the media, and elected representatives for changing their recommendations based upon new scientific data and findings \parencite[]{vanderbleseffects2020,kreps_model_2020}.
+A similar tension between how certain scientific knowledge is perceived and how uncertain it really is concerns vaccines for COVID-19--a tension that involves reconciling robust but still initial clinical trial data and new data from population-level vaccination efforts.
+This tension also involves regulatory organizations, politicians, and medical, pharmaceutical, and scientific experts making policy-level recommendations (and individual decisions) amid uncertainty \parencite[]{vanderbleseffects2020, kreps_model_2020}.
+\\
 
-In this paper, we argue that these challenges concerning trust in scientific knowledge that not only can but must change can be addressed in a principled, flexible way, but that doing so requires making use of a Bayesian perspective that heretofore has been under-used outside of particular scientific and engineering domains--domains that often make little contact with science educators and researchers with an interest in building trust in science.
-We do so by drawing on research from the work of individuals from diverse disciplines: psychologists using Bayesian models of cognition \parencite{g12, tgk06}, statisticians and statistics educators advancing tools and practices for Bayesian methods \parencite{a02, b02, h_a20, h_j20,me14, s07} and science education scholars who have advanced a Bayesian perspective on argumentation \parencite{n11, s07}.
-In drawing on research from psychologists, learning scientists, and educational researchers, we make this conjecture not only on the merits of Bayes' Theorem as a mathematical or statistical tool, but primarily on evidence that the perspective we are advocating can provide an intuitive way for people to think about data and evidence in the world \parencite{g12} and in classroom contexts \parencite{ll20, n11, s07, so12, w17}.
-The primary contribution of this paper lies in drawing together different areas of past research to argue for the use of Bayesian methods in a manner and context in which Bayes has not yet become well-known or widely-used: teaching and learning in pre-collegiate (or K-12) settings.
+In this paper, we argue that these challenges concerning trust in scientific knowledge can be addressed by considering scientific knowledge not as correct or incorrect or true or false, but rather to consider the degrees of belief that one might express in light of prior information and new evidence.
+In other words, we argue for the use of a Bayesian perspective that emphasizes \ital{subjective probability} \parencite[]{batanero2016research,, konold1991understanding}.
+Particularly, we argue that some of the challenges around trust in science can begin to be addressed by taking a Bayesian perspective to how people--and students--understand uncertainty.
+Thus, this paper is \ital{not} about how scientists, statisticians might use Bayesian methods; books address this topic \parencite[]{gelman1995bayesian, Kruschke2015Book}, and papers now exist on how educational researchers can use such methods \parencite{kubsch2021beyond, levy2016advances}.
+Instead, we claim that \ital{science learners} can use Bayesian reasoning to be able to reason about uncertainty in a more principled and flexible manner.
+We make this argument by drawing on research from the work of individuals from diverse disciplines: a) psychologists using Bayesian models of cognition \parencite{g12, tgk06}, b) statisticians and statistics educators advancing tools and practices for Bayesian methods \parencite{a02, b02, h_a20, h_j20,me14, s07} and c) science education scholars who have advanced a Bayesian perspective on argumentation \parencite{n11, s07} and scientific reasoning \parencite{warren_quantitative_2018,warren_impact_2020}.
+In doing so, we claim that Bayes' Theorem does not only have use as a mathematical or statistical tool, but also that Bayesian reasoning can provide an intuitive way for people to think about data and evidence under uncertainty \parencite{g12} and in classroom contexts \parencite{ll20, n11, s07, so12, w17, warren_impact_2020}.
+\\
 
-Central to the contributions of Bayesian methods is viewing uncertainty \ital{probabilistically}, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains \parencite{g12}, as we discuss next.
+As this introduction has suggested, central to the contributions of Bayesian methods is viewing uncertainty \ital{probabilistically}, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains \parencite{g12}, as we discuss next.
+\\
 
-\section{Uncertainty in Science and the Central Role of Probability}
+\section{Uncertainty in Science}
 
 Probability and uncertainty are ubiquitous in scientific methodologies, scientific concepts, and in how science is communicated \parencite{gsoobmc17}.
-The practice of science often begins with measurement for which measurement error--which has a random component--is endemic.
-This random component arises from random variations in the measurement process and limits the certainty that one can have in the measurements.
-Reducing that measurement error has been critical for many discoveries in science, such as sorting out the periodic table as it is known today \parencite{fco15}.
-However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., a recent study on the relationship between SARS-CoV-2 viral load and patient age by \textcite{jmvbzhd20} that was heavily debated in the scientific community \parencite{frick_peer-review_2020}.
-
-Probability and uncertainty are not only key concepts in science practices (such as analyzing and interpreting data) but are also an integral part of explanatory models of and theories for scientific concepts.
-For example, in the context of Heisenberg’s uncertainty principle and--more generally--quantum mechanics \parencite{feynman1951operator}, confidence is limited not only by measurement error but in principle--even in the perfect experimental setup, there is no deterministic outcome.
-Instead, one has to calculate the probability of each possible outcome using the Born rule.
-\
-
-To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunneling \parencite{c19}.
-Another example of how probability and uncertainty are central to learning scientific concepts is evolution \parencite{fsnh19}.
-Evolution is modeled as a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically.
-Further, the continued survival of these variations is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modeled probabilistically.
-Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers \parencite{m16}.
+The practice of science often begins with \ital{measurement}, which has a random component \parencite{fuller2009measurement} that arises from random variations in the measurement process.
+Measurement error limits the certainty that one can have in the measurements and reducing that measurement error has been critical for many discoveries in science, such as the sorting out of the periodic table that has resulted in the representation as it is known today \parencite{fco15}.
+Deciding on adequate statistical procedures to determine measurement uncertainty is not always easy and it regularly sparks debate, e.g., consider a recent study on the relationship between SARS-CoV-2 viral load and patient age by \textcite{jmvbzhd20} that was heavily debated in the scientific community \parencite{frick_peer-review_2020} because the authors initially discretized a continuous variable, reducing statistical power.
 \\
 
-Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public \parencite[]{gkv04, c14, mg17}.
-For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign a higher probability to the center of the interval \parencite{kl18} or \emph{p}-values are interpreted as effect sizes \parencite{gc17, n14} (although \ital{p}-values describe how \ital{incompatible} a set of data is with a set of assumptions).
-As \textcite[][]{jeffreys1961} describes "what the use of P implies, therefore, is that a hypothesis that may be true may be rejected because it has not predicted observable results that have not occurred" (p.
-385).
-In short, one reason why learners and people may have difficulty understanding \ital{p}-values is that they are used in a way that runs counter to how many intuitively think of them.
-In sum, misunderstandings of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work--and these can lead to misinterpretations that can diminish trust in science and the process through which scientific knowledge is constructed.
+Probability and uncertainty are not only integral to some science practices (such as analyzing and interpreting data; see \textcite[]{nrc12}) but are also a part of explanatory models of and theories for scientific concepts.
+For example, in the context of Heisenberg’s uncertainty principle--more generally in quantum mechanics \parencite{feynman1951operator}--confidence is limited not only by measurement error but in principle--even in the perfect experimental setup, there is no deterministic outcome.
+Instead, one has to calculate the probability of each possible outcome using the Born rule.
+To this day, there is an ongoing argument about how to interpret the puzzling quantum mechanical phenomena such as quantum tunneling \parencite{c19}.
+\\
+
+Another example of how probability and uncertainty are central to scientific concepts and respective learning activities is evolution \parencite{fsnh19}.
+Evolution is modeled as a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically.
+Further, the continued survival of these variations with the population is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modeled probabilistically.
+Similar to quantum physics, how to interpret this probability has also sparked debate among researchers \parencite{m16}.
+\\
+
+Given that probability and uncertainty spark and sustain debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public \parencite[]{gkv04, c14, mg17}.
+For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign a higher probability to the center of the interval \parencite{kl18}.
+Another example concerns very low \ital{p}-values, which are sometimes misinterpreted as effect sizes \parencite{gc17, n14} although \ital{p}-values describe how \ital{incompatible} a set of data is with a set of assumptions).
+Futher, p-values and Null-Hypothesis Significance Testing (NHST) have been criticized for fostering black or white thinking: either accept or reject the null hypothesis, leaving little room for uncertainty even when there is \parencite[]{cohen_earth_1994,aczel_quantifying_2017}.
+
+In short, misunderstandings of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work--and these can lead to misinterpretations that can diminish trust in science and the process through which scientific knowledge is constructed \parencite{vanderbleseffects2020,kreps_model_2020}.
+\\
+
+We next discuss the role of uncertainty not in science, but in science education contexts. \\
 
@@
 
-Given the above-noted roles of probability in scientific concepts and theories, it is unsurprising that educational researchers have documented how students learn about probability.
-As one example, probability and chance events play an important role in the life sciences \parencite{g03, gk08}, particularly in evolutionary processes and learning about them \parencite{th17}.
+Probability and chance events play an important role in the life sciences \parencite{g03, gk08}, particularly in learning about evolutionary processes \parencite{th17}, and scholars have explored how knowing about probability might impact knowledge about scientific ideas.
 Recently, \textcite{fsnh19} investigated the relation between statistical reasoning and acceptance and knowledge about evolution in a large sample of nearly 500 university students in the US.
 They found that students’ statistical reasoning capabilities are a strong predictor of both acceptance of evolution and knowledge about evolution.
-Research on teachers' conceptions of evolution has suggested that exposure to curricular materials that emphasized the role of randomness in evolution bolstered teachers' \ital{acceptance} of evolution, but not their understanding of it; the researchers conjectured that the difficulties learners face when working to understand probability and uncertainty may require extensive instruction \parencite{nadelson2010shifting}.
+Research on teachers' conceptions of evolution has suggested that exposure to curricular materials that emphasized the role of randomness in evolution bolstered teachers' \ital{acceptance} of evolution, but not their understanding of it.
+The authors conjectured that difficulties learners face when working to understand probability and uncertainty may require extensive instruction \parencite{nadelson2010shifting}.
+\\
 
-In the physical sciences, students' struggles with the probabilistic nature of quantum physics \parencite{br02} and nuclear decay \parencite{smmv17} have been established, but the underlying reasons and mechanisms remain less researched than in evolution.
-Research into students’ misconceptions inspired by conceptual change research has started to address this gap \parencite{ms15, alra14, st09}.
+In the physical sciences, students' struggles with the probabilistic nature of quantum physics \parencite{br02} and nuclear decay \parencite{smmv17} have been established by prior research, but the underlying reasons and mechanisms remain less researched than in evolution.
+Research into students’ misconceptions has started to address this gap \parencite{ms15, alra14, st09}.
 Findings mirror the position of \textcite{bgv94}.
@@
 Thus, these studies affirm students' struggles with scientific concepts that are probabilistic but hint at an underlying issue of more foundational challenges with reasoning about probability beyond science education contexts.
+\\
 
-Independent of the topics of quantum physics and nuclear decay, probability may generally be difficult for people to understand.
-Past research has shown that the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners.
-The work of \textcite{tk74}, especially, points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit.
-A prominent example is the neglect of base rate information in the (timely) context of medical testing, that is, the probability of having a medical condition based on a positive test result is often hugely overestimated because information about the base rate of people affected from the condition is neglected \parencite{kahneman_thinking_2012}.
+Adding to this difficult state of affairs is that when scientists try to communicate their uncertainty, prior research has demonstrated that many people–learners and experts alike–struggle to interpret the statistics and derivations of statistics (such as standard errors/margins of error, confidence intervals, and p-values that establish statistical significance) and that are often used to make inferences from scientific data \parencite{gkv04, s07, tk74}.
+Similarly, difficulties in understanding the technical language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution or nuclear decay that are modeled in terms of probabilistic statements \parencite[e.g.][]{fth17}.
+\\
 
-Responding to and building upon Kahneman and Tversky's work, Gigerenzer and colleagues provided important insights into the underlying mechanisms for the occurrence of these biases.
+The work of \textcite{tk74}, especially, points to a range of biases that can be elicited by the language in which statements about probability and uncertainty are framed.
+A prominent example is the neglect of base rate information in the (timely) context of medical testing.
+The probability of having a medical condition based on a positive test result is often overestimated because information about the (base) rate of people affected by the condition is not considered \parencite{kahneman_thinking_2012}; this base rate can be considered a form of prior knowledge about how probable it is that a person has a disease before considering the result of the test.
+The neglect of base rate information means that diagnostic tests that return positive results may not indicate that the individual truly has the disease the test is intended to detect.
+This also manifests itself in the prosecutor's fallacy \parencite{thompson1987interpretation}.
+\\
+
+Responding to and building upon Kahneman and Tversky's work that suggests that people may not reason in a \ital{proper} (or correct) Bayesian manner, Gigerenzer and colleagues provided important insights into the underlying mechanisms for the occurrence of these biases.
 Their research suggests that the biases are at least partly the result of information about probability and uncertainty being provided in a way that is incompatible with the heuristics that people develop from their everyday experiences \parencite{gh95, jkg18}.
 Based upon this premise, \textcite{jkg18} were able to demonstrate that students judge probability significantly better when language is adopted that better aligns with the heuristics that most learners develop from their everyday experiences.
-In sum, it seems feasible to change representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that these support, rather than hinder, learning.
+In sum, it seems feasible to change representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that they support, rather than hinder, learning.
+\\
 
-\subsection{Uncertainty in the Context of Analyzing and Interpreting Data}
+Given that people struggle to understand probability, is the Bayesian account of reasoning under uncertainty a prescriptive rather than a descriptive account?
+We explore this question in the next section drawing on work from scholars studying children's causal reasoning under conditions of uncertainty.
+\\
 
-Recent work--primarily by developmental psychologists and scientists--has applied Bayesian ideas to a new context: how children make sense of and world \parencite{g12, tgk06, tgk06}.
-This notion that children reason about the world in a way that is analogous to Bayes Theorem is in some ways surprising because Bayesian methods are associated with highly advanced statistical analyses, such as deciphering the Enigma machine during World War II \parencite{mcgrayne2011theory}.
-However, the notion that children (and people) weigh evidence about the world in a Bayesian way also builds upon, refines, and in cases refutes aspects of much earlier work that \ital{also} documented the surprisingly complex ways that children could reason about the world \textcite{pi69}.
-Different from earlier work on the stages through which children develop, recent work using Bayesian models considers the differences in the reasoning ability of children and adults to be a matter of degree, rather than kind.
-In sum, children hold initial ideas and understandings that they change (or update) in ways that are concordant with how Bayes Theorem represents the updating of an initial belief in light of data (and how the Bayesian Learning Cycle described earlier allows individuals to integrate inductive and deductive reasoning).
+\subsection{Bayes' Theorem and Research on How Children Learn to Reason}
 
-In addition to developmental research that uses Bayes Theorem as an analog for how children make sense of data and information in light of their initial ideas and beliefs, other work bears a striking semblance to educational and developmental research that explicitly refers to and draws upon a Bayesian perspective.
-In particular, research on how learners’ prior understanding influence how they make sense of how likely different patterns in data are centers the same tension around interpreting data in light of initial beliefs that a Bayesian perspective highlights \parencite[e.g., ][]{kd88, mkm07, ssc07}.
-In addition, prior research on learners' perceptions of the plausibility of a scientific explanation in light of alternative explanations highlights the importance of the degrees of belief (and re-assessing these (probabilistic, rather than dichotomous) degrees of belief in light of new information or experiences) that learners hold \parencite[e.g., ][]{lombardi2013plausibility, lombardi2016plausibility}.
+In the past two decades, developmental scientists have used Bayes' Theorem to understand how children make sense of and reason about the world \parencite{g12, tgk06, tgk06, bonawitz_sticking_2019}.
+The notion that children (and people) weigh evidence about the world in a Bayesian way also builds upon, refines, and in some cases refutes aspects of much earlier work that documented the surprisingly complex ways that children could reason about the world \parencite{pi69}.
+Different from earlier work on the stages through which children develop (e.g., the work of Piaget and other developmental \ital{stage} theorists), recent work using Bayesian models considers the differences in the reasoning ability of children and adults to be a matter of degree, rather than one of kind.
+From contemporary views of development, children hold initial ideas and understandings that they change (or update) in ways that are concordant with how Bayes' Theorem represents the updating of an initial belief in light of data \parencite{g12}.
+\\
+
+In addition to developmental research that uses Bayes' Theorem as a way to understand how children make sense of data and information in light of their initial ideas and beliefs, other, related research is also relevant.
+Specifically, research that studies how learners’ prior understanding influences probability judgements of different patterns in data centers on the same tension around interpreting data in light of initial beliefs that a Bayesian perspective highlights \parencite[e.g., ][]{kd88, mkm07, ssc07}.
+For example, the work of \textcite{kd88} considers the scientific reasoning process in terms of the two “searches” people undertake, of their beliefs and hypotheses in their mind and data and evidence collected or analyzed as a part of some investigation.
+Scientific reasoning involves the coordination of these two “problem” spaces, but neither entirely outweighs the other, in a manner similar to the Bayesian process of updating initial beliefs in light of data.
+We know that people--even children--can and do update their beliefs in light of data, though the extent to which they do depends on many factors, including how much people \ital{could} know about the topic to begin with \parencite{mkk17}.
+\\
+
+In addition, prior research on learners' perceptions of the plausibility of scientific explanations highlights the importance of the degrees of belief that learners hold about a particular scientific explanation \parencite[e.g., ][]{lombardi2013plausibility, lombardi2016plausibility}.
 This work suggests the merit of a Bayesian perspective as a potential frame or theoretical account for (in these key instances) analyzing patterns in data and conceptual change; it also has other benefits for teaching and learning.
@@
 
-Because of the advances made my research using a Bayesian perspective to understand human \ital{development} \parencite{gt07, gw12}, one might believe that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education and developmental scientists have pointed to this opportunity.
-For instance, \textcite{g12} wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p.
+Because of the advances made by research using a Bayesian perspective to understand human \ital{development} \parencite{gt07, gw12}, one might conjecture that there has been a ready application of these ideas to education.
+For instance, \textcite{g12} wrote that the use of Bayesian methods to understand child development can serve as ``a scientific foundation for a long tradition on `inquiry-based science education''', and that adopting a Bayesian perspective ``could lead us to much more specific and scientifically supported proposals for education'' (p.
 1627).
-However, it has largely not been the case that “science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” \textcite[p.
-1627]{g12}, as some science education scholars have bemoaned \parencite{ls15}).
-Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas \parencite[e.g., ][]{pi69} that represent the greatest contribution of developmental ideas to education \textcite{ls07}.
+However, it has largely not been the case that ``science itself could help turn young children's natural curiosity and brilliance into better science teaching and learning'' \parencite[p.
+1627]{g12}, as some science education scholars have bemoaned \parencite{ls15}.
+Indeed, scholars have bemoaned that it is often obsolete applications of Piaget’s ideas \parencite[e.g., ][]{pi69} that represent the greatest contribution of developmental ideas to education \parencite{ls07}.
 \\
 
-At the same time that opportunities to connect developmental science with science education, scholars have emphasized how important students' prior knowledge and understandings are to their understanding of scientific concepts and processes \parencite[e.g., ][]{ssc07, mkm07, mkk17, ls04}--highlight a potentially powerful connection between Bayesian methods and the priorities of science educators and science education researchers.
-Developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in education--especially science education--to design instruction based around eliciting and understanding students’ ideas \parencite{gb16, hbavb20, wtbs12}.
-There is an opportunity to use Bayesian ideas as a framework for formalizing these ideas about how prior ideas affect students’ reasoning about data and how teachers can leverage students’ ideas to design instruction, but a Bayesian perspective has--with a few exceptions focusing on how Bayes Theorem can provide a model for the science practice of argumentation \parencite[e.g., ][]{n11, so12}--has been absent from research in science education.
-A Bayesian perspective may also have a role in building trust in science by representing uncertainty in a principled yet flexible way.
+At the same time that opportunities to connect developmental science with science education exist, scholars have emphasized how important students' prior knowledge and understandings are to their ability to update their understanding of scientific concepts and processes \parencite[e.g., ][]{ssc07, mkm07, mkk17, ls04}--highlighting a potentially powerful connection between Bayesian methods and the priorities of science educators.
+Developmental and psychological research into Bayesian models of cognition supports and bolsters these educational efforts--especially science education efforts--to design instruction based around eliciting and understanding students’ ideas \parencite{gb16, hbavb20, wtbs12}.
 \\
 
-\section{A Primer on Bayes Theorem and a Bayesian Epistemology}
+Taken together, there is an opportunity to use Bayes' Theorem and Bayesian ideas to bring together and to formalize a collection of ideas that are relevant to science teaching and learning.
+However, a Bayesian perspective--with a few exceptions \parencite[e.g., ][]{n11, so12}--has been mostly absent from research in science education.
+Moreover, a Bayesian perspective may also have a role in building trust in science by representing uncertainty in a principled yet flexible way--in this way, having a potential impact on people's scientific literacy.\\
 
-\subsection{An Accessible Introduction to Bayes Theorem}
+Before elaborating on the education- and science literacy-related roles for a Bayesian perspective, we introduce Bayes' Theorem from an accessible mathematical and statistical perspective (with connections to the axioms that follow from this perspective).
 
-At its core, Bayes' theorem provides a means to update one's beliefs about the world from observations in such a way that prior knowledge can be used in a principled manner.
-As represented in Equation 1, Bayes' Theorem states that at any point in time, our knowledge about the world after making a new observation (i.e., our posterior knowledge) can be obtained by multiplying our prior knowledge with a predictive updating factor (e.g., \cite{RouderMorey2019,WagenmakersEtAl2016CD} ):
+\section{A Primer on Bayes' Theorem}
+
+Fundamentally, Bayes' Theorem is a mathematical procedure to understand new evidence in light of prior information.
+In this way, the process of updating what is already known is \ital{mathematical}, but Bayes' Theorem also includes an \ital{epistemological} component--a component related to what is known about the world.
+What counts as prior information should be construed very broadly--it encompasses subjective judgments of how likely an event is as well as knowledge of empirical data from other, related events.
+\\
+
+The usefulness of Bayes' Theorem lies in how it presents a flexible yet principled way to update what is known in light of the evidence.
+Bayes' Theorem is not a cure-all; it is, instead, a formalization of something scientists and people alike do: interpret evidence in light of what is already known.
+It is epistemologically \ital{normative} to the extent that it is the derivation of the laws of conditional probability.
+\\
+
+As represented in Equation 1, Bayes' Theorem states that at any point in time, our knowledge about the world after observing new evidence can be obtained by multiplying our prior information with something called a \emph{predictive updating factor} \parencite{RouderMorey2019,WagenmakersEtAl2016CD}:
+
 \begin{equation*}
@@
 \end{equation*}
-Mathematically, this can be expressed as  
+
+Mathematically, this can be expressed as:  
 \begin{equation}
@@
 \end{equation}
-where $\theta$ is a hypothesis, proposition, claim, parameter value, or in general any particular account of the world about which we are uncertain.
+where $\theta$ is prior information that can take the form of a  proposition, claim, hypothesis, parameter value, and any account of the world about which we are uncertain.
 The predictive updating factor consists of two components.
-First, $p(\text{data} \given \theta)$ is the likelihood, that is, the extent to which the observed data were expected under $\theta$.
-Second, $p(\text{data})$ is the marginal likelihood, that is, the extent to which the observed data were expected without conditioning on $\theta$.
+First, $p(\text{data} \given \theta)$ is the \ital{likelihood}, that is, the extent to which the observed data were expected under $\theta$.
+Second, $p(\text{data})$ is the \ital{marginal likelihood}, that is, the extent to which the observed data were expected without conditioning on $\theta$.
 If the observed data are more expected (i.e., better predicted) when conditioning on $\theta$, the predictive updating factor is higher than 1, and $\theta$ becomes more credible.
 
-To illustrate the above equation, consider the following: An educator is interested in the proportion of students that use Windows or MacOS as their operating system.
-The proportion of Windows users $\theta$ ranges between 0 and 1.
-Before you start our Bayesian analysis, you need to quantify your background knowledge -- the relative plausibility of the different values for $\theta$.
-Given their previous experiences as a teacher, their best guess is that about half of the students use Windows.
-They would not be surprised if the proportion of Windows users was between 30\% and 70\% but you would be rather surprised to find either no Windows or exclusive Windows usage.
+To illustrate the above equation, consider the following.
+The Hemlock Wooly Adelgid is an invasive insect that feeds on Eastern Hemlock trees, a tall-growing pine tree found in the Eastern United States.
+Because infestations can kill Eastern Hemlocks fairly rapidly (within several years), in many affected areas, people take steps to protect Eastern Hemlocks that are (or could be) affected by the Hemlock Wooly Adelgid; some systemic treatments can be effective \parencite{nps2021}.
+\\
 
-This background knowledge motivates the use of the dome-shaped prior distribution shown as the dashed line in Figure~\ref{fig:WindowsMacPriorPosterior}.\footnote{This is the beta$(2,2)$ distribution.} Now suppose you start to collect data by asking students about the operating systems of their computers.
-The first ten observations of Windows (W) and MacOS (M) usage that you make are $\{W, W, M, W, M, W, W, M, W, M\}$: six students use Windows and four use MacOS.
+A student (or a scientist) may be interested in the proportion of Eastern Hemlock trees within a specific area that exhibit signs of infection: diseased trees with visible white ``cotton ball'' clumps at the base of the needles of affected trees.
+Let's consider that in the location we are examining, each Eastern Hemlock tree grows about the same distance apart from every other tree (such that we may reasonably consider the distance between trees to not be a critical factor to consider in an investigation).
+Before beginning an analysis, we need to establish what prior information we have - the relative plausibility of the different values for the proportion of affected Eastern Hemlock trees, or $\theta$.
+Given what they learned in class and what they noticed walking into their school, the student's best guess may be that about half of the Eastern Hemlock trees on the playground of their school are infected.
+They would not be surprised if the proportion of affected trees was between 30\% and 70\%, but--based on what they learned and observed--they would be rather surprised to find either no affected trees or that all of the trees were affected.
+\\
 
-With these data in hand, we then combine our prior knowledge (i.e., the dome-shaped prior distribution) with the predictive updating factor (assuming a binomial likelihood) and obtain the posterior distribution shown as the solid line in Figure~\ref{fig:WindowsMacPriorPosterior}.
+This background knowledge motivates the use of the dome-shaped prior distribution shown as the dashed line in Figure~\ref{fig:WindowsMacPriorPosterior}.\footnote{This is the beta$(2,2)$ distribution.} Now suppose the student starts to collect data by observing trees on the playground of their school.
+The first ten observations (with \ital{Y} representing infected and \ital{N} representing not infected) are $\{Y, Y, N, Y, N, Y, Y, N, Y, N\}$: six infected trees and four not infected trees.
+\\
+
+With these data in hand, we then combine our prior knowledge with the predictive updating factor and obtain the posterior distribution shown as the solid line in Figure~\ref{fig:WindowsMacPriorPosterior}.
 The posterior distribution is more peaked than the prior distribution, indicating that the data have sharpened our knowledge about $\theta$.
 In addition, the posterior distribution is higher than the prior distribution for values of $\theta$ between approximately $.4$ and $.75$: these values of $\theta$ predicted the data relatively well and have therefore gained credibility.
-In contrast, values of $\theta$ lower than $.4$ and higher than $.75$ predicted the data relatively poorly, which is why they have lost credibility compared to the prior.\footnote{Note that the area under a probability distribution equals 1, so that posterior gains for specific regions of $\theta$ cancel out exactly against posterior losses for the other regions.}
+In contrast, values of $\theta$ lower than $.4$ and higher than $.75$ predicted the data relatively poorly, which is why they have lost credibility compared to the prior.\\
 
@@
 \includegraphics[width = .785\paperwidth]{WindowsMacPriorPosterior.pdf}
-\caption{Simple example of Bayesian learning.
-A dome-shaped prior distribution captures the background knowledge concerning the proportion of students using Windows vs.
-MacOS.
-Observing fictitious data from 10 students (six using Windows, four using MacOS) drives a knowledge update that results in a bell-shaped posterior distribution.
-Figure based on the \emph{Learn Bayes} module in JASP.}
+\caption{Example of the use of Bayes' Theorem to update prior information in light of new evidence.
+A dome-shaped prior distribution captures the background knowledge concerning the proportion of affected trees.
+Observing ten trees (six affected and four not affected) drives a knowledge update that results in a bell-shaped posterior distribution.
+Figure based on the \emph{Learn Bayes'} module in JASP.}
 \label{fig:WindowsMacPriorPosterior}
@@
 
-Although we have demonstrated this updating process in a single step, it could also be executed sequentially, one student at a time.
-For instance, the first observation you make is a $W$, and in light of your prior, this makes Windows usage a little more likely than MacOS usage.
+Although we have demonstrated this updating process in a single step, it could also be executed sequentially, one tree at a time.
+For instance, the first observation we make is a $Y$, and in light of the prior, this makes the proportion of infected trees a bit higher.
 This slight change in knowledge is reflected in the difference between the distributions on the top two lines in Figure~\ref{fig:WindowsMacSequential}; line ``0'' represents the dome-shaped prior distribution and line ``1'' represents the posterior distribution after the first observations.
-Note that observing a $W$ has nudged the distribution a little to the right.
+Note that observing a $Y$ has nudged the distribution a little to the right.
+\\
+
 Crucially, the `nudged' posterior distribution after the first observation now takes the role of the prior distribution ready to be updated by our second observation.
-The second observation is again a $W$, and line ``2'' shows that the resulting posterior distribution is nudged to the right once more.
+The second observation is again a $Y$, and line ``2'' shows that the resulting posterior distribution is nudged to the right once more.
 At every stage in the sequential updating process, the posterior distribution based on the observations seen so far becomes the prior distribution for incorporating the information from the next observation.
-Specifically, the changing distributions in Figure~\ref{fig:WindowsMacSequential} show our knowledge about the world is subject to constant change, by making observations and integrating incoming information with current knowledge using Bayes' theorem.
+Specifically, the changing distributions in Figure~\ref{fig:WindowsMacSequential} show our knowledge about the world is subject to constant change, by making observations and integrating incoming information with current knowledge using Bayes' Theorem.
 It should be emphasized that although the updating process is conceptually different, the final outcome is identical: the shape of the posterior distribution is unaffected by whether data arrive simultaneously or sequentially.
@@
 \includegraphics[width = .65\paperwidth]{WindowsMacSequential.pdf}
-\caption{Sequential Bayesian learning.
-A dome-shaped prior distribution (line ``0'') captures the background knowledge concerning the proportion of students using Windows vs.
-MacOS.
+\caption{Updating what is known sequentially using Bayes' Theorem.
+A dome-shaped prior distribution (line ``0'') captures the background knowledge concerning the proportion of infected trees.
 Each new observation results in an update to a posterior distribution, which then becomes the prior distribution for the analysis of the next observation.
-Figure based on the \emph{Learn Bayes} module in JASP.}
+Figure based on the \emph{Learn Bayes'} module in JASP.}
 \label{fig:WindowsMacSequential}
@@
 
+We expand on this primer--specifically mathematical axioms that follow from Bayes' Theorem--in Appendix A.
+
+\section{Supporting Science Learners to Navigate Uncertainty}
+
+\subsection{The Principles That Follow From Bayes' Theorem}
+
+As long as science instruction progresses mostly \ital{qualitatively}--our assumption for most K-12/pre-collegiate science classrooms--the features of Bayesian reasoning can help to bring a sense of coherence to the many ways in which uncertainty is manifest when students engage in scientific practices such as constructing explanations, analyzing and interpreting data, or developing and using models.
+\\
+
+Thus, we argue that Bayesian reasoning can provide this coherence as its distinctive feature, through the use of ``concept of degree of belief to describe epistemic attitudes about uncertain propositions'' \parencite[, p.
+2]{sprenger2019bayesian}.
+Such a perspective reflects an embrace of the epistemic stance that scientific knowledge is not immutable but in fact open to revision.
+In other words, scientific knowledge always comes attached with a degree of belief that can be updated following the rules of Bayes'’ Theorem when new evidence becomes available.
+The result of that updating process depends on the uncertainty of existing scientific knowledge which is then expressed in the prior.
+\\
+
+In this way, Bayes' Theorem applied to the scientific process, or what we refer to as \ital{Bayesian reasoning}, emphasizes that science and scientific knowledge is always situated in context \parencite{so12} and society \parencite{driver1994constructing}.
+\\
+
+These ideas about what Bayesian reasoning is can be captured in the following three principles:
+
+\begin{enumerate}
+    \item \ital{Be open to new evidence}: Scientific knowledge always comes with some uncertainty and priors that denoting absolute certainty or impossibility --by definition--prevent scientific progress.
+This principle expresses the Bayesian degree of belief epistemology, exemplifying that scientific knowledge is tentative and that scientists should not make absolute or certain claims.
+Lindley popularized this idea in statistics and coined it ``Cromwell's rule'' \parencite[p.
+104]{Lindley1985}.
+    \item \ital{Account for what is already known}: Evaluate new evidence in light of prior information.
+This emphasizes how scientific knowledge is not constructed in isolation but rather is built upon earlier information and evidence.
+    \item \ital{Consider alternative explanations}: Consider the evidence in terms of compatibility with all possible outcomes; in other words, consider counterfactuals.
+This expresses what in Bayesian philosophy is referred to as the \ital{Simple Principle of Conditionalization} \parencite{adams1965logic}, i.e., when we weigh evidence we must consider to what extent it supports the range of possible explanations for the data.
+\end{enumerate}
+
+Emphasizing these principles and their relations to the nature of scientific knowledge within the context of science teaching and learning can support students in connecting the frequent but often isolated references to uncertainty and limits of scientific knowledge in the Framework for K-12 Science Education \parencite{nrc12} and the NGSS \parencite{ngss2012}.
+The descriptions of the science practices in the \ital{Framework} \parencite{nrc12}, particularly, contain expressions such as revising models, refining explanations, critiquing arguments, and considering the limitations of the precision of the data.
+\\
+
+However, what is lacking in these statements is an explicit and coherent rationale for why revising, refining, considering limitations is a necessary part--and in fact a feature--of science, rather than a limitation or constraint of science.
+Familiarizing students with the principles as core components of Bayesian reasoning can foster ideas about scientific knowledge--epistemological ideas--from which the elements of the practices that pertain to uncertainty and the limits of scientific knowledge follow naturally.
+Moreover, this assumption is supported by the work of \parencite{warren_impact_2020,warren_quantitative_2018} which demonstrates that integrating Bayes' reasoning into university science courses positively shifts students epistemic attitudes regarding the nature of knowing and learning.
+This assumption is also supported by research at the upper elementary (ages 10-11) grade levels \parencite{kazak2015bayesian, kazak2018emergent}.
+\\
+
+In the following section, we provide an example of how the quantitative Bayesian updating activities in \textcite{warren_impact_2020,warren_quantitative_2018} can be made accessible to students in middle and high school by using a computational tool--specifically tailored toward students at the middle and high school grade levels (ages 12 and older).
+In both studies, Warren implemented Bayesian updating activities into a introductory university physics courses.
+The activities were added to in-class and homework tasks and generally asked students to evaluate their answers or results using Bayesian Reasoning.
+For example, in addition to all the usual parts of a lab report students were asked record their initial confidence in the hypothesis they would test, asked to estimate how the data they collected aligns with their initial hypothesis, and update their confidence accordingly using Bayes' theorem.
+In a quasi-experimental setting,  \textcite{warren_impact_2020} found that these activities positively impact students' epistemic beliefs, including beliefs regarding the nature of scientific knowledge and the presence and importance of uncertainties.
+\\
+
+\subsection{Making it Practical: An Example with the Confidence Updater Widget}
+
+\subsection{Background on the Confidence Updater Widget}
+
+\textcite{warren_impact_2020,warren_quantitative_2018} frames the Bayesian reasoning as part of a hypothetico-deductive process \parencite[][pp.
+30, 360]{popper_objective_1979}.
+In this process, a model is evaluated by deriving a testable hypothesis and testing it against yet unknown data.
+Depending on the ratio between the likelihood of observing the data if the hypothesis is true and the likelihood of observing the data if the hypothesis is false, confidence in the model decreases or increases relative to the initial confidence in the model.
+\\
+
+This confidence updating can be described stating Bayes' Theorem in the following way where $\theta$ from equation (1) takes the form of a hypothesis H: 
+\begin{equation}
+p(H|data) = \frac{p(H)\times R}{p(H)\times R+1-p(H)}
+\end{equation}
+
+p(H|data), or our \ital{posterior}, is the confidence we can have in the hypothesis H after updating our initial confidence, or our \ital{prior}, $p(H)$ based on consideration of the \ital{new evidence}, expressed in the updating factor \ital{R}\footnote{Technically a Bayes Factor.} where \\
+
+\begin{equation}
+R = \frac{p(data|H)}{p(data|\neg H)}
+\end{equation}
+
+\ital{R} > 1 represents confirmatory evidence, while \ital{R} = 1 represents inconclusive evidence, and \ital{R} < 1 represents disconfirmatory evidence.
+Guidelines for choosing an adequate R exist \parencite{kass1995bayes}, e.g., 20 < \ital{R} < 150 can be interpreted as the evidence strongly favoring $H$.
+$p(H)$ the initial confidence in the hypothesis, ranges from 0 to 1 with $p(H)$ = 0.5 representing maximum uncertainty, i.e., having no idea about the validity of the hypothesis.
+\\
+
+To support, students focus on the conceptual elements of updating their confidence in a hypothesis, the Shiny--a framework available through the \ital{R} language \parencite{rlang} for creating interactive web applications--Confidence Updater app represented in Figure~\ref{fig:confidence-updater-input}.
+This app allows students to choose values for the strength of the new evidence--\ital{R}--based on their interpretation of the evidence and p(H) based on their initial confidence and after performing the needed calculations returns a textual statement about updated confidence $p(H|E)$ in the hypothesis after considering the evidence and an optional numeric value for the confidence level as well.
+\\
+
+\begin{figure*}[h]
+\begin{center}
+\includegraphics[width = .785\paperwidth]{confidence-updater-input.png}
+\caption{Shiny web application for updating one’s confidence in a hypothesis following Bayes' Theorem.
+$P(H)$ corresponds to the second question (How sure are you about your hypothesis?); \ital{R} corresponds to the second question (How compatible is the evidence with your hypothesis relative to an alternative hypothesis?).}
+\label{fig:confidence-updater-input}
+\end{center}
+\end{figure*}
+
+Using the app, one can obtain an estimate--based on Bayes' Theorem--for the confidence one can have in a hypothesis based upon the degree of one's belief in the initial hypothesis--one's prior--$p(H)$, as well as the predictive updating power of the new evidence, \ital{R}.
+One can see the example output in Figure~\ref{fig:confidence-updater-output}.
+
+\begin{figure*}[h]
+\begin{center}
+\includegraphics[width = .785\paperwidth]{confidence-updater-output.png}
+\caption{Output from the Shiny app}
+\label{fig:confidence-updater-output}
+\end{center}
+\end{figure*}
+
+The principles from the previous section are reflected implicitly and explicitly in the app.
+
+The first principle, \ital{Be open to new evidence}, is emphasized as a natural consequence of Bayes' Theorem and the options \ital{absolutely certain that it is correct} and \ital{absolutely certain that it is incorrect} in the "How sure are you about your hypothesis?" section of the confidence updater.
+\\ When students select from among these options, $p(H)$ is set to 1 (absolutely certain that it is correct) or 0 (absolutely certain that it is incorrect) respectively.
+In such cases, neither the tentative (or changeable) nature of scientific knowledge nor the empirical basis of science are manifest, though both are key elements of science \parencite{abd1998nature}.
+
+In consequence, whatever updating factor \ital{R} students choose, the updated confidence will always be the same as the initial hypothesis--the prior--$p(H)$ and the consideration of the new evidence becomes pointless.
+In other words, in such cases, no amount of data may change one's belief.
+From this perspective, a question could be considered to be not scientific, as empirical evidence has no bearing on the answer to the question.
+\\
+
+If students selected any of the options that are equivalent to $p(H)$ = 1, then observe strong contrary evidence, and wonder how this does not affect the updated confidence are observing the evidence, a teacher could refer to principle a) and point out to students how this degree of belief can inhibit changes in beliefs even when confronted with strong evidence--as the students may have just experienced when using the app.
+\\
+
+As \textcite{warren_quantitative_2018} points out, ``the fact that we can never reach a probability of zero is the quantitative expression of the maxim that we should never rule out any hypothesis with an absolute certainty'' (p.
+371).
+At the same time, even ``in cases where a hypothesis is confirmed by multiple experiments, the posterior probability will approach but never reach, a value of 1; this expresses the maxim that one ought never to have absolute confidence in the truth of a hypothesis.'' (p.
+371).
+In other words, there are no certainties where it comes to Bayesian epistemology.
+This reflects philosophical and historical accounts of science that emphasize its mutability over time \parencite{Kuhn1962} as expressed in principle 1).
+\\
+
+The second principle, \itel{account for what is already known} appears in the wording of the prompt for selecting $p(H)$ and becomes visible to students as it effectively moderates the power of the evidence.
+When prior beliefs and evidence align, the updated confidence is larger in comparison to either having no idea” before or having prior beliefs that do not align.
+Students could notice this if they compare the updated confidence $p(H|E)$ for students that had the same prior belief but evaluated the evidence differently or had different prior beliefs but evaluated the evidence in the same way.
+Again, as \textcite{warren_quantitative_2018} points out, “the fact that everyone starts with subjective prior probabilities and yet inevitably converges to a single asymptotic result illustrates the objective aspect of science” (p.
+371), which can help build trust in science.
+\\
+
+Finally, the last principle, \ital{consider alternative explanations} is reflected in how the prompt for \ital{R} is worded.
+Further, when students use the confidence updater, students should be encouraged to argue for their choice of \ital{R} and be able to explain why they choose a specific option.
+In such an argument, students following the prompt should not only consider how their evidence supports their own hypothesis but also how compatible it is with alternative hypotheses.
+That students must select the value for \ital{R}, the predictive updating factor, is notable.
+Students must figure out the extent to which any single laboratory activity contributes to what we collectively know about a scientific phenomenon: As \textcite{warren_quantitative_2018} writes, it leads to a more realistic view of what laboratory activities can accomplish (i.e., few single introductory physics experiments are likely to overturn established knowledge about the physical world, but they can lead to us updating our understanding)--as well as one that emphasizes scientific sense-making to a greater extent.
+\\
+
+\subsection{Using the Confidence Updater app}
+
+How does using the confidence updater change a typical science class activity?
+Here we sketch how using the confidence updater can provide rich opportunities to discuss and reflect on epistemic attitudes.
+A standard activity at the middle grades levels is to work to understand the relations between amperage, \ital{I}, resistance, \ital{RE}\footnote{We use \ital{RE} for electric resistance--instead of the traditional \ital{R}--to avoid confusion with the updating factor \ital{R}}, and voltage \ital{V}, i.e., \ital{Ohm’s law}, $V=RE*I$.
+\\
+
+To explore the relations reflected in Ohm's law, students can measure the amperage in a circuit at different voltages, keeping the resistance fixed, or constant.
+At the beginning of the activity, students can be asked to generate different hypotheses for the relationship between amperage and voltage.
+In our experience, students often propose a proportional (or linear), inverse, or quadratic relationship for this relationship--when resistance is held constant.
+Based upon their initial ideas--that the relationship is proportional, inverse, or quadratic, they may test their hypotheses using the app.
+\\
+
+Specifically, following the second principle, students can now specify their prior confidence in their group’s hypothesis.
+Students may very well choose different priors based on their individual experiences.
+Now, they may proceed and record a number of measurements of the current for different voltages and then graph the data.
+\\
+
+Figure~\ref{fig:example-student-graph} shows an example graph. \\
+
+\begin{figure*}[h]
+\begin{center}
+\includegraphics[width = .5\paperwidth]{ohm-graph.png}
+\caption{An example student graph}
+\label{fig:example-student-graph}
+\end{center}
+\end{figure*}
+
+When students are asked to evaluate the Evidence and set a value for \ital{R}, they are primed to consider to what extent the evidence is consistent with their hypothesis relative to an alternative hypothesis, reflecting principle 3.
+Looking at the graph one could see evidence for a proportional relationship as well as a higher-order positive relationship (blue and red dashed line in Figure~\ref{fig:example-student-graph}).
+\\
+
+Thus, the evidence is barely compatible with an inverse relationship between I and U but supportive of both hypotheses that suggest a positive relationship between the variables.
+Thus, depending on what hypothesis students choose to investigate, they should select “data favors an alternative hypothesis” for R if their hypothesis is an inverse relationship and select “data somewhat favors my hypothesis” if their hypothesis is proportional or quadratic relationship.
+\\
+
+The updated confidence for every student will depend on their initial confidence--the prior ($p(H)$)--in their hypothesis and the evaluation of the data which will be different across groups.
+If students compare their results within groups, they should see how the effect of their initial confidence influences the effect of the evidence.
+However, if they repeat the activity, i.e., use their updated confidence as a new initial belief and collect more data, they should see how consensus is approached over a few iterations unless anyone choose $p(H) = 0$ or $p(H) = 1$, which points to the importance of principle 1).\\
+
+\subsection{How Can Younger Students Update Their Beliefs in a Bayesian Way?}
+
+It is worth briefly considering whether--and if so, how--younger children may engage in a Bayesian activity like that described above.
+The work of \textcite{kazak2015bayesian, kazak2018emergent} demonstrates how Bayesian activities can be designed for younger students.
+For a study with fifth grade (10-11-year-old) students, \textcite{kazak2015bayesian} created a game in which students drew one coin/token from each of two bags.
+If the colors of the two tokens matched, then students won.
+The question students were tasked to answer was, "Is the game fair or not?".
+They played four different rounds of the game (using four different sets of bags).
+\\
+
+Prior to playing the bag, through a worksheet, students a) articulated whether or not the game was fair and b) their confidence in their beliefs about the fairness of the game.
+Three of the four games were designed to make the task of determining their fairness challenging.
+Over the course of playing the games--and using a statistical software tool for learners, TinkerPlots \parencite{konold2005tinkerplots}--students updated both their beliefs about the fairness of the game and their confidence shifted in conjunction with their analysis of data: both generally improved as students collected more data and simulated the collection of data through TinkerPlots.
+\\
+
+Students first expressed their initial beliefs about the fairness of the game before playing it, updating these beliefs over time.
+As \textcite{kazak2015bayesian} explained, "since these beliefs can change based on new evidence, it is important to assess the personal degree of confidence in the initial hypothesis or prediction and look at how it changes over the course of gathering new relevant information" (p.
+704).
+This work shows that students can express their \ital{subjective probability} beliefs \parencite[]{batanero2016research,, konold1991understanding} and--critically--update them over time.
+An instructional implication of this research is that educators can elicit students' ideas about what they think about a phenomenon prior to conducting an investigation (and collecting data).
+This is an instructional practice already familiar to science educators and science teacher educators \parencite{windschitl2018mbitious, windschitl2012proposing}.
+\\
+
+In addition to eliciting students' ideas, though, educators can also prompt students to consider how confident they are in their beliefs or ideas.
+Reflecting upon and "updating" these beliefs can be supported over the course of investigations (and data analyses) by prompting students to consider both their beliefs and confidence over the course of an activity, which can support students to begin to understand the three principles that characterize Bayesian reasoning.
+\\
+
+\section{Toward a Bayesian Research and Development Agenda in Science Education}
+
+In the last section, we provided an example of how Bayesian principles can be made accessible to a large part of the K-12 science classroom when we use tools like the Confidence Updater app or approaches that prompt students to consider their beliefs and their confidence in them and how they can be modified over time.
+Based on findings with older students, this work has the potential to support students' epistemic understandings about science \parencite{warren2018quantitative, warren2020impact}.
+We also have findings from past classroom-based research that "modifying predictions based on previous results appeared to be intuitive for young children" \parencite[, p.
+53]{kazak2018emergent}.
+In this way, beliefs  about the nature of knowing and knowledge-creation could in turn help to build trust in science.
+But, there are other ways in which a Bayesian perspective could support reasoning about uncertainty and probability more generally.
+There are several avenues through which this can be approached.
+\\
+
+First, research on the role of Bayes' in education has mostly been carried out in the context of undergraduate statistics education, in which there have been reviews \parencite{dogucu2021}, recommendations and calls to action \parencite{gpkpwc18, h_a20, dogucu2021}, examples, and design options \parencite{a02, b02,g08, w17, h_j20} and debate \parencite{jrhrr20} over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs.
+Why has there been this attention?
+The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power \parencite{gpkpwc18} as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers \parencite{ah20}.
+As evidenced by the recent special issue of the \emph{Journal of Statistics Education} of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.
+\\
+
+Despite attention from statistics educators and statistics education researchers, there has been little research outside of statistics education in the broader science education community.
+Two publications merit mention, though.
+The work of \textcite{so12} and \textcite{n11} both applied Bayesian perspectives to the science practice of argumentation.
+Szu and Osborne presented the case for how and why Bayes' Theorem can apply to research and practice on scientific reasoning, broadly, and argumentation, particularly.
+This paper represents the most comprehensive account of the relevance of Bayesian methods for science education.
+Szu and Osborne argue that Bayes' is useful as both a formal mathematical tool and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes' Theorem that follow more or less deterministically from ``external, objectively probabilistic systems'' (p.
+61)—is ``the key leap that characterizes the debate about the value of Bayesian inference as a model of scientific reasoning'' (p.
+61).
+In this way, Szu and Osborne argue that the greatest use of Bayes' Theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community \parencite{batanero2016research, mr18}.
+They offer some research-related backing for the use of Bayes' Theorem (e.g., noting how conceptual change research, particularly, and constructivist research, broadly, can both be explained coherently within a Bayesian approach) as well as some practical, instructional recommendations, some of which we detail later in this section.
+\\
+
+\textcite{n11} work on Bayesian approaches to argumentation describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation.
+Concerning the latter, Nussbaum describes how the social issue of raising taxes to provide resources to homeless individuals was a rich context for students to engage in forms of Bayesian reasoning.
+In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what \textcite{so12} characterize as the more externally objective use of Bayes' Theorem—and how the estimates that result from applying Bayes' Theorem led students to re-evaluate their initial arguments.
+\\
+
+The work of \textcite{so12} and \textcite{n11} begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process.
+Neither, though, speaks to how a Bayesian perspective could have a role in bolstering understanding or trust in science by learners and others.
+More generally, these two contributions are steps toward conceiving what a Bayesian perspective may offer science education, but there remains room to build further.
+For instance, despite their association with statistical methods that are growing in use in post-secondary contexts and in industry \parencite{mcgrayne2011theory}, no science education scholarship has applied Bayesian methods to another science practice, that of analyzing and interpreting data.
+In this way, being clear about what we mean by a \ital{Bayesian perspective} is important: this perspective could be associated not only with a \ital{practice} (analyzing data) but also as a cross-cutting concept \parencite{nrc12} that can, in the context of learning about particular scientific ideas, unite the practice of argumentation with analyzing and interpreting data with the notions of probability and (un)certainty.
+Bayesian methods can also provide a framework that includes a set of epistemological principles and a partial explanation for the process of how and why people change their ideas in light of new information or data--two key components of (and potential barriers to) how members of the public understand and act upon scientific knowledge \parencite{sinatra2014addressing}.
+
+\section{Conclusion}
+
+In light of how scientific knowledge can be presented to learners as immutable \parencite{d90, manz2018supporting, carey1993understanding}, it may be unsurprising that learners--and people--are dismayed when scientific knowledge that was once considered to be the consensus is modified or even nullified.
+When learners do not have a way to think about uncertainty, they may adopt an approach that considers scientific knowledge to be correct or incorrect--or for scientific sources to provide evidence that is true or false.
+From a Bayesian perspective, uncertainty about scientific knowledge can be viewed in terms of the degrees of belief that a person holds toward a belief, claim, or research--any prior information--in light of new evidence.
+We conjecture in this paper that such a paper can bolster individuals' trust in science by providing a principled, flexible way to reason about uncertainty.
+
+Indeed, a key potential benefit of Bayes' may be that it provides a framework for helping students to transition from their conceptual understanding of scientific ideas to quantitatively analyzing data.
+In this way, we argue for the use of a Bayesian perspective that could influence research about (and teaching to support) learners' epistemic considerations and how learners' ideas change over time--factors that may have a bearing on how they understand science outside of school, or later as adults \parencite{sinatra2014addressing}.
+Thus, we see there is a number of ways in which a Bayesian perspective can influence science education--and, therefore, we see  a number of mechanisms through which Bayesian methods could bolster trust in science, from specific teaching approaches, such as those that could be constructed around the Confidence Updater widget \parencite[]{warren_impact_2020, warren_quantitative_2018}, to approaches tailored to younger children \parencite[e.g., ]{kazak2015bayesian, kazak2018emergent}.
+Somewhat different from the past research focused on the connection between scientific argumentation and Bayes' Theorem \parencite[]{n11, so12}, our aim with this paper was to highlight the broad potential of a Bayesian perspective for reasoning about uncertainty in data and evidence and to offer a few directions for how the potential of Bayes' in future research.
+\\
+
+Even at a time in which science has made and continues to make great advances, the history of science suggests that uncertainty will not be removed (or even reduced) given scientific progress \parencite{fara2010science}; looking ahead, being able to effectively navigate, counsel, and act in the midst of uncertainty may be as important as it has been at other uncertain points in history.
+Like the surprising utility of a Bayesian perspective in domains that are similar to education (e.g., developmental science), might an unexpected feature of a Bayesian perspective be that embracing uncertainty--equipped with the tools that Bayesian methods provide--makes individuals and societies more confident?
+In short, can we build trust in science by embracing uncertainty?
+We think that these are questions that are \ital{likely} to yield positive results for science educators and those advocating for a more scientifically-informed citizenry and science learners.
+\\
+
+\printbibliography
+
+\newpage
+
+\section{Appendix A: The Epistemological Axioms that Follow from Bayes' Theorem}
+
 \subsection{Integrating Inductive and Deductive Reasoning With the Bayesian Learning Cycle}
 
-A recent case from the physical sciences can serve to illustrate how the principles from the example in the previous section apply to science and science education.
-In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light \parencite{b11} - and thus defying Einstein’s theory of relativity.
+Several years ago, physicists reported that they had observed elementary particles moving faster than the speed of light \parencite{b11}--thus defying Einstein’s theory of relativity.
 However, instead of celebrating a historic discovery, the scientists started to look for what had gone wrong in their observations.
@@
 
-Thus, rather than changing their belief about the speed of light, scientists looked for things that could have gone wrong with the measurement.
-This example shows how a qualitative take on Bayes’ theorem can explain and predict many aspects of how animals \parencite{o13}, people \parencite{tgk06}, and organizations \parencite{kaj12} learn from observations.
+Rather than changing their belief about the speed of light, scientists looked for things that could have gone wrong with the measurement.
+This example shows how a qualitative take on Bayes'’ theorem can explain and predict many aspects of how animals \parencite{o13}, people \parencite{tgk06}, and organizations \parencite{kaj12} learn from observations.
 Recently, advances in computational power and techniques have started the more and more widespread application of this principle in a quantitative way that allows researchers to precisely express their knowledge about the world and then learn from observations.
+\\
+
 Expressed in the Bayesian learning cycle (Figure \ref{fig:The Bayesian Learning Cycle}), a Bayesian reasoning framework allows scientists to integrate the strengths of both, inductive and deductive research approaches, in a formalized way.
 Further, when specifying the prior, scientists openly and explicitly need to describe the uncertainty in the current state of knowledge, and the posterior captures how the data has changed the knowledge and the uncertainty in that knowledge.
-Thus, Bayesian reasoning is, at least in principle, open and transparent by design.
+Thus, Bayesian reasoning is--in principle--open and transparent by design.
+\\
+
 \begin{figure*}[h]
@@
 \end{figure*}
-Having introduced Bayes Theorem and an application of Bayes Theorem as a cycle through which individuals can learn, we show how Bayes Theorem can be used to demonstrate several principles that are relevant to building trust in science.
 
-\section{The Epistemological Principles that Follow from Bayes Theorem}
+\subsection{Mathematical Axioms That Follow From Bayes' Theorem}
 
@@
 We illustrate these to later show how they can have a role in the science classroom and how they can be used by individuals outside of classrooms and schools to understand and act upon scientific information.
+\\
 
@@
 Thus, Bayes' rule states that our posterior knowledge $p(\theta \given \text{data})$ is proportional to the likelihood $p(\text{data} \given \theta)$ (or the extent to which the observed data are expected given $\theta$) multiplied by our prior knowledge $p(\theta)$.
+\\
 
@@
 As the data accumulate, the posterior uncertainty is more and more determined by predictive performance, and the impact of the initial uncertainty about $\theta$ is increasingly watered down: `the data overwhelm the prior' \parencite{WrinchJeffreys1919}.
+\\
 
 This implies (but does not dictate, as we discuss next) that a constant stream of accumulating data ought to bring any two people into an arbitrarily close agreement, no matter how divergent their opinions may have been at the outset.
-There are two important caveats here.
-
+Two caveats exist.
 First and foremost, the data only overwhelm the prior if that prior is neither equal to zero (denoting an impossibility) nor one (denoting absolute certainty).
@@
 104]{Lindley1985}.
+\\
 
@@
 This appears irrational, but several Bayesian accounts have been offered to explain the phenomenon (e.g., \cite{CookLewandowsky2016,JernEtAl2014}).
+\\
 
@@
 This way of writing Bayes' rule highlights that extraordinary claims require extraordinary evidence -- if a particular hypothesis $\mathcal{H}_1$ is extremely unlikely a priori, the prior odds $\frac{p(\mathcal{H}_1)}{p(\mathcal{H}_0)}$ are stacked against it, and for the posterior odds to favor $\mathcal{H}_1$ the support from the data (i.e, the degree to which $\mathcal{H}_1$ predicts $\mathcal{H}_0$) needs to be overwhelmingly strong.
+\\
 
@@
 The alternative hypothesis that we entertain here specifies that the coin may be any of the following options with equivalent likelihood: double-tails, fair, or double-heads, $\mathcal{H}_1: \xi \in \{0, \nicefrac{1}{2}, 1\}$.
+\\
 
@@
 Specifically, we now know that $\xi$ cannot be zero; moreover, the datum was twice as likely under $\xi=1$ as under $\xi = \nicefrac{1}{2}$, so that our posterior distribution for $\xi$ under $\mathcal{H}_1$ is now $p(\xi = \nicefrac{1}{2}) = \nicefrac{1}{3}, p(\xi = 1) = \nicefrac{2}{3}$.
+\\
 
@@
 In contrast, the simple model $\mathcal{H}_0$ made a precise prediction.
+\\
 
@@
 This is the penalty for complexity that $\mathcal{H}_1$ pays for entertaining three different values of $\xi$ from the outset.
-Thus, in addition to reflecting the principle of parsimony and highlighting the utility of parsimonious explanations, Bayes rule includes a automatic ``Ockham's razor'' \parencite{Jeffreys1939,JefferysBerger1992} in the sense that daring predictions are rewarded when they come true.
+Thus, in addition to reflecting the principle of parsimony and highlighting the utility of parsimonious explanations, Bayes' rule includes a automatic ``Ockham's razor'' \parencite{Jeffreys1939,JefferysBerger1992} in the sense that daring predictions are rewarded when they come true.
+\\
 
-Finally, we can also write Bayes' rule as follows:
+Lastly, we can also write Bayes' rule as follows:
 \begin{equation}
@@
 
-Having introduced Bayes Theorem and some principles that may follow from it, we next make the case how Bayes could support science literacy at large before delving into implications for science teaching and learning.
-
-\section{How Bayesian Principles Could Bolster Science Literacy}
-
-We conceive that Bayesian methods can enhance \ital{science literacy}, capabilities range from core \ital{literacies}, including numeracy and textual literacy (what is mostly thought of in common parlance concerning literacy), content knowledge, and epistemic knowledge \parencite{council2016science}.
-The specific ways we envision a Bayesian perspective bolstering science literacy are also techniques that teachers can appropriate in their planning and teaching; they may already align with how they teach about science through engaging in scientific practices \parencite{nrc12} or how they articulate the nature of science \parencite{lederman2020avoiding}.
-
-Consider a salient topic as a context to apply these principles, medical vaccines for diseases, such as the disease caused by the coronavirus SARS-CoV-2, COVID-19.
-Most evidence suggests that vaccines have been safe, historically, and are generally safe, at present (Centers for Disease Control and Prevention, 2021).
-Nevertheless, they are not perfectly safe, as unlikely, adverse outcomes do happen, and there may be adverse side effects from some vaccines that research has not yet revealed.
-In light of this evidence, consider the development of a new vaccine, such as those available at the time of this writing for COVID-19.
-Some use long-established methods of vaccine development (namely, deactivating SARS-CoV-2 particles but retaining cell signaling proteins to which humans develop an immune response—and, after, a degree of immunity) and injecting these deactivated virus particles into people), whereas others use a new mechanism (namely, injecting messenger RNA into people; this messenger RNA is then transcribed by cells in people’s bodies to produce a particular protein common to SARS-CoV-2 particles, which human’s bodies then learn to recognize to develop immunity.
-We note that--at present--diseases for which it has been difficult to develop effective vaccines, including Malaria, may have viable mRNA vaccines in the near future \parencite{datoo2021high}.
-
-Before any data from clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and based on similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are.
-After initial trials were completed, the degree of safety that individuals associate with these vaccines—in light of the prior and the limited evidence—could increase, but still not be near the level of safety that many would consider acceptable.
-Following the completion of clinical trials that show minimal side effects (holding aside their efficacy), one’s estimates of the safety of these vaccines could increase by a substantial amount, but concerns could remain, particularly for messenger RNA vaccines, which are newer, and may, plausibly, have unintended side effects.
-
-While the weighting of the evidence in light of prior research and theoretical accounts above may be natural to a scientist, it is challenging for many.
-In the following, we showcase how several important insights for weighting evidence follow directly from Bayes' rule.
-We do not think or expect that members of the public can do this, but members of the public (and learners) do have viable ideas and hypotheses about how the world works.
-Thus, we argue that emphasizing Bayesian reasoning can support science literacy by providing people with a principled way to weight evidence that is accessible because it builds on peoples' prior ideas.
-
-In the context of the above example on SARS-CoV-2, the result of applying Bayes Theorem in a heuristic mode is a non-binary inference, one that doesn't speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty.
-Though a small difference, discussing the safety of vaccines on these terms can provide a more solid foundation for principled disagreements and arguments.
-In addition, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in line with how most of us likely already interpret data \parencite{tgk06, gw12, gh95}; A Bayesian perspective can provide people with a way to talk about weighing between ideas and their prior beliefs and new information and data.
-In short, Bayes gives people a chance to deal with uncertainty in a principled but accessible manner.
-
-In addition, many themes introduced in the section on the epistemological principles that follow from Bayes Theorem are relevant to the aim of supporting science literacy.
-First, \ital{even individuals with different initial beliefs may be brought closer to an agreement in light of data} that serves as evidence for a particular belief or hypothesis.
-In this way, Bayes Theorem illustrates--at the conceptual level as well as statistically--how individuals may be unlikely to agree based on the limited or weak information, but that this can shift in light of accumulating data.
-As we noted earlier, this is far from a deterministic process, and \ital{holding initial beliefs that a hypothesis has a probability of zero or one means that data does not affect one's beliefs}.
-This suggests that a cultural norm and principle for science literacy can be to always reserve a degree of uncertainty such that one could be influenced by (new) data or evidence: From a Bayesian perspective, \ital{our knowledge is uncertain and open to change and revision} so long as we remain open to new evidence.
-
-Related to the importance of not holding scientific beliefs with a probability of zero or one, another principle that follows from Bayes Theorem is that \ital{extraordinary claims require extraordinary evidence}, and that unlikely or implausible hypotheses or claims are likely to necessitate strong empirical data to be convincing to others.
-A corollary of this principle is that we should strong initial or prior beliefs can serve as a tool through which we can discuss extraordinary claims: Except in light of strong evidence, "ordinary" claims can serve as the default for citizens and scientists making sense of unusual and potentially biased claims.
-Similarly, Bayes Theorem demonstrates that the \ital{simplest possible explanation--the most parsimonious explanation--is often the best explanation}.
-
-In sum, a Bayesian perspective offers a general approach to science literacy that can account for uncertainty in a rigorous and principled manner.
-In addition, a number of principles that follow from Bayes Theorem, such as being receptive to new evidence and weighting that evidence per its strength in light of the strength of one's initial beliefs or understanding could bolster science literacy--and how the public comes to understand and gain trust in science.
-
-\section{Toward a Bayesian Perspective in Science Education}
-
-Research on the role of Bayes in science education has mostly been carried out in the context of undergraduate statistics education, in which there have been several calls to action \parencite{gpkpwc18, h_a20}, examples and design options \parencite{a02, b02,g08, w17, h_j20} and debate \parencite{jrhrr20} over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs.
-The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power \parencite{gpkpwc18} as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers \parencite{ah20}.
-As evidenced by the recent special issue of the \emph{Journal of Statistics Education} of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.
-
-There is little research outside of statistics education in the broader science education community.
-Two publications merit mention though.
-The work of \textcite{so12} and \textcite{n11} both applied Bayesian perspectives to the science practice of argumentation.
-Szu and Osborne presented the case for how and why Bayes Theorem can apply to research and practice on scientific reasoning, broadly, and argumentation, particularly.
-This paper represents the most comprehensive account of the relevance of Bayesian methods for science education.
-Szu and Osborne argue that Bayes is useful as both a formal mathematical tool and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes Theorem that follow more or less deterministically from ``external, objectively probabilistic systems'' (p.
-61)—is ``the key leap that characterizes the debate about the value of Bayesian inference as a model of scientific reasoning'' (p.
-61).
-In this way, Szu and Osborne argue that the greatest use of Bayes Theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community \parencite{batanero2016research, mr18}.
-They offer some research-related backing for the use of Bayes Theorem (e.g., noting how conceptual change research, particularly, and constructivist research, broadly, can both be explained coherently within a Bayesian approach) as well as some practical, instructional recommendations, some of which we detail later in this section.
-
-\textcite{n11} work on Bayesian approaches to argumentation describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation.
-Concerning the latter, Nussbaum describes how the social issue of raising taxes to provide resources to homeless individuals was a rich context for students to engage in forms of Bayesian reasoning.
-In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what \textcite{so12} characterize as the more externally objective use of Bayes Theorem—and how the estimates that result from applying Bayes Theorem led students to re-evaluate their initial arguments.
-
-The work of \textcite{so12} and \textcite{n11} begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process.
-Neither, though, speaks to how a Bayesian perspective could have a role in bolstering understanding or trust in science by learners and others.
-More generally, these two contributions are steps toward conceiving what a Bayesian perspective may offer science education, but there remains room to build further.
-For instance, despite their association with statistical methods that are growing in use in post-secondary contexts and in industry \parencite{mcgrayne2011theory}, no science education scholarship has applied Bayesian methods to another science practice, that of analyzing and interpreting data.
-In this way, being clear about what we mean by a \ital{Bayesian perspective} is important: this perspective could be associated not only with a \ital{practice} (analyzing data) but also as a cross-cutting concept \parencite{nrc12} that can, in the context of learning about particular scientific ideas, unite the practice of argumentation with analyzing and interpreting data with the notions of probability and (un)certainty.
-Bayesian methods can also provide a frame that includes a set of epistemological principles and a partial explanation for the process of how and why people change their ideas in light of new information or data--two key components of (and potential barriers to) how members of the public understand and act upon scientific knowledge \parencite{sinatra2014addressing}.
-
-First, curricular resources and pedagogical approaches that take a Bayesian perspective can be developed.
-These could elaborate on the ideas of \textcite{so12} and \textcite{n11}.
-At a high-level, Szu and Osborne write that to reason about data in light of their initial beliefs (and to construct a kind of informal likelihood statistic), students “need to see judgments about data and evidence being an assessment not only of the probability of the hypothesis being correct but also of it being wrong” (p.
-67).
-Filling in more details, speaking from both the perspective of a researcher (of argumentation) and an educator (or, a designer of classroom activities), Nussbaum summarizes that using Bayes “forces one to (a) think about an issue from different perspectives, (b) reflect on one’s range of certainty and uncertainty and (c) conduct a simple computer simulation (e.g., calculating the product) that integrates various facets of the issue together and affords further reflection” (p.
-99).
-
-If a Bayesian perspective is to be realized in classroom contexts, then specific, fine-grained instructional techniques will need to be developed.
-Such instructional techniques do not necessarily need to include technology or statistical software; \ital{unplugged} activities may be useful, and science education scholars could draw from research in other disciplines \parencite[e.g., ][]{caeli2020unplugged} as well as research on the role of epistemic tools in science education \parencite[e.g., ][]{stroupe2019introduction} to design these.
-
-Activities could also be adapted from those designed by statistics educators to support students to reason about uncertainty probabilistically.
-A benefit of these strategies is that they can help to address some of the ingrained challenges that people and learners face when attempting to reason probabilistically \parencite[e.g., ][]{tk74}.
-Particularly, \textcite{bkbm18} for instance, suggest replacing probabilities with natural frequencies.
-This suggestion is in alignment with prior work by \textcite{gh95} on how different representations of probabilities can explain some of the challenges that people and learners face.
-In addition, \textcite{me14}—building on much earlier experimental work by \textcite{wason1971natural}—provide a simple example of how even fourth-graders, with the right tools to represent probabilities, can support students to transition from logical to probabilistic reasoning.
-These could also support the aim of developing unplugged versions of Bayesian analyses.
-\\
-
-Science education researchers may consider how teachers can support students to reason about uncertainty in a way that aligns with a Bayesian perspective.
-While some scholars have contributed foundational work demonstrating how this uncertainty can be productive as a part of teaching practice \parencite{manz2018supporting}, a broad and formal perspective on how uncertainty can be approached in science education is needed.
-Further, a better understanding of uncertainty can help to learn about scientific--and other--concepts where probability plays a key role; it may be that learning about how Bayes rule can be used in legal or medical contexts could impress upon students the cross-cutting nature of probability and uncertainty.
-\\
-
-Finally, recently developed tools, particularly the JASP statistical software \parencite{JASP2020}, may be helpful.
-JASP allows for both conventional frequentist as well as Bayesian versions of commonly used statistical tests, such as \ital{t}-tests and regression analyses.
-There are many accessible guides available for specifying Bayesian versions of analyses \parencite{v20, wagenmakers2018bayesian}, and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms.
-Another possibly useful tool could be Bayesian networks \parencite[e.g., ][]{bayesbox}, which could be used to model complex systems in a way that is similar to other tools used in science education for modeling complex systems, such as \textcite{20}.
-
-\section{Conclusion}
-
-Probability and uncertainty are two concepts that are ubiquitous in the learning, doing, and communicating of science and therefore critical aspects of science literacy.
-However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike \parencite{gkv04, s07, tk74}.
-Thus, it is not surprising when the public misinterprets statements about probability and uncertainty but the consequences may be dire when trust in science is eroded, and, sadly, examples of this are plentiful during current the COVID-19, e.g., the efficiency of masks in preventing the spread of COVID and the efficiency of the AstraZeneca vaccine were strongly questioned, delaying efforts to combat the pandemic.
-\\
-
-We think the blame does not squarely lie with the public in light of findings that the traditional representations of probability and uncertainty are challenging for learners and people alike to comprehend.
-As research from cognitive and psychological science as well as emerging work from statistics and statistics education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions \parencite{kl18, so12, jeffreys1961} about these concepts and thus has the potential to support learning about probability and uncertainty in science.
-Moreover, a Bayesian perspective is unique in that it is akin to a cross-cutting concept in the sense of a style of reasoning \parencite{so12, ork18} that blends conceptual ideas with practices: Bayes is both a philosophical and psychological perspective on how to learn from data as well as a statistical framework.
-
-Indeed, a key potential benefit of Bayes may be that it provides a framework for helping students to transition from their conceptual understanding of scientific ideas to quantitatively analyzing data.
-In this way, we argue for the use of a Bayesian \ital{perspective}, one that could influence research about (and teaching to support) learners' epistemic considerations and how learners' ideas change over time--factors that may have a bearing on how they understand science outside of school, or later as adults \parencite{sinatra2014addressing}.
-Thus, we see there is a number of ways in which a Bayesian perspective can influence science education--and, therefore, we see  a number of mechanisms through which Bayesian methods could bolster trust in science.
-Research and theory suggest that Bayesian ideas could have an important role in science education, but the number of ways that it could make this prospect daunting.
-Different from past research focused on the connection between argumentation and Bayes Theorem, our aim with this paper was to highlight the broad potential of a Bayesian perspective for considering variability and uncertainty in data and evidence and to offer a few directions for how the potential of Bayes in future research.
-
-Even at a time in which science has made and continues to make great advances, the history of science suggests that uncertainty will not be removed (or even reduced) given scientific progress \parencite{fara2010science}; looking ahead, being able to effectively navigate, counsel, and act in the midst of uncertainty may be as important as it has been at other uncertain points in history.
-Like the surprising utility of a Bayesian perspective in domains that are similar to education (e.g., developmental science), might an unexpected feature of a Bayesian perspective be that embracing uncertainty--equipped with the tools that Bayesian methods provide--makes individuals and societies more confident?
-In short, can we build trust in science by embracing uncertainty?
-We think that these are questions that are \ital{likely} to yield positive results for science educators and those advocating for a more scientifically-informed citizenry--and for science learners.
-
-\printbibliography
-
 \end{document}

### 2021-09-22 new-section-for-revised-manuscript.tex
updates for revision

\section{Supporting students in navigating uncertainty with Bayes}

How can Bayes’ Theorem and the principles that follow from it help students to navigate the ubiquitous uncertainty in science? As long as science instruction progresses mostly qualitatively the features of Bayesian reasoning can help to bring a sense of coherence to the many ways in which uncertainty is manifest when students engage scientific practices such as constructing explanations, analyzing and interpreting data, or developing and using models. We argue that Bayesian reasoning can provide this coherence as its distinctive feature, that is, the “concept of degree of belief to describe epistemic attitudes about uncertain propositions” (Sprenger & Hartmann, 2019), embraces the epistemic stance that scientific knowledge is not immutable but in fact open to revision. In other words, scientific knowledge always comes attached with a degree of belief that can be updated following the rules of Bayes’ Theorem when new evidence becomes available. The result of that updating process depends on the uncertainty of existing scientific knowledge which is then expressed in the prior. In this way Bayesian reasoning emphasizes that science and scientific knowledge is always situated in context (Szu & Osborne, 2012) and society (Driver et al., 1994). These ideas can be captured in the following three statements:

a) Be skeptical of absolutes and preserve an open mind – scientific knowledge always comes with some uncertainty and priors that denoting absolute certainty or impossibility permit scientific progress. This first principle expresses the Bayesian degree of belief epistemology which reflects that scientific knowledge is tentative. 
b) Consider what you already know – do not evaluate data/information without considering prior evidence. This emphasizes how science is situated in context and society which both inform how evidence is evaluated.
c) Consider counterfactuals – consider evidence in terms of compatibility with all possible outcomes. This expresses the what in Bayesian philosophy is referred to as the Simple Principle of Conditionalization (Adams, 1965), i.e., when we weigh evidence we must consider to what extent it supports the range of possible explanations for the data.

We argue that emphasizing these statements (or variations thereof) and how they relate to the nature of scientific knowledge can support students in connecting the frequent but often isolated references to uncertainty and limits of scientific knowledge in the scientific practices as expressed, e.g., in the Framework for K-12 Science Education (NRC, 2012). The descriptions of the practices in the Framework are filled with expressions like revising model, refining explanations, critiquing arguments, or considering the limitations of the precision of the data. However, what is lacking is an explicit rationale of why all this revising, refining, considering of limitations is necessary and in fact a feature of science, rather than a bug. Familiarizing students with the principles and core components of Bayesian reasoning can foster a habit of mind from which the elements of the practices that pertain to uncertainty and the limits of scientific knowledge follow naturally. This assumption is supported by the work of (Warren, 2018, 2020) which demonstrates that integrating Bayes reasoning into university science courses positively shifts students epistemic attitudes regarding the nature of knowing and learning. In the following we provide an example of how the quantitative Bayesian updating activities in (Warren, 2020) can be made accessible to students in middle and high school by using a computational tool.

\section{Making it tangible with the confidence updater}

\subsection{The Confidence Updater Shiny app}

Warren (2020) frames the Bayesian reasoning as part of a hypothetico-deductive (HD) process (Popper, 1979). In the HD process, a model is evaluated by deriving a testable hypothesis and testing it against yet unknown data. Depending on ratio between the likelihood of observing the data if the hypothesis is true and the likelihood of observing the data if the hypothesis is false, confidence in the model decreases or increases relative to the initial confidence in the model. This confidence updating can be described using Bayes theorem, writing it in the following way:

P(H|E) is confidence we can have in the hypothesis H after updating our initial confidence P(H) based on consideration of the evidence expressed in R. R > 1 represents confirmatory evidence, R = 1 represents inconclusive evidence, and 0 < R < 1 represents disconfirmatory evidence. Guidelines for choosing an adequate R exist (Kass & Raftery, 1995), e.g., 20 < R < 150 can be interpreted as the evidence strongly favoring H. P(H) the initial confidence in the hypothesis ranges from 0 to 1 with P(H) = 0.5 representing maximum epistemic uncertainty, i.e., having no idea about the validity of the hypothesis. To help students focus on the conceptual elements of updating their confidence in a hypothesis, the shiny confidence updater app in Figure XXX allows students to choose values for R based on their interpretation of the evidence and P(H) based on their initial confidence and after performing the needed calculations returns a textual statement about updated confidence P(H|E) in the Hypothesis after considering the evidence and an optional numeric value for the confidence level as well.

Figure XXX. Shiny web application for updating one’s confidence in a hypothesis following Bayes Theorem. The orange P(H) and R are not displayed to students but imposed on the image to help orient the reader.

The principles a)-c) from the previous section are emphasized either implicitly or explicitly in the design of the confidence updater app. 
Principle a) “Be skeptical of absolutes and preserve an open mind” is emphasized as a natural consequence of Bayes Theorem and the options “absolutely certain that it is correct” and “absolutely certain that it is incorrect” in the “How sure are you about your hypothesis?” part of the confidence updater. When students select these options, P(H) is set to 1 or 0 respectively. In consequence, whatever updating factor R students choose, the updated confidence will always be the same as P(H) and the consideration of the evidence becomes pointless. Thus, if students selected any of the options that are equivalent to P(H) = 1, then observe strong contrary evidence, and wonder how this does not effect, the updated confidence are observing the evidence, a teacher could refer to principle a) and point out to students how fixed prior believes inhibit changes in believes even when confronted with strong evidence – as the students just experienced when using the confidence updater.
Principle b) “Consider what you already know” shows is emphasized in the wording of the prompt for selecting P(H) and becomes visible to students as it effectively moderates the power of the evidence. When prior believes and evidence align, the updated confidence is larger compared to either having “no idea” before or having prior believes that do not align. Students could notice this if they compare the updated confidence P(H|E) for students that had the same prior believe but evaluated the evidence differently or had different prior believes but evaluated the evidence in the same way.

Finally, principle c) "Consider counterfactuals" is emphasized in how the prompt for R is worded. Further, when students use the confidence updater, students should be encouraged to argue for their choice of R and be able to explain why they choose a specific option. In such an argument, students following the prompt should not only consider how their evidence supports their own hypothesis but also how compatible it is with alternative hypotheses.

\subsection{An example}

How does using the confidence updater change a typical science class activity? Here we sketch how using the confidence updater can provide rich opportunities to discuss and reflect on epistemic attitudes. A standard activity in German middle school is to figure out the relation between electric current I, resistance RE, and voltage U, i.e., Ohm’s law (U=RE*I). To do so, students one can measure the resulting current for different voltages in a circuit with fixed resistance. At the beginning of the activity, students are asked to come up with different hypotheses for the relationship. They often propose a proportional, inverse, or quadratic relationship. Now, the students may form groups, select a hypothesis, and set out to test it. Following principle a) students can now specify their prior confidence in their group’s hypothesis. As the prior is subjective students may very well choose different priors based on their individual experiences. Now, they may proceed and record a number of measurements of the current for different voltages and then graph the data. Figure XXX shows an example graph. Now, when students are asked to evaluate the Evidence and set a value for R, they are urged to consider to what extent the evidence is consistent with their hypothesis vs. an alternative which is emphasized in principle c). Looking at the graph one could see evidence for a proportional relationship as well as a higher order positive relationship (blue and red dashed line in Figure XXX). Thus, the evidence is barely compatible with an inverse relationship between I and U but supportive of both hypotheses that suggest a positive relationship between the variables. Thus, depending on what hypothesis students choose to investigate, they should select “data favors an alternative hypothesis” for R if their hypothesis is inverse relationship and select “data somewhat favors my hypothesis” if their hypothesis is proportional or quadratic relationship. Now, the updated confidence for every student will depend on their initial confidence in their hypothesis and the evaluation of the data which will be different across groups. If students compare their results within groups, they should see how the effect of their initial confidence influences the effect of the evidence. However, if they repeat the activity, i.e., use their updated confidence as a new initial believe and collect more data, they should see how consensus is approached over a few iterations unless anyone choose P(H) = 0 or P(H) = 1.  Frist, this emphasizes the importance of principle a) for science again. Second, as (Warren, 2018) points out, “the fact that everyone starts with subjective prior probabilities and yet inevitably converges to a single asymptotic result illustrates the objective aspect of science” which can help build trust in science. Lastly, if student compare their results between the groups, students can note how disconfirmatory evidence has more epistemic power than confirmatory evidence, which is also another direct corollary of Bayes Theorem.

\subsection{Figure XXX. Possible graph of current and voltage data.}

### 2021-09-22 old-text/axioms.tex
updates for revision

--- 2021-05-01 manuscript.tex
@@
-%%
-%% This is file `./samples/longsample.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% apa7.dtx  (with options: `longsample')
-%% ----------------------------------------------------------------------
-%% 
-%% apa7 - A LaTeX class for formatting documents in compliance with the
-%% American Psychological Association's Publication Manual, 7th edition
-%% 
-%% Copyright (C) 2019 by Daniel A. Weiss <daniel.weiss.led at gmail.com>
-%% 
-%% This work may be distributed and/or modified under the
-%% conditions of the LaTeX Project Public License (LPPL), either
-%% version 1.3c of this license or (at your option) any later
-%% version.  The latest version of this license is in the file:
-%% 
-%% http://www.latex-project.org/lppl.txt
-%% 
-%% Users may freely modify these files without permission, as long as the
-%% copyright line and this statement are maintained intact.
-%% 
-%% This work is not endorsed by, affiliated with, or probably even known
-%% by, the American Psychological Association.
-%% 
-%% ----------------------------------------------------------------------
-%% 
-\documentclass[man]{apa7}
-% \documentclass[man, mask, floatsintext]{apa7}
+\section{Appendix A: The Epistemological Axioms that Follow from Bayes Theorem}
 
-\usepackage[american]{babel}
-\usepackage{nicefrac}
-\usepackage{todonotes}
-\usepackage{epigraph}
-\usepackage{amsfonts, amsmath,amssymb}
-
-\newcommand{\SR}[1]{} %for EJ's bibtex file
-\newcommand{\given}{\, | \,}
-\newcommand{\EJ}[1]{\todo[inline, color=green]{  #1 }}
-\newcommand{\JR}[1]{\todo[inline, color=gray]{  #1 }}
-\newcommand{\MK}[1]{\todo[inline, color=red]{  #1 }}
-
-\usepackage{graphicx}
-\graphicspath{ {images/} }
-
-\usepackage{csquotes}
-\usepackage[style=apa,sortcites=true,sorting=nyt,backend=biber]{biblatex}
-\DeclareLanguageMapping{american}{american-apa}
-\addbibresource{bibliography.bib}
-\addbibresource{referenties.bib}
-
-% \setlength{\parindent}{8ex}
-
-% \usepackage{authblk}
-
-\title{Building Trust in Science Through a Bayesian Perspective on Probability and Uncertainty}
-\shorttitle{Building Trust Through a Bayesian Perspective}
-
-\threeauthors{Joshua M. Rosenberg}{Marcus Kubsch}{Eric-Jan Wagenmakers}
-\threeaffiliations{University of Tennessee, Knoxville}{IPN - Leibniz Institute for Science and Mathematics Education}{University of Amsterdam}
-
-% \authorsnames[1,2,3]{Joshua M. Rosenberg, Marcus Kubsch, Eric-Jan Wagenmakers}
-% \authorsaffiliations{{University of Tennessee, Knoxville},{IPN - Leibniz Institute for Science and Mathematics Education},{University of Amsterdam}}
-
-\leftheader{Weiss}
-
-\abstract{
-Uncertainty is ubiquitous in science, but scientific understanding is often represented to the public and in educational contexts as certain and immutable.
-Adding to this problem is that when scientists try to communicate their uncertainty, research has demonstrated that many people--learners and experts alike--struggle to interpret the statistics that are often used to make inferences about evidence from scientific data.
-We conjecture that a Bayesian perspective can offer a foundation for teaching and learning about a persistent but sometimes unsettling tension for science, making sense of and communicating about uncertainty.
-Central to this approach is viewing uncertainty \ital{probabilistically}, as doing so provides a language for expressing the degree of uncertainty in knowledge across a range of scientific domains and socio-scientific issues.
-In this paper, we elaborate on the importance of probability and uncertainty in science (and science education) and review how people learn about these concepts---and when and why they face difficulties when doing so.
-Then, we provide an accessible primer on Bayes Theorem and more general principles following from the application of Bayes Theorem.
-Next, we provide examples of how a Bayesian perspective could inform efforts around both science literacy and science education with an emphasis upon building trust in science.
-We conclude with directions for future research focused on bolstering trust in science outside of and within pre-collegiate science classrooms, where a Bayesian perspective can provide a coherent approach to engaging in science to make sense of scientific ideas in a principled yet flexible manner.
-}
-
-\keywords{Bayesian, uncertainty, probability, trustworthiness, statistics}
-
-\authornote{
-   \addORCIDlink{Joshua M. Rosenberg}{0000-0003-2170-0447}
-   \addORCIDlink{Marcus Kubsch}{0000-0001-5497-8336}
-   \addORCIDlink{Eric-Jan Wagenmakers}{0000-0003-1596-1034} 
+In this section, we build on the idea of the Bayesian Learning Cycle--that a Bayesian perspective aligns with a way of systematically taking in new information to improve what one knows (and how certain one is about what one knows--to elaborate on some general principles related to the nature of knowledge and knowing, or epistemological principles.
+We illustrate these to later show how they can have a role in the science classroom and how they can be used by individuals outside of classrooms and schools to understand and act upon scientific information.
 \\
-   Joshua M. Rosenberg and Marcus Kubsch contributed equally to this manuscript.
-}
-
-\begin{document}
-\maketitle
-
-\setlength{\epigraphwidth}{1.0\textwidth} % set for individual epigraph
-\epigraph{It is seen in this essay that the theory of probabilities is at bottom only common sense reduced to calculus; it makes us appreciate with exactitute that which exact minds feel by a sort of instinct without being able ofttimes to give a reason for it .
-.
-.
-we shall see that there is no science more worthy of our meditations, and that no more useful one could be incorporated in the system of public instruction.}{Pierre-Simon Laplace, 1829}
-\setlength{\epigraphwidth}{0.4\textwidth} % return to default for the next epigraphs
-%\nocite{Laplace18291902} %[p. 196]
-
-Uncertainty is ubiquitous in science and creating ways to account for this uncertainty has helped science to progress.
-Consider, for instance, uncertainty due to the measurements or instruments used to study phenomena: In such cases, error due to measurement can be incorporated into models through the repeated measures of the same object or phenomenon \parencite{pls03}.
-Measurement is also present when the object that is measured--like a plant--is variable in its characteristics because of environmental and genetic variation (and interactions among these sources of variation) between the individual plants \parencite{pls03}.
-Another form of uncertainty is accounted for in terms of its central role within explanatory theories--such as those associated with evolution and quantum physics--for which models take the form of probabilistic statements about the world \parencite{g00}.
-
-Negotiating the many forms of uncertainty through (constructively) arguing—-and building toward consensus at the same time that uncertainty is present—-is a key part of the scientific process \parencite{duschl2008science, t00, nrc12}.
-Concomitantly, some philosophers of science have suggested that science is a process that builds increasingly better models which allows us to make increasingly more accurate predictions \parencite{g10, g06, n02, r77}.
-However, while scientists and historians and philosophers of science are well aware of the uncertainty (and limitations) around scientific knowledge and the practice of negotiating and arguing from evidence, scientific knowledge is often represented to the public and in educational contexts as certain and immutable \parencite{d90}.
-Trust in science can erode and has eroded when scientists individually or collectively change their views, such as during the first phase of the COVID-19 crisis, when scientists faced serious criticism from the public, the media, and elected representatives for changing their recommendations based upon new scientific data and findings.
-A similar tension is now present concerning vaccines for COVID-19--a tension that involves reconciling robust but still initial clinical trial data and new data from population-level vaccination efforts and making policy-level recommendations (and individual decisions) amid this uncertainty.
-Adding to this difficult state of affairs regarding the role of uncertainty in science is that when scientists try to communicate their uncertainty, prior research has demonstrated that many people–learners and experts alike–struggle to interpret the statistics and derivations of statistics (such as standard errors/margins of error and confidence intervals) and that are often used to make inferences from scientific data \parencite{gkv04, s07, tk74}.
-Similarly, difficulties in understanding the technical language of probability can contribute to a struggle on the part of students when learning about concepts such as evolution and nuclear decay that are modeled in terms of probabilistic statements \parencite{fth17}.
-
-In this paper, we argue that these challenges concerning trust in scientific knowledge that not only can but must change can be addressed in a principled, flexible way, but that doing so requires making use of a Bayesian perspective that heretofore has been under-used outside of particular scientific and engineering domains---domains that often make little contact with science educators and researchers with an interest in building trust in science.
-Yet, the opportunity to do so now exists for technical and social reasons.
-Based on accumulating evidence from psychologists using Bayesian models of cognition \parencite{g12, tgk06}, statisticians and statistics educators advancing tools and practices for Bayesian methods \parencite{a02, b02, h_a20, h_j20,me14, s07} and science education scholars who have advanced a Bayesian perspective on argumentation \parencite{n11, s07}, we conjecture that a Bayesian perspective can help people--and science learners--to understand and reason about uncertainty in a way that leverages their existing ways of thinking.
-Central to this conceptualization is viewing uncertainty \ital{probabilistically}, as probabilities can provide a language for expressing the degree of uncertainty in our knowledge in a wide range of domains \parencite{g12}.
-In drawing on research from psychologists, learning scientists, and educational researchers, we make this conjecture not only on the merits of Bayes' Theorem as a mathematical or statistical tool, but primarily on evidence that the perspective we are advocating can provide an intuitive way for people to think about data and evidence in the world \parencite{g12} and in classroom contexts \parencite{ll20, n11, s07, so12, w17}.
-
-\section{Uncertainty in Science and the Central Role of Probability}
-
-Probability and uncertainty are ubiquitous in scientific methodologies, scientific concepts, and in how science is communicated \textcite{gsoobmc17}.
-The practice of science often begins with measurement for which measurement error--which has a random component--is endemic.
-This random component arises from random variations in the measurement process and limits the certainty that one can have in the measurements.
-Reducing that measurement error has been critical for many discoveries in science, such as in sorting out the periodic table as we know it today \parencite{fco15}.
-However, deciding on adequate statistical procedures to determine measurement uncertainty regularly sparks debate, e.g., the recent study on the relationship between SARS-CoV-2 viral load and patient age by \textcite{jmvbzhd20} was heavily debated in the scientific community \parencite{frick_peer-review_2020}.
-
-Probability and uncertainty are not only key concepts in science practices (such as analyzing and interpreting data) but are also an integral part of explanatory models of and theories for scientific concepts.
-For example, in the context of Heisenberg’s uncertainty principle and--more generally--quantum mechanics \parencite{feynman1951operator}, confidence is limited not only by measurement error but in principle---even in the perfect experimental setup there is no deterministic outcome.
-Instead, one has to calculate the probability of each possible outcome using the Born rule.
-To this day, there is an ongoing argument among physicists and philosophers about how to interpret probability and the puzzling quantum mechanical phenomena such as quantum tunneling \parencite{c19}.
-Another example of how probability and uncertainty are central to learning scientific concepts is evolution \parencite{fsnh19}.
-Evolution is modeled as a probabilistic process as the emergence of new variations (e.g., through mutations) can only be described probabilistically.
-Further, the continued survival of these variations is governed by phenomena such as natural selection or random drift which also escape a deterministic description and are thus modeled probabilistically.
-Similar to within quantum physics, how to interpret this probability has also sparked debate among researchers \parencite{m16}.
-Given that probability and uncertainty spark debate among scientists, it is perhaps unsurprising that misinterpretations happen when scientists in the natural and engineering sciences communicate about their findings with their peers or the public \parencite[]{gkv04, c14, mg17}.
-For example, error bars in the form of confidence intervals are routinely interpreted as distributions that assign a higher probability to the center of the interval \parencite{kl18} or \emph{p}-values are interpreted as effect sizes \parencite{gc17, n14} (although p-values describe how (in)compatible a set of data is with a set of assumptions.
-In sum, misinterpretations of statements about probability and uncertainty are common across the range of modes in which scientists communicate about their work.
-
-\section{Uncertainty in Science Education}
-
-\subsection{The Roles of Probability Across Scientific Disciplines}
-
-Given the above-noted roles of probability in scientific concepts and theories, it is unsurprising that educational researchers have documented how students learn about probability.
-As one example, probability and chance events play an important role in the life sciences \parencite{g03, gk08}, particularly in evolutionary processes and learning about them \parencite{th17}.
-Recently, \textcite{fsnh19} investigated the relation between statistical reasoning and acceptance and knowledge about evolution in a large sample of nearly 500 university students in the US.
-They found that students’ statistical reasoning capabilities are a strong predictor of both acceptance of evolution and knowledge about evolution.
-
-In the physical sciences, students' struggles with the probabilistic nature of quantum physics \parencite{br02} and nuclear decay \parencite{smmv17} have been established, but the underlying reasons and mechanisms remain less researched than in evolution.
-Research into students’ misconceptions inspired by conceptual change research has started to address this gap \parencite{ms15, alra14, st09}.
-Findings mirror the position of \textcite{bgv94}.
-Specifically, the concepts of probability are taught abstractly and in such a way that does not build upon the ideas that students already hold, the experiences that students have, and the language that students use.
-Thus, these studies affirm students' struggles with scientific concepts that are probabilistic but hint at an underlying issue of more foundational challenges with reasoning about probability beyond science education contexts.
-
-Indeed, past research has shown that the language and representations used to communicate probability and uncertainty independent of specific scientific content domains or scientific practices are also known to present difficulties to learners.
-The work of \textcite{tk74}, especially, points to a whole range of biases that the language in which statements about probability and uncertainty are framed can elicit.
-A prominent example is the neglect of base rate information in the (timely) context of medical testing, that is, the probability of having a medical condition based on a positive test result is often hugely overestimated because information about the base rate of people affected from the condition is neglected \parencite{kahneman_thinking_2012}.
-
-Responding to and building upon Kahneman and Tversky's work, Gigerenzer and colleagues provided important insights into the underlying mechanisms for the occurrence of these biases.
-Their research suggests that the biases are at least partly the result of information about probability and uncertainty being provided in a way that is incompatible with the heuristics that people develop from their everyday experiences \parencite{gh95, jkg18}.
-Based upon this premise, \textcite{jkg18} were able to demonstrate that students judge probability significantly better when language is adopted that better aligns with the heuristics that most learners develop from their everyday experiences.
-In sum, it seems feasible to change representations of probability and uncertainty so that they better align with ideas and heuristics that learners hold so that these support, rather than hinder, learning.
-
-\subsection{Uncertainty in the Context of Analyzing and Interpreting Data}
-
-Recent work--primarily by developmental psychologists and scientists--has applied Bayesian ideas to a new context: how children make sense of and world \parencite{g12, tgk06, tgk06}.
-This notion that children reason about the world in a way that is analogous to Bayes Theorem is in some ways surprising, but it also builds upon much earlier work that also documented the surprisingly complex ways that children could reason about the world---as well as the assumed limitations of children's reasoning, the work of \textcite{pi69}.
-Different from earlier work on the stages through which children develop, recent work using Bayesian models considers the differences in the reasoning ability of children and adults to be a matter of degree, rather than kind.
-In short, children hold initial ideas and understandings that they change (or update) in ways that are concordant with how Bayes Theorem represents the updating of an initial belief in light of data (and how the Bayesian Learning Cycle described earlier allows individuals to integrate inductive and deductive reasoning).
-
-In addition to developmental research that uses Bayes Theorem as an analog for how children make sense of data and information in light of their initial ideas and beliefs, other work bears a striking semblance to developmental research that does.
-In particular, research on how learners’ prior understanding influence how they make sense of how likely different patterns in data are centers the same tension around interpreting data in light of initial beliefs that a Bayesian perspective highlights \parencite[e.g., ][]{kd88, mkm07, ssc07}.
-While these studies have focused on how students’ prior ideas and data interact, Bayesian frameworks were not used to formalize this process of integrating data.
-
-A Bayesian view of how children learn from their experiences and update their beliefs about the world has become especially prominent over the past 10-15 years \parencite{gt07, gw12}.
-Because of this, one might assume that there has been a ready application of these ideas about how children think about the world in a Bayesian way to education and developmental scientists have pointed to this opportunity.
-For instance, \textcite{g12} wrote that the use of Bayesian methods to understand child development can serve as “a scientific foundation for a long tradition on ‘inquiry-based’ science education,” that “could lead us to much more specific and scientifically supported proposals for education” (p.
-1627).
-
-However, despite arguments such as Gopnik's (2012) for integrating developmental and educational research, it has largely not been the case that “science itself could help turn young children’s natural curiosity and brilliance into better science teaching and learning” \textcite[p.
-1627]{g12}, as some science education scholars have bemoaned \parencite{ls15}).
-Indeed, scholars have argued that it is often obsolete applications of Piaget’s ideas \parencite[e.g., ][]{pi69} that represent the greatest contribution of developmental ideas to education \textcite{ls07}.
-We argue that this is still largely the case in the present, despite advances in both developmental research as well as statistical methodology related to Bayesian ideas—and that it is time to start exploring how bringing out the inherent Bayesian ways of thinking in education may provide support for learners’ efforts to make sense of the world.
-
-At the same time that opportunities to connect developmental science with education (especially science education), science education research has emphasized how important students' prior knowledge and understandings are to their understanding of scientific concepts and processes \parencite[e.g., ][]{ssc07}.
-For example, \textcite{mkm07} summarized findings across several studies showing that when elementary grade students reasoned with data in a context about which they had strong and valid prior understandings that informed their interpretation of data (about force and motion in the context of ramps), and they were able to explain sources of error in the data \parencite{mk03}, but when the data was in a context they did not understand well, they struggled to distinguish between random variation due to sources of error and sources of variation due to something that was causing a difference in the data (generated by measuring the motion of pendulums).
-Later design- and laboratory-based research has shown that students can develop an understanding of scientific ideas through the analysis of data \parencite{ls04,mkk17}.
-
-In summary, developmental and psychological research into Bayesian models of cognition supports and bolsters efforts in education--especially science education--to design instruction based around eliciting and understanding students’ ideas \parencite{gb16, hbavb20, wtbs12}.
-There is an opportunity to use Bayesian ideas as a framework for formalizing these ideas about how prior ideas affect students’ reasoning about data and how teachers can leverage students’ ideas to design instruction, but a Bayesian perspective has--with a few exceptions focusing on how Bayes Theorem can provide a model for the science practice of argumentation \parencite[e.g., ][]{n11, so12}--has been absent from research in science education.
-
-After having reviewed how people learn about probability and uncertainty and where they face difficulties and research on uncertainty in the context of analyzing data, we next formally introduce Bayes Theorem and describe the broader process through which one's beliefs can be revised, or updated, in a way that leverages both inductive reasoning from data and deductive reasoning from beliefs or principles--what we refer to as the \ital{Bayesian Learning Cycle}.
-We also describe some of the ways of knowing, or epistemological considerations, that follow from Bayes Theorem and which comprise a core part of the Bayesian perspective that we argue has a role in building trust in science.
-
-\section{A Primer on Bayes Theorem and a Bayesian Epistemology}
-
-\subsection{An Accessible Introduction to Bayes Theorem}
-
-At its core, Bayes' theorem provides a means to update one's beliefs about the world from observations in such a way that prior knowledge can be used in a principled manner.
-As represented in Equation 1, Bayes' Theorem states that at any point in time, our knowledge about the world after making a new observation (i.e., our posterior knowledge) can be obtained by multiplying our prior knowledge with a predictive updating factor (e.g., \cite{RouderMorey2019,WagenmakersEtAl2016CD} ):
-\begin{equation*}
-        \text{Posterior uncertainty} = \text{Prior uncertainty} \times \text{Predictive updating factor}.
-\end{equation*}
-Mathematically, this can be expressed as  
-\begin{equation}
-\label{eq:BayesRule}
-        p(\theta \mid \text{data}) = p(\theta) \times \frac{p(\text{data} \mid \theta)}{p(\text{data})},
-\end{equation}
-where $\theta$ is a hypothesis, proposition, claim, parameter value, or in general any particular account of the world about which we are uncertain.
-The predictive updating factor consists of two components.
-First, $p(\text{data} \given \theta)$ is the likelihood, that is, the extent to which the observed data were expected under $\theta$.
-Second, $p(\text{data}$ is the marginal likelihood, that is, the extent to which the observed data were expected without conditioning on $\theta$.
-If the observed data are more expected (i.e., better predicted) when conditioning on $\theta$, the predictive updating factor is higher than 1, and $\theta$ becomes more credible.
-
-To illustrate the above equation, consider the following: An educator is interested in the proportion of students that use Windows or MacOS as their operating system.
-The proportion of Windows users $\theta$ ranges between 0 and 1.
-Before you start our Bayesian analysis, you need to quantify your background knowledge -- the relative plausibility of the different values for $\theta$.
-Given their previous experiences as a teacher, their best guess is that about half of the students use Windows.
-They would not be surprised if the proportion of Windows users was between 30\% and 70\% but you would be rather surprised to find either no Windows or exclusive Windows usage.
-
-This background knowledge motivates the use of the dome-shaped prior distribution shown as the dashed line in Figure~\ref{fig:WindowsMacPriorPosterior}.\footnote{This is the beta$(2,2)$ distribution.} Now suppose you start to collect data by asking students about the operating systems of their computers.
-The first ten observations of Windows (W) and MacOS (M) usage that you make are $\{W, W, M, W, M, W, W, M, W, M\}$: six students use Windows and four use MacOS.
-
-With these data in hand, we then combine our prior knowledge (i.e., the dome-shaped prior distribution) with the predictive updating factor (assuming a binomial likelihood) and obtain the posterior distribution shown as the solid line in Figure~\ref{fig:WindowsMacPriorPosterior}.
-The posterior distribution is more peaked than the prior distribution, indicating that the data have sharpened our knowledge about $\theta$.
-In addition, the posterior distribution is higher than the prior distribution for values of $\theta$ between approximately $.4$ and $.75$: these values of $\theta$ predicted the data relatively well and have therefore gained credibility.
-In contrast, values of $\theta$ lower than $.4$ and higher than $.75$ predicted the data relatively poorly, which is why they have lost credibility compared to the prior.\footnote{Note that the area under a probability distribution equals 1, so that posterior gains for specific regions of $\theta$ cancel out exactly against posterior losses for the other regions.}
-
-\begin{figure*}[h]
-\begin{center}
-\includegraphics[width = .785\paperwidth]{WindowsMacPriorPosterior.pdf}
-\caption{Simple example of Bayesian learning.
-A dome-shaped prior distribution captures the background knowledge concerning the proportion of students using Windows vs.
-MacOS.
-Observing fictitious data from 10 students (six using Windows, four using MacOS) drives a knowledge update that results in a bell-shaped posterior distribution.
-Figure based on the \emph{Learn Bayes} module in JASP.}
-\label{fig:WindowsMacPriorPosterior}
-\end{center}
-\end{figure*}
-
-Although we demonstrated this updating process in a single step, it could also be executed sequentially, one student at a time.
-For instance, the first observation you make is a $W$ and in light of your prior this makes Windows usage a little more likely than MacOS usage.
-This slight change in knowledge is reflected in the difference between the distributions on the top two lines in Figure~\ref{fig:WindowsMacSequential}; line ``0'' represents the dome-shaped prior distribution and line ``1'' represents the posterior distribution after the first observations.
-Note that observing a $W$ has nudged the distribution a little to the right.
-Crucially, the `nudged' posterior distribution after the first observation now takes the role of the prior distribution ready to be updated by our second observation.
-The second observation is again a $W$, and line ``2'' shows that the resulting posterior distribution is nudged to the right once more.
-At every stage in the sequential updating process, the posterior distribution based on the observations seen so far becomes the prior distribution for incorporating the information from the next observation.
-Specifically, the changing distributions in Figure~\ref{fig:WindowsMacSequential} show our knowledge about the world is subject to constant change, by making observations and integrating incoming information with current knowledge using Bayes' theorem.
-It should be emphasized that although the updating process is conceptually different, the final outcome is identical: the shape of the posterior distribution is unaffected by whether data arrive simultaneously or sequentially.
-
-\begin{figure*}[h]
-\begin{center}
-\includegraphics[width = .65\paperwidth]{WindowsMacSequential.pdf}
-\caption{Sequential Bayesian learning.
-A dome-shaped prior distribution (line ``0'') captures the background knowledge concerning the proportion of students using Windows vs.
-MacOS.
-Each new observation results in an update to a posterior distribution, which then becomes the prior distribution for the analysis of the next observation.
-Figure based on the \emph{Learn Bayes} module in JASP.}
-\label{fig:WindowsMacSequential}
-\end{center}
-\end{figure*}
-
-\subsection{Integrating Inductive and Deductive Reasoning With the Bayesian Learning Cycle}
-
-How do the principles from the above example translate to scientific data?
-A recent case from the physical sciences can serve to illustrate this approach.
-In 2011, physicists reported that they had observed elementary particles moving faster than the speed of light \parencite{b11} - and thus defying Einstein’s theory of relativity.
-However, instead of celebrating a historic discovery, the scientists started to look for what had gone wrong in their observations.
-With nearly a century of accumulated evidence that the theory of relativity holds, from a Bayesian perspective, scientists had a very strong prior belief about the speed of light.
-To change this belief, extraordinary evidence would have been required.
-Thus, a single instance of apparently faster-than-light elementary particles does not give a likelihood strong enough to change the prior in a meaningful way, and in consequence, the posterior--the knowledge about the world after making an observation--does not change in a meaningful way.
-Thus, rather than changing their belief about the speed of light, scientists looked for things that could have gone wrong with the measurement.
-This example shows how a qualitative take on Bayes’ theorem can explain and predict many aspects of how animals \parencite{o13}, people \parencite{tgk06}, and organizations \parencite{kaj12} learn from observations.
-Recently, advances in computational power and techniques have started the more and more widespread application of this principle in a quantitative way that allows researchers to precisely express their knowledge about the world and then learn from observations.
-Expressed in the Bayesian learning cycle (Figure \ref{fig:The Bayesian Learning Cycle}), a Bayesian reasoning framework allows scientists to integrate the strengths of both, inductive and deductive research approaches, in a formalized way.
-Further, when specifying the prior, scientists openly and explicitly need to describe the uncertainty in the current state of knowledge, and the posterior captures how the data has changed the knowledge and the uncertainty in that knowledge.
-Thus, Bayesian reasoning is, at least in principle, open and transparent by design.
-\begin{figure*}[h]
-\begin{center}
-\includegraphics[width = .65\paperwidth]{learning.png}
-\caption{The Bayesian learning cycle is a reasoning framework that allows scientists, citizens, and learners to integrate the strengths of inductive and deductive reasoning while making sense of data in light of prior beliefs.}
-\label{fig:The Bayesian Learning Cycle}
-\end{center}
-\end{figure*}
-Having introduced Bayes Theorem and an application of Bayes Theorem as a cycle through which individuals can learn, we show how Bayes Theorem can be used to demonstrate several principles that are relevant to building trust in science.
-
-\section{The Epistemological Principles that Follow from Bayes Theorem}
 
@@
 Thus, Bayes' rule states that our posterior knowledge $p(\theta \given \text{data})$ is proportional to the likelihood $p(\text{data} \given \theta)$ (or the extent to which the observed data are expected given $\theta$) multiplied by our prior knowledge $p(\theta)$.
+\\
 
 Equation~\ref{eq:BayesRulePropto} emphasizes that the posterior uncertainty about $\theta$ is a \emph{compromise} between our prior uncertainty about $\theta$ and the predictive performance of $\theta$.
-But, the posterior uncertainty after having observed the first datum becomes the prior uncertainty for the next datum---and any other data (cf.
+But, the posterior uncertainty after having observed the first datum becomes the prior uncertainty for the next datum--and any other data (cf.
 Figure~\ref{fig:WindowsMacSequential}).
@@
 As the data accumulate, the posterior uncertainty is more and more determined by predictive performance, and the impact of the initial uncertainty about $\theta$ is increasingly watered down: `the data overwhelm the prior' \parencite{WrinchJeffreys1919}.
+\\
 
 This implies (but does not dictate, as we discuss next) that a constant stream of accumulating data ought to bring any two people into an arbitrarily close agreement, no matter how divergent their opinions may have been at the outset.
-There are two important caveats here.
-
+Two caveats exist.
 First and foremost, the data only overwhelm the prior if that prior is neither equal to zero (denoting an impossibility) nor one (denoting absolute certainty).
 If someone already knows for certain that a hypothesis is true or false, then you cannot adjust your opinion in light of the data: your initial opinion will be your final opinion, no matter what the data may indicate.
-Philosophically, adopting such priors is a dangerous practice; for instance, in antiquity the adage of the \emph{New Academy} --a school of skeptics headed by Carneades-- was ``never assert absolutely''.
+Philosophically, adopting such priors is a dangerous practice; for instance, in antiquity, the adage of the \emph{New Academy} --a school of skeptics headed by Carneades-- was ``never assert absolutely''.
 In modern times, Lindley popularized this idea in statistics and coined it ``Cromwell's rule'' \parencite[p.
 104]{Lindley1985}.
+\\
 
@@
 Thus, in addition to reflecting the principle of parsimony and highlighting the utility of parsimonious explanations, Bayes rule includes a automatic ``Ockham's razor'' \parencite{Jeffreys1939,JefferysBerger1992} in the sense that daring predictions are rewarded when they come true.
+\\
 
-Finally, we can also write Bayes' rule as follows:
+Lastly, we can also write Bayes' rule as follows:
 \begin{equation}
@@
 This is in stark contrast to $p$-value statistical hypothesis testing, in which a statistical model (the null hypothesis) is judged in isolation.
-Having introduced Bayes Theorem and some principles that may follow from it, we next make the case how Bayes could support science literacy at large before delving into implications for science teaching and learning.
-
-\section{A Bayesian Perspective and Science Literacy}
-
-We conceive that Bayesian methods can enhance science literacy or a range of capabilities that people may possess which have value (individually or collectively) to them \parencite{council2016science}.
-These capabilities range from core \ital{literacies}, including numeracy and textual literacy (what is mostly thought of in common parlance concerning literacy), content knowledge, and epistemic knowledge \parencite{council2016science}.
-A potentially powerful way in which Bayesian ideas can inform discussions of science literacy is to provide individuals with a principle for combining their prior understanding with new evidence.
-Along this line, Bayes and the Bayesian Learning Cycle can provide a tool for people (and learners) to combine their understanding with new evidence (and data) rigorously and coherently.
-
-A timely example can help to illustrate how a Bayesian perspective on a topic relevant to science literacy may offer a different way forward.
-Consider a salient topic, medical vaccines for diseases, such as Polio and the disease caused by the coronavirus SARS-CoV-2, COVID-19.
-Most evidence suggests that vaccines have been safe, historically, and are generally safe, at present (Centers for Disease Control and Prevention, 2021).
-Nevertheless, they are not perfectly safe, as unlikely, adverse outcomes do happen, and there may be adverse side effects from some vaccines that research has not yet revealed.
-In light of this evidence, consider the development of a new vaccine, such as those available at the time of this writing for COVID-19.
-Some use long-established methods of vaccine development (namely, deactivating SARS-CoV-2 particles but retaining cell signaling proteins to which humans develop an immune response—and, after, a degree of immunity) and injecting these deactivated virus particles into people), whereas others use a new mechanism (namely, injecting messenger RNA into people; this messenger RNA is then transcribed by cells in people’s bodies to produce a particular protein common to SARS-CoV-2 particles, which human’s bodies then learn to recognize to develop immunity.
-
-How should we consider evidence about the vaccines?
-Before any data from medical clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and based on similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are.
-After initial trials were completed, the degree of safety that individuals associate with these vaccines—in light of the prior and the limited evidence—could increase, but still not be near the level of safety that many would consider acceptable.
-Following the completion of clinical trials that show minimal side effects (holding aside their efficacy), one’s estimates of the safety of these vaccines could increase by a substantial amount, but concerns could remain, particularly for messenger RNA vaccines, which are newer, and may, plausibly, have unintended side effects.
-
-While the complex weighting of the evidence above may seem natural to a scientist, it is challenging for many.
-In the following, we showcase how several important insights for weighting evidence follow directly from Bayes' rule.
-Thus, we argue that emphasizing Bayesian reasoning can support science literacy by providing people with a principled way to weighting evidence that is accessible because it builds on peoples' prior ideas.
-
-In the context of the above example on SARS-CoV-2, the result of applying Bayes Theorem in a heuristic mode is a non-binary inference, one that doesn't speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty.
-Though a small difference, discussing the safety of vaccines on these terms can provide a more solid foundation for principled disagreements and arguments.
-In addition, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in line with how most of us likely already interpret data \parencite{tgk06, gw12, gh95}; A Bayesian perspective can provide people with a way to talk about weighing between ideas and their prior beliefs and new information and data.
-In short, Bayes gives people a chance to deal with uncertainty in a principled but accessible manner.
-
-In addition, many themes introduced in the section on the epistemological principles that follow from Bayes Theorem are relevant to the aim of supporting science literacy.
-First, \ital{even individuals with different initial beliefs may be brought closer to an agreement in light of data} that serves as evidence for a particular belief or hypothesis.
-In this way, Bayes Theorem illustrates--at the conceptual level as well as statistically--how individuals may be unlikely to agree based on the limited or weak information, but that this can shift in light of accumulating data.
-As we noted earlier, this is far from a deterministic process, and \ital{holding initial beliefs that a hypothesis has a probability of zero or one means that data does not affect one's beliefs}.
-This suggests that a cultural norm and principle for science literacy can be to always reserve a degree of uncertainty such that one could be influenced by (new) data or evidence: From a Bayesian perspective, \ital{our knowledge is uncertain and open to change and revision} so long as we remain open to new evidence.
-
-Related to the importance of not holding scientific beliefs with a probability of zero or one, another principle that follows from Bayes Theorem is that \ital{extraordinary claims require extraordinary evidence}, and that unlikely or implausible hypotheses or claims are likely to necessitate strong empirical data to be convincing to others.
-A corollary of this principle is that we should strong initial or prior beliefs can serve as a tool through which we can discuss extraordinary claims: Except in light of strong evidence, "ordinary" claims can serve as the default for citizens and scientists making sense of unusual and potentially biased claims.
-Similarly, Bayes Theorem demonstrates that the \ital{simplest possible explanation--the most parsimonious explanation--is often the best explanation}.
-
-In sum, a Bayesian perspective offers a general approach to science literacy that can account for uncertainty in a rigorous and principled manner.
-In addition, a number of principles that follow from Bayes Theorem, such as being receptive to new evidence and weighting that evidence per its strength in light of the strength of one's initial beliefs or understanding could bolster science literacy---and how the public comes to understand and gain trust in science.
-
-\section{A Bayesian Perspective and Science Education}
-
-Research on the role of Bayes in science education has mostly been carried out in the context of undergraduate statistics education, in which there have been several calls to action \parencite{gpkpwc18, h_a20}, examples and design options \parencite{a02, b02,g08, w17, h_j20} and debate \parencite{jrhrr20} over how to advance the place of Bayesian methods in undergraduate statistics and data science degree programs.
-The accessibility of Bayesian methods in undergraduate classes is the result of advances in the necessary (for many uses) computer power \parencite{gpkpwc18} as well as the availability of tools that facilitate Bayesian analysis, especially for newcomers \parencite{ah20}.
-As evidenced by the recent special issue of the \emph{Journal of Statistics Education} of which several of the above-referenced articles are part, the pedagogy of Bayesian statistics is an active area of research in statistics education at the undergraduate level.
-
-There is little research outside of statistics education in the broader science education community.
-Two publications merit mention though.
-The work of \textcite{so12} and \textcite{n11} both applied Bayesian perspectives to the science practice of argumentation.
-Szu and Osborne presented the case for how and why Bayes Theorem can apply to research and practice on scientific reasoning, broadly, and argumentation, particularly.
-This paper represents the most comprehensive account of the relevance of Bayesian methods for science education.
-Szu and Osborne argue that Bayes is useful as both a formal mathematical tool and a conceptual one; indeed, they write that considering the degrees of certainty in beliefs that individual students hold—different from applications of Bayes Theorem that follow more or less deterministically from ``external, objectively probabilistic systems'' (p.
-61)—is ``the key leap that characterizes the debate about the value of Bayesian inference as a model of scientific reasoning'' (p.
-61).
-In this way, Szu and Osborne argue that the greatest use of Bayes Theorem in science classrooms is as a model of informal scientific reasoning, aligning with similar (informal) approaches to inference within the statistics education research community \parencite{batanero2016research, mr18}.
-They offer some research-related backing for the use of Bayes Theorem (e.g., noting how conceptual change research, particularly, and constructivist research, broadly, can both be explained coherently within a Bayesian approach) as well as some practical, instructional recommendations, some of which we detail later in this section.
-
-\textcite{n11} work on Bayesian approaches to argumentation complement the work of Szu and Osborne in that Nussbaum describes both an application of Bayesian methods in K-12 classroom contexts and ideas about how Bayesian methods can serve as an analytic framework for students’ argumentation.
-Concerning the latter, Nussbaum describes how the social issue of raising taxes to provide resources to homeless individuals was a rich context for students to engage in forms of Bayesian reasoning.
-In this application, Nussbaum describes how the prior and likelihood could be obtained from empirical evidence—in this way, illustrating what \textcite{so12} characterize as the more externally objective use of Bayes Theorem—and how the estimates that result from applying Bayes Theorem led students to re-evaluate their initial arguments.
-
-In summary, the work of \textcite{so12} and \textcite{n11} begin to apply Bayesian ideas to science education and other relevant classroom contexts, making Bayesian ideas more vivid in the process.
-Curiously, no work has applied Bayesian methods to another science practice, that of analyzing and interpreting data; one for which Bayesian methods have clear conceptual applicability and utility, but one for which the absence of suitable tools may provide to be a barrier.
-In this way, Bayesian methods could serve as a cross-cutting concept \parencite{nrc12} that unites the practice of argumentation with analyzing and interpreting data and using mathematics and computational thinking.
-
-Bayesian methods have some utility in science education, even if that utility is under-realized at present.
-Next, we present some of the specific ways Bayesian methods could enhance science education—and could do so in a way that makes Bayes more tractable.
-
-First, curricular resources and pedagogical approaches that take a Bayesian perspective can be developed.
-These could elaborate on the ideas of \textcite{so12} and \textcite{n11}.
-At a high-level, Szu and Osborne write that to reason about data in light of their initial beliefs (and to construct a kind of informal likelihood statistic), students “need to see judgments about data and evidence being an assessment not only of the probability of the hypothesis being correct but also of it being wrong” (p.
-67).
-Filling in more details, speaking from both the perspective of a researcher (of argumentation) and an educator (or, a designer of classroom activities), Nussbaum summarizes that using Bayes “forces one to (a) think about an issue from different perspectives, (b) reflect on one’s range of certainty and uncertainty and (c) conduct a simple computer simulation (e.g., calculating the product) that integrates various facets of the issue together and affords further reflection” (p.
-99).
-
-The work of \textcite{so12} and \textcite{n11} can be complemented by finer-grained instructional techniques, especially those advanced by statistics educators to support students to reason about uncertainty probabilistically.
-A benefit of these strategies is that they can help to address some of the ingrained challenges that people and learners face when attempting to reason probabilistically \parencite[e.g., ][]{tk74}.
-Particularly, \textcite{bkbm18} for instance, suggest replacing probabilities with natural frequencies.
-This suggestion is in alignment with prior work by \textcite{gh95} on how different representations of probabilities can explain some of the challenges that people and learners face.
-In addition, \textcite{me14}—building on much earlier experimental work by \textcite{wason1971natural}—provide a simple example of how even fourth-graders, with the right tools to represent probabilities, can support students to transition from logical to probabilistic reasoning.
-These could also support the aim of developing unplugged versions of Bayesian analyses.
-
-Science education researchers may consider how teachers can support students to reason about uncertainty in a way that aligns with a Bayesian perspective.
-While some scholars have contributed foundational work demonstrating how this uncertainty can be productive as a part of teaching practice \parencite{manz2018supporting}, a broad and formal perspective on how uncertainty can be approached in science education is needed.
-Further, a better understanding of uncertainty can help to learn about scientific--and other--concepts where probability plays a key role; it may be that learning about how Bayes rule can be used in legal or medical contexts could impress upon students the cross-cutting nature of probability and uncertainty.
-
-Finally, recently developed tools, particularly the JASP statistical software \parencite{JASP2020}, may be helpful.
-JASP allows for both conventional frequentist as well as Bayesian versions of commonly used statistical tests, such as t-tests and regression analyses.
-There are many accessible guides available for specifying Bayesian versions of analyses \parencite{v20, wagenmakers2018bayesian}, and there is a great potential in building particular examples within JASP (that are aligned to science education curricular standards) that could support science educators to have greater access to the power of Bayesian methods in their classrooms.
-Another possibly useful tool could be Bayesian networks (see Bayes Box; Bayes Box, 2021), which could be used to model complex systems in a way that is similar to other tools used in science education for modeling complex systems, such as \textcite{20}.
-
-\section{Conclusion}
-
-In summary--to this point--we have introduced Bayes as a powerful way of conceptualizing probability and uncertainty.
-These are two concepts that are ubiquitous in the learning, doing, and communicating of science and therefore critical aspects of science literacy.
-However, probability and uncertainty are also infamously challenging concepts—for experts and laypersons alike \parencite{gkv04, s07, tk74}.
-Thus, it is not surprising when the public misinterprets statements about probability and uncertainty but the consequences may be dire when trust in science is eroded, and, sadly, examples of this are plentiful during current the COVID-19, e.g., the efficiency of masks in preventing the spread of COVID and the efficiency of the AstraZeneca vaccine were strongly questioned, delaying efforts to combat the pandemic.
-The public, however, can hardly be blamed for this in light of the prevalent findings that the frequentist conceptualizations of probability and uncertainty in school are challenging and uncommonly align with or build on peoples’ intuitions.
-As research from cognitive and psychological science as well as emerging work from statistics and statistics data science education demonstrates, conceptualizing probability and uncertainty from a Bayesian perspective better aligns with peoples’ intuitions \parencite{kl18, so12} about these concepts and thus has the potential to support learning about probability and uncertainty in science.
-Moreover, a Bayesian perspective is unique in that it is akin to a cross-cutting concept in the sense of a style of reasoning \parencite{so12, ork18} that blends conceptual ideas with practices: Bayes is both a philosophical and psychological perspective on how to learn from data as well as a statistical framework.
-Indeed, a key potential benefit of Bayes may be that it provides a framework for helping students to transition from their conceptual understanding of scientific ideas to quantitatively analyzing data.
-
-Even at a time in which science has made and continues to make great advances, the history of science suggests that uncertainty will not be removed (or even reduced) given scientific progress \parencite{fara2010science}; looking ahead, being able to effectively navigate, counsel, and act in the midst of uncertainty may be as important as it has been at other uncertain points in history.
-Like the surprising utility of a Bayesian perspective in domains that are similar to education (e.g., developmental science), might an unexpected feature of a Bayesian perspective be that embracing uncertainty--equipped with the tools that Bayesian methods provide--makes individuals and societies more confident?
-In short, can we build trust in science by embracing uncertainty?
-We think that these are questions that are \ital{likely} to yield positive results for science educators and those advocating for a more rationale and scientific citizenry and for science learners.
-
-\printbibliography
-
-\end{document}
+\\

### 2021-09-22 old-text/science-literacy.tex
updates for revision

\section{How Bayesian Principles Could Bolster Science Literacy}

We conceive that Bayesian methods can enhance \ital{science literacy}, capabilities range from core \ital{literacies}, including numeracy and textual literacy (what is mostly thought of in common parlance concerning literacy), content knowledge, and epistemic knowledge \parencite{council2016science}. The specific ways we envision a Bayesian perspective bolstering science literacy are also techniques that teachers can appropriate in their planning and teaching; they may already align with how they teach about science through engaging in scientific practices \parencite{nrc12} or how they articulate the nature of science \parencite{lederman2020avoiding}. \\

Consider a salient topic as a context to apply these principles, medical vaccines for diseases, such as the disease caused by the coronavirus SARS-CoV-2, COVID-19. Most evidence suggests that vaccines have been safe, historically, and are generally safe, at present (Centers for Disease Control and Prevention, 2021). Nevertheless, they are not perfectly safe, as unlikely, adverse outcomes do happen, and there may be adverse side effects from some vaccines that research has not yet revealed. In light of this evidence, consider the development of a new vaccine, such as those available at the time of this writing for COVID-19. Some use long-established methods of vaccine development (namely, deactivating SARS-CoV-2 particles but retaining cell signaling proteins to which humans develop an immune response—and, after, a degree of immunity) and injecting these deactivated virus particles into people), whereas others use a new mechanism (namely, injecting messenger RNA into people; this messenger RNA is then transcribed by cells in people’s bodies to produce a particular protein common to SARS-CoV-2 particles, which human’s bodies then learn to recognize to develop immunity. We note that--at present--diseases for which it has been difficult to develop effective vaccines, including Malaria, may have viable mRNA vaccines in the near future \parencite{datoo2021high}. \\

Before any data from clinical trials, it would be reasonable to consider these vaccines to probably be safe in light of the safety of other vaccines (and based on similarities in their mechanisms of action), but to have a strong degree of caution about how safe they are. After initial trials were completed, the degree of safety that individuals associate with these vaccines—in light of the prior and the limited evidence—could increase, but still not be near the level of safety that many would consider acceptable. Following the completion of clinical trials that show minimal side effects (holding aside their efficacy), one’s estimates of the safety of these vaccines could increase by a substantial amount, but concerns could remain, particularly for messenger RNA vaccines, which are newer, and may, plausibly, have unintended side effects. \\

While the weighting of the evidence in light of prior research and theoretical accounts above may be natural to a scientist, it is challenging for many. In the following, we showcase how several important insights for weighting evidence follow directly from Bayes' rule. We do not think or expect that members of the public can do this, but members of the public (and learners) do have viable ideas and hypotheses about how the world works. Thus, we argue that emphasizing Bayesian reasoning can support science literacy by providing people with a principled way to weight evidence that is accessible because it builds on peoples' prior ideas. \\

In the context of the above example on SARS-CoV-2, the result of applying Bayes Theorem in a heuristic mode is a non-binary inference, one that doesn't speak to whether or not the vaccine is safe, but to what degree it is safe, doing so with a language of probability and uncertainty. Though a small difference, discussing the safety of vaccines on these terms can provide a more solid foundation for principled disagreements and arguments. In addition, Bayes provides a heuristic and a language for engaging with uncertainty in a science literacy context in a principled way that is in line with how most of us likely already interpret data \parencite{tgk06, gw12, gh95}; A Bayesian perspective can provide people with a way to talk about weighing between ideas and their prior beliefs and new information and data. In short, Bayes gives people a chance to deal with uncertainty in a principled but accessible manner. \\

In addition, many themes introduced in the section on the epistemological principles that follow from Bayes Theorem are relevant to the aim of supporting science literacy. First, \ital{even individuals with different initial beliefs may be brought closer to an agreement in light of data} that serves as evidence for a particular belief or hypothesis. In this way, Bayes Theorem illustrates--at the conceptual level as well as statistically--how individuals may be unlikely to agree based on the limited or weak information, but that this can shift in light of accumulating data. As we noted earlier, this is far from a deterministic process, and \ital{holding initial beliefs that a hypothesis has a probability of zero or one means that data does not affect one's beliefs}. This suggests that a cultural norm and principle for science literacy can be to always reserve a degree of uncertainty such that one could be influenced by (new) data or evidence: From a Bayesian perspective, \ital{our knowledge is uncertain and open to change and revision} so long as we remain open to new evidence. \\

Related to the importance of not holding scientific beliefs with a probability of zero or one, another principle that follows from Bayes Theorem is that \ital{extraordinary claims require extraordinary evidence}, and that unlikely or implausible hypotheses or claims are likely to necessitate strong empirical data to be convincing to others. A corollary of this principle is that we should strong initial or prior beliefs can serve as a tool through which we can discuss extraordinary claims: Except in light of strong evidence, "ordinary" claims can serve as the default for citizens and scientists making sense of unusual and potentially biased claims. Similarly, Bayes Theorem demonstrates that the \ital{simplest possible explanation--the most parsimonious explanation--is often the best explanation}. \\

In sum, a Bayesian perspective offers a general approach to science literacy that can account for uncertainty in a rigorous and principled manner. In addition, a number of principles that follow from Bayes Theorem, such as being receptive to new evidence and weighting that evidence per its strength in light of the strength of one's initial beliefs or understanding could bolster science literacy--and how the public comes to understand and gain trust in science. \\
